### 2024-10-24 template/edumer_template.qmd
Create template first version

---
pagetitle: "EDUMER Template"
Author: Equipo EDUMER
bibliography: "../input/bib/refs.bib"
csl: "../input/bib/apa6.csl"
lang: es

knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
    R.options:
      knitr.graphics.auto_pdf: true

format:
  revealjs:
    #logo: images/edumer.png
    slide-number: true
    theme: "../libs/edumer.scss"
    auto-stretch: false
    title-slide-attributes:
      visibility: false
    transition: fade
    transition-speed: slow
# data-background-image: images/cover.jpg
# data-background-size: cover
    auto-play-media: true
  pdf:
    format: beamer 
    keep-tex: true  # Debug LaTeX output if needed
---

```{r}
#| label: setup
#| include: false
library(knitr)
knitr::opts_chunk$set(echo = F,
                      warning = F,
                      error = F, 
                      message = F) 
```

```{r}
#| label: packages
#| include: false

if (! require("pacman")) install.packages("pacman")

pacman::p_load(tidyverse, 
               sjmisc, 
               sjPlot, 
               sjlabelled, 
               here,
               kableExtra,
               ggdist,
               summarytools)

options(scipen=999)
rm(list = ls())
```

```{r}
#| label: data
#| include: false

load(url("https://github.com/educacion-meritocracia/market-justice-preferences-school/raw/refs/heads/main/input/data/mlm_models.RData"))
load(url("https://github.com/educacion-meritocracia/market-justice-preferences-school/raw/refs/heads/main/input/data/dependientes.RData"))

```

::: columns
::: {.column width="15%"}

![](images/coes.png)

![](images/edumer.png)

:::

::: {.column .column-right width="85%"}
<br>

## **Mi increíble presentación**

------------------------------------------------------------------------

**Autores**

##### Afiliaciones

::: {.red2 .medium}

**FONDECYT No. 1210847 "Meritocracia en la escuela: fundamentos morales del mercado educativo y sus implicancias para la formación ciudadana en Chile"**

:::

Conferencia X

:::
:::

::: {.notes}
Aquí mis notas
:::

# Contexto y motivación {.xlarge}

## Antecedentes

::: {.incremental .highlight-last style="font-size: 100%;"}

- The justification of social inequality based on market-type criteria has been conceptualized as the individuals' adherence to the deservingness of social goods and services (such as health, education, and pensions) based on prices and individuals' ability to pay [@boltanski_new_2005; @lane_market_1986; @streeck_citizens_2012].

- Research on social stratification beliefs [@kluegel_beliefs_1987], highlights that people's explanations and justifications of social inequality are closely tied to their judgments of deservingness.

-  The expansion of marketization has been related to a larger justification of market mechanisms, whereby societies with larger private spending on services show larger market justice preferences [@lindh_public_2015].

:::

## Antecedentes

::::: columns
::: {.column width="50%" .content-box-red}

#### A

- The original definition of merit is a combination of effort and talent [@young_rise_1958]
- 2
- 3

:::

::: {.column width="50%" .content-box-yellow}

#### B

- Adolescents from low-socioeconomic-status families are likelier to have an egalitarian fairness view and consider an equal distribution as fair in a situation with unequal merits [@almas_fairness_2017]
- 2
- 3

:::
:::::

## Hipótesis

::: {.incremental .highlight-last style="font-size: 100%;"}

$H_{1}$

$H_{2}$

$H_{3}$

$H_{n}$

:::

# Datos, variables y métodos {.xlarge}

## Datos

::: {.incremental .highlight-last style="font-size: 100%;"}
-   Datos
:::

## Variables

```{r}
#| label: tbl-desc-dependientes
#| tbl-cap: "Variables dependientes"
#| results: asis
#| echo: false

df<-dfSummary(dependientes,
               plain.ascii = FALSE,
               style = "multiline",
               tmp.img.dir = "/tmp",
               headings = F,  # encabezado
               varnumbers = F, # num variable
               labels.col = T, # etiquetas
               na.col = F,    # missing
               graph.col = F, # plot
               valid.col = T) # n valido

df$Variable <- NULL # delete variable column

print(df)
```

:::

## Métodos

- Métodos

# Resultados {.xlarge}

# Modelos
::: {style="font-size: 100%;"}

```{r}
#| label: fig-lineal-reg
#| fig-cap: "Linear mixed-effects models for meritocracy perception and market justice preferences"
#| fig-cap-location: top
#| fig-align: center
#| out-width: '80%'
#| echo: false
#| warning: false
#| message: false
#| results: asis

theme_set(theme_ggdist())

sjPlot::plot_model(model = reg1_bienestar, show.values = T, show.p = T)

```

:::

# Discusión y conclusiones {data-background-color="#F2A0A7"}

# Discusión y conclusiones {data-background-color="#F2A0A7"}

::: { .incremental .highlight-last style="font-size: 100%;"}

- Discusión

  - Conclusión
:::

# Gracias por su atención!

-   **Github del proyecto:** <https://github.com/educacion-meritocracia>

## Referencias