### 2025-01-29 processing/01-proc-data.qmd
Up primeros ajustes

---
title: "Data preparation"
subtitle: "Perceptions of Inequality and Meritocracy: Their Interplay in Shaping Preferences for Market Justice in Chile (2016-2023)"
author: "Andreas Laffert, Research asistant"
date: today
lang: en
fontsize: 12pt
format:
  html:
    toc: true
    toc-location: right
    toc-depth: 2
    toc-expand: 2
    toc-title: Contents
    number-sections: true
    number-depth: 3
    theme:
      - cosmo
      - edumer_html.scss
    code-link: true
    title-block-banner: true
  pdf:
    number-sections: true
    number-depth: 3
editor_options: 
  chunk_output_type: console
---

# Presentation

This is the data preparation code for the paper "Perceptions of Inequality and Meritocracy: Their Interplay in Shaping Preferences for Market Justice in Chile (2016-2023)". The prepared dataset is `ELSOC_Long_2016_2023_1.00.RData`.

```{r}
#| label: set
#| echo: false
#| message: false
#| warning: false

library(knitr)
knitr::opts_chunk$set(echo = TRUE, include = TRUE, warning = FALSE, message = FALSE)

table_format <- if(is_html_output()) {
  "html"
} else if(is_latex_output()) {
  "latex"
}
table_format2 <- if(is_html_output()) {
  T
} else if(is_latex_output()) {
  F
}

options(kableExtra.html.bsTable = T)
options(knitr.kable.NA = "")
```

# Libraries

```{r}
#| label: packages
#| include: false

if (! require("pacman")) install.packages("pacman")

pacman::p_load(tidyverse,
               sjmisc, 
               here,
               sjlabelled)

options(scipen=999)
rm(list = ls())
```

# Data

```{r}
#| label: data
#| echo: true
#| output: false

load(file = here("input", "data", "original", "ELSOC_Long_2016_2023_1.00.RData"))

glimpse(elsoc_long_2016_2023)

```

# Processing

## Select

## Filter

## Recode and transform

## Missing values

# Save and export