### 2025-03-04 processing/01-proc-data.qmd
Up data preparation and analysis V1

---
title: "Data preparation"
subtitle: "Perceptions of Inequality and Meritocracy: Their Interplay in Shaping Preferences for Market Justice in Chile (2016-2023)"
author: "Andreas Laffert, Research asistant"
date: today
lang: en
fontsize: 12pt
format:
  html:
    toc: true
    toc-location: right
    toc-depth: 2
    toc-expand: 2
    toc-title: Contents
    number-sections: true
    number-depth: 3
    theme:
      - cosmo
      - edumer_html.scss
    code-link: true
    title-block-banner: true
  pdf:
    number-sections: true
    number-depth: 3
editor_options: 
  chunk_output_type: console
---

# Presentation

This is the data preparation code for the paper "Perceptions of Inequality and Meritocracy: Their Interplay in Shaping Preferences for Market Justice in Chile (2016-2023)". The prepared dataset is `ELSOC_Long_2016_2023_1.00.RData`.

```{r}
#| label: set
#| echo: false
#| message: false
#| warning: false

library(knitr)
knitr::opts_chunk$set(echo = TRUE, include = TRUE, warning = FALSE, message = FALSE)

table_format <- if(is_html_output()) {
  "html"
} else if(is_latex_output()) {
  "latex"
}
table_format2 <- if(is_html_output()) {
  T
} else if(is_latex_output()) {
  F
}

options(kableExtra.html.bsTable = T)
options(knitr.kable.NA = "")
```

# Libraries

```{r}
#| label: packages
#| include: true

if (! require("pacman")) install.packages("pacman")

pacman::p_load(tidyverse,
               sjmisc, 
               here,
               sjlabelled,
               SciViews,
               naniar)

options(scipen=999)
rm(list = ls())
```

# Data

```{r}
#| label: data
#| echo: true
#| output: false

load(url("https://dataverse.harvard.edu/api/access/datafile/10797987"))

glimpse(elsoc_long_2016_2023)

```

# Processing

## Select

```{r}
db_proc <- elsoc_long_2016_2023 %>% 
  select(idencuesta, ola, 
         ponderador_long_total,segmento, estrato,
         just_educ = d02_02, just_pension = d02_01, 
         just_health = d02_03, merit_effort = c18_09,
         merit_talent = c18_10, perc_sal_gerente = d03_01,
         perc_sal_obrero = d03_02, just_sal_gerente = d04_01, 
         just_sal_obrero = d04_02, age = m0_edad, m01, 
         sex = m0_sexo, ess = d01_01, ideo = c15
         ) %>% 
  as_tibble()
```

## Filter

```{r}
#db_proc %>% 
#  group_by(ola) %>% 
#  count(just_educ,just_pension,just_health) %>% 
#  na.omit() %>% 
#  print(n = nrow(.))

db_proc <- db_proc %>% dplyr::filter(ola %in% c(1,2,3,4,6,7))

```

## Recode and transform

```{r}
# general na's
db_proc[db_proc ==-999] <- NA
db_proc[db_proc ==-888] <- NA
db_proc[db_proc ==-777] <- NA
db_proc[db_proc ==-666] <- NA

# mjp

frq(db_proc$just_health)
frq(db_proc$just_pension)
frq(db_proc$just_educ)

labels1 <- c("Strongly desagree" = 1, 
             "Desagree" = 2, 
             "Neither agree nor desagree" = 3, 
             "Agree" = 4, 
             "Strongly agree" = 5)

db_proc <- db_proc %>% 
  mutate(
    across(
      .cols = c(just_health, just_pension, just_educ),
      .fns = ~ sjlabelled::set_labels(., labels = labels1)
      )
    )

db_proc <- cbind(db_proc, "mjp" = rowMeans(db_proc %>% select(just_health, just_pension, just_educ), na.rm=TRUE))

summary(db_proc$mjp)

db_proc <- db_proc %>% 
  mutate(
    across(
      .cols = c(just_health, just_pension, just_educ),
      .fns = ~ car::recode(., recodes = c("1='Strongly desagree'; 2='Desagree';
                                          3='Neither agree nor desagree'; 4='Agree';
                                          5='Strongly agree'"), 
                           levels = c("Strongly desagree", "Desagree", "Neither agree nor desagree", "Agree", "Strongly agree"),
                           as.factor = T)
    )
  )

db_proc$just_health <- sjlabelled::set_label(db_proc$just_health, 
                        label = "Health distributive justice")

db_proc$just_pension <- sjlabelled::set_label(db_proc$just_pension, 
                        label = "Pension distributive justice")

db_proc$just_educ <- sjlabelled::set_label(db_proc$just_educ, 
                        label = "Education distributive justice")

db_proc$mjp <- sjlabelled::set_label(db_proc$mjp, 
                        label = "Market justice preferences")

# merit

frq(db_proc$merit_effort)
frq(db_proc$merit_talent)

db_proc <- db_proc %>% 
  mutate(
    across(
      .cols = c(merit_effort, merit_talent),
      .fns = ~ sjlabelled::set_labels(., labels = labels1)
      )
    )

db_proc <- db_proc %>% 
  mutate(
    across(
      .cols = c(merit_effort, merit_talent),
      .fns = ~ car::recode(., recodes = c("1='Strongly desagree'; 2='Desagree';
                                          3='Neither agree nor desagree'; 4='Agree';
                                          5='Strongly agree'"), 
                           levels = c("Strongly desagree", "Desagree", "Neither agree nor desagree", "Agree", "Strongly agree"),
                           as.factor = T)
    )
  )

db_proc$merit_effort <- sjlabelled::set_label(db_proc$merit_effort, 
                        label = "People are rewarded for their efforts")

db_proc$merit_talent <- sjlabelled::set_label(db_proc$merit_talent, 
                        label = "People are rewarded for their intelligence")

# perc inequality gap

frq(db_proc$perc_sal_gerente)
frq(db_proc$perc_sal_obrero)

db_proc <- db_proc %>% 
  mutate(
    across(
      .cols = c(perc_sal_gerente, perc_sal_obrero),
           .fns = ~ sjlabelled::set_na(., na = 0)
      )
    )

db_proc$perc_inequality <- SciViews::ln(db_proc$perc_sal_gerente/db_proc$perc_sal_obrero)

summary(db_proc$perc_inequality)

db_proc$perc_inequality <- sjlabelled::set_label(db_proc$perc_inequality, 
                        label = "Inequality gap perception")

# just inequality gap

frq(db_proc$just_sal_gerente)
frq(db_proc$just_sal_obrero)

db_proc <- db_proc %>% 
  mutate(
    across(
      .cols = c(just_sal_gerente, just_sal_obrero),
           .fns = ~ sjlabelled::set_na(., na = 0)
      )
    )

db_proc$just_inequality <- SciViews::ln(db_proc$just_sal_gerente/db_proc$just_sal_obrero)

summary(db_proc$just_inequality)

db_proc$perc_inequality <- sjlabelled::set_label(db_proc$perc_inequality, 
                        label = "Inequality gap justification")

# sex
db_proc$sex <- car::recode(db_proc$sex, 
                           recodes = c("1='Male'; 2='Female'"), 
                           levels = c("Male", "Female"),
                           as.factor = T)

db_proc$sex <- sjlabelled::set_label(db_proc$sex, 
                        label = "Gender")

# wave 
db_proc <- db_proc %>% 
  rename(wave = ola) %>% 
  mutate(wave = case_when(wave == 1 ~ 1,
                          wave == 2 ~ 2,
                          wave == 3 ~ 3,
                          wave == 4 ~ 4,
                          wave == 6 ~ 5,
                          wave == 7 ~ 6,
                          TRUE ~ NA_real_))

db_proc$wave <- sjlabelled::set_label(db_proc$wave, 
                        label = "Wave")

# age
frq(db_proc$age)

db_proc$age <- 
  factor(car::recode(db_proc$age, 
                     "18:29=1;30:49=2;50:64=3;65:150=4"),
         labels = c('18-29', '30-49', '50-64', '65 or more'))
db_proc$age <-
  sjlabelled::set_label(db_proc$age, 
                        label = c("Age groups"))

# political indentification

frq(db_proc$ideo)

db_proc$ideo<-
factor(
  car::recode(
    db_proc$ideo,
    "c(11,12,-888,-999)='Does not identify';c(0,1,2,3,4)='Left';
     c(5)='Center';c(6,7,8,9,10)='Right'"
  ),
  levels = c('Left', 'Center', 'Right', 'Does not identify')
)

db_proc$ideo<- factor(db_proc$ideo,levels = levels(db_proc$ideo))

db_proc$ideo <- 
sjlabelled::set_label(x = db_proc$ideo, 
                      label = "Political identification")

frq(db_proc$ideo)

# ess
sjmisc::frq(db_proc$ess)

db_proc$ess <- as.numeric(db_proc$ess)

db_proc$ess <-  sjlabelled::set_label(x = db_proc$ess,
                      label = "Subjective Social Status")

# educ
frq(db_proc$m01)

db_proc$educ <- 
  car::recode(db_proc$m01,
              "c(1,2,3,4,5,6,7)=1;c(8,9,10)=2; c(-888,-999)=NA")
db_proc$educ <-
  factor(db_proc$educ,
         labels = c("Less than Universitary","Universitary"))

db_proc$educ <- 
sjlabelled::set_label(x = db_proc$educ,
                      label = "Education")

#Recoding of education to years based on casen 2017.
db_proc$educyear<- as.numeric(
  car::recode(db_proc$m01, 
              "1=0;2=4.3;3=7.5;4=9.8;5=12.02;6=13.9;
               7=14.8;8=14.9;9=16.9;10=19.07;c(-888,-999)=NA", 
              as.numeric = T))

db_proc$educyear <- 
sjlabelled::set_label(x = db_proc$educyear,
                      label = "Education in years")

class(db_proc$educyear)
frq(db_proc$educyear)
```

## Missing values

```{r}
colSums(is.na(db_proc))

n_miss(db_proc)

prop_miss(db_proc)*100

miss_var_summary(db_proc)

miss_var_table(db_proc)

vis_miss(db_proc) + theme(axis.text.x = element_text(angle=80))

```

# Save and export

```{r}
db_proc <- db_proc %>% 
  select(-c(perc_sal_gerente, perc_sal_obrero, just_sal_gerente, just_sal_obrero, m01))

df1 <- db_proc %>% na.omit()

df1 <- sjlabelled::copy_labels(df_new = df1, df_origin = db_proc)

save(df1, file = here("input/data/proc/df1.RData"))
```