### 2025-03-04 processing/02-analysis.qmd

---
title: "Data analysis"
subtitle: "Perceptions of Inequality and Meritocracy: Their Interplay in Shaping Preferences for Market Justice in Chile (2016-2023)"
author: "Andreas Laffert, Research asistant"
date: today
lang: en
fontsize: 12pt
format:
  html:
    code-fold: true
    code-summary: "Show the code"
    toc: true
    toc-location: right
    toc-depth: 2
    toc-expand: 2
    toc-title: Contents
    number-sections: true
    number-depth: 3
    theme:
      - cosmo
      - edumer_html.scss
    code-link: true
    title-block-banner: true
  pdf:
    number-sections: true
    number-depth: 3
editor_options: 
  chunk_output_type: console
---

# Presentation

This is the analysis code for the paper "Perceptions of Inequality and Meritocracy: Their Interplay in Shaping Preferences for Market Justice in Chile (2016-2023)". The dataset used is `df1.RData`.

```{r}
#| label: set
#| echo: false
#| message: false
#| warning: false

library(knitr)
knitr::opts_chunk$set(echo = TRUE, include = TRUE, warning = FALSE, message = FALSE)

table_format <- if(is_html_output()) {
  "html"
} else if(is_latex_output()) {
  "latex"
}
table_format2 <- if(is_html_output()) {
  T
} else if(is_latex_output()) {
  F
}

options(kableExtra.html.bsTable = T)
options(knitr.kable.NA = "")
```

# Libraries

```{r}
#| label: packages
#| include: true

if (! require("pacman")) install.packages("pacman")

pacman::p_load(tidyverse, 
               sjmisc, 
               sjPlot, 
               lme4, 
               here, 
               performance,
               influence.ME, 
               marginaleffects,
               MLMusingR,
               texreg, 
               ggdist,
               misty,
               kableExtra,
               ggalluvial, 
               shadowtext,
               MetBrewer,
               patchwork,
               sjlabelled)

options(scipen=999)
rm(list = ls())
```

# Data

```{r}
#| label: data
#| echo: true

load(file = here("input/data/proc/df1.RData"))

glimpse(df1)

```

# Analysis

## Descriptives

```{r}
#| label: tbl-summary
#| tbl-cap: "Estadísticos descriptivos"
#| tbl-cap-location: top
t1 <- df1 %>% 
  select(just_health, just_pension, just_educ, mjp, perc_inequality, just_inequality, merit_effort, merit_talent, ess, ideo, educ, sex, age)

print(summarytools::dfSummary(t1), method="render")
```

```{r}
#| label: fig-alluvial
#| out-width: '100%'
#| fig-asp: 1
#| fig-cap: "Change in the justification of inequality in health, pensions and education over time (2016-2023)"
#| fig-cap-location: top

datos.health <- df1 %>% 
  group_by(idencuesta, wave) %>% 
  count(just_health) %>% 
  group_by(wave) %>% 
  mutate(porcentaje=n/sum(n)) %>% 
  ungroup() %>% 
  na.omit() %>% 
  mutate(wave = case_when(wave == 1 ~ "2016",
                          wave == 2 ~ "2017",
                          wave == 3 ~ "2018",
                          wave == 4 ~ "2019",
                          wave == 5 ~ "2022",
                          wave == 6 ~ "2023"),
         wave = factor(wave, levels = c("2016",
                                      "2017",
                                      "2018",
                                      "2019",
                                      "2022",
                                      "2023")))

etiquetas.health <- df1 %>%
  group_by(wave, just_health) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(wave) %>%
  mutate(porcentaje = count / sum(count)) %>% 
  na.omit() %>% 
  mutate(idencuesta = 1,
         wave = case_when(wave == 1 ~ "2016",
                          wave == 2 ~ "2017",
                          wave == 3 ~ "2018",
                          wave == 4 ~ "2019",
                          wave == 5 ~ "2022",
                          wave == 6 ~ "2023"),
         wave = factor(wave, levels = c("2016",
                                        "2017",
                                        "2018",
                                        "2019",
                                        "2022",
                                        "2023")))

p1 <- datos.health %>% 
  ggplot(aes(x = wave, fill = just_health, stratum = just_health,
             alluvium = idencuesta, y = porcentaje)) +
  ggalluvial::geom_flow(alpha = .4) + 
  ggalluvial::geom_stratum(linetype = 0) +
  scale_y_continuous(labels = scales::percent) + 
  scale_fill_manual(values =  c("#CA0020","#F4A582","#b3b3b3ff","#92C5DE","#0571B0")) +
  geom_shadowtext(data = etiquetas.health,
                  aes(label = ifelse(porcentaje > 0 , scales::percent(porcentaje, accuracy = .1),"")),
                  position = position_stack(vjust = .5),
                  show.legend = FALSE,
                  size = 3,
                  color = rep('white'),
                  bg.colour='grey30')+
  labs(y = "%",
       x = NULL,
       fill = NULL,
       title = "a. Health") +
  theme_ggdist() +
  theme(legend.position = "none")

datos.pension <- df1 %>% 
  group_by(idencuesta, wave) %>% 
  count(just_pension) %>% 
  group_by(wave) %>% 
  mutate(porcentaje=n/sum(n)) %>% 
  ungroup() %>% 
  na.omit() %>% 
  mutate(wave = case_when(wave == 1 ~ "2016",
                          wave == 2 ~ "2017",
                          wave == 3 ~ "2018",
                          wave == 4 ~ "2019",
                          wave == 5 ~ "2022",
                          wave == 6 ~ "2023"),
         wave = factor(wave, levels = c("2016",
                                        "2017",
                                        "2018",
                                        "2019",
                                        "2022",
                                        "2023")))

etiquetas.pension <- df1 %>%
  group_by(wave, just_pension) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(wave) %>%
  mutate(porcentaje = count / sum(count)) %>% 
  na.omit() %>% 
  mutate(idencuesta = 1,
         wave = case_when(wave == 1 ~ "2016",
                          wave == 2 ~ "2017",
                          wave == 3 ~ "2018",
                          wave == 4 ~ "2019",
                          wave == 5 ~ "2022",
                          wave == 6 ~ "2023"),
         wave = factor(wave, levels = c("2016",
                                        "2017",
                                        "2018",
                                        "2019",
                                        "2022",
                                        "2023")))

p2 <- datos.pension %>% 
  ggplot(aes(x = wave, fill = just_pension, stratum = just_pension,
             alluvium = idencuesta, y = porcentaje)) +
  ggalluvial::geom_flow(alpha = .4) + 
  ggalluvial::geom_stratum(linetype = 0) +
  scale_y_continuous(labels = scales::percent) + 
  scale_fill_manual(values =  c("#CA0020","#F4A582","#b3b3b3ff","#92C5DE","#0571B0")) +
  geom_shadowtext(data = etiquetas.pension,
                  aes(label = ifelse(porcentaje > 0 , scales::percent(porcentaje, accuracy = .1),"")),
                  position = position_stack(vjust = .5),
                  show.legend = FALSE,
                  size = 3,
                  color = rep('white'),
                  bg.colour='grey30')+
  labs(y = "%",
       x = NULL,
       fill = NULL,
       title = "b. Pension") +
  theme_ggdist() +
  theme(legend.position = "none")

datos.educ <- df1 %>% 
  group_by(idencuesta, wave) %>% 
  count(just_educ) %>% 
  group_by(wave) %>% 
  mutate(porcentaje=n/sum(n)) %>% 
  ungroup() %>% 
  na.omit() %>% 
  mutate(wave = case_when(wave == 1 ~ "2016",
                          wave == 2 ~ "2017",
                          wave == 3 ~ "2018",
                          wave == 4 ~ "2019",
                          wave == 5 ~ "2022",
                          wave == 6 ~ "2023"),
         wave = factor(wave, levels = c("2016",
                                        "2017",
                                        "2018",
                                        "2019",
                                        "2022",
                                        "2023")))

etiquetas.educ <- df1 %>%
  group_by(wave, just_educ) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(wave) %>%
  mutate(porcentaje = count / sum(count)) %>% 
  na.omit() %>% 
  mutate(idencuesta = 1,
         wave = case_when(wave == 1 ~ "2016",
                          wave == 2 ~ "2017",
                          wave == 3 ~ "2018",
                          wave == 4 ~ "2019",
                          wave == 5 ~ "2022",
                          wave == 6 ~ "2023"),
         wave = factor(wave, levels = c("2016",
                                        "2017",
                                        "2018",
                                        "2019",
                                        "2022",
                                        "2023")))

p3 <- datos.educ %>% 
  ggplot(aes(x = wave, fill = just_educ, stratum = just_educ,
             alluvium = idencuesta, y = porcentaje)) +
  ggalluvial::geom_flow(alpha = .4) + 
  ggalluvial::geom_stratum(linetype = 0) +
  scale_y_continuous(labels = scales::percent) + 
  scale_fill_manual(values =  c("#CA0020","#F4A582","#b3b3b3ff","#92C5DE","#0571B0")) +
  geom_shadowtext(data = etiquetas.educ,
                  aes(label = ifelse(porcentaje > 0 , scales::percent(porcentaje, accuracy = .1),"")),
                  position = position_stack(vjust = .5),
                  show.legend = FALSE,
                  size = 3,
                  color = rep('white'),
                  bg.colour='grey30')+
  labs(y = "%",
       x = NULL,
       fill = NULL,
       title = "c. Education") +
  theme_ggdist() +
  theme(legend.position = "bottom")

(p1 / p2 / p3)

```

## Correlations

```{r}
df1 %>% 
  select(mjp, perc_inequality, just_inequality, merit_effort, merit_talent) %>% 
  mutate_all(.funs = ~ as.numeric(.)) %>% 
  sjPlot::tab_corr(., triangle = "lower")
```

## Longitudinal multilevel models

```{r}
m0 <- lmer(mjp ~ 1 + (1 | idencuesta), 
                data = df1)

performance::icc(m0, by_group = T)
```

```{r}
df1$merit_effort <- as_numeric(df1$merit_effort)
df1$merit_talent <- as_numeric(df1$merit_talent)
df1 <- df1 %>% 
  mutate(wave_f = case_when(wave == 1 ~ "2016",
                          wave == 2 ~ "2017",
                          wave == 3 ~ "2018",
                          wave == 4 ~ "2019",
                          wave == 5 ~ "2022",
                          wave == 6 ~ "2023"),
         wave_f = factor(wave_f, levels = c("2016",
                                        "2017",
                                        "2018",
                                        "2019",
                                        "2022",
                                        "2023")))

df1 <- df1 %>% 
  group_by(idencuesta) %>% 
  mutate(perc_inequality_mean = mean(perc_inequality, na.rm = T),
         perc_inequality_cwc = perc_inequality - perc_inequality_mean,
         just_inequality_mean = mean(just_inequality, na.rm = T),
         just_inequality_cwc = just_inequality - just_inequality_mean,
         merit_effort_mean = mean(merit_effort, na.rm = T),
         merit_effort_cwc = merit_effort - merit_effort_mean,
         merit_talent_mean = mean(merit_talent, na.rm = T),
         merit_talent_cwc = merit_talent - merit_talent_mean,
         ) %>% 
  ungroup()

m1 <- lmer(mjp ~ 1 + perc_inequality + (1 | idencuesta),
                data = df1)

m2 <- lmer(mjp ~ 1 + just_inequality + (1 | idencuesta),
                data = df1)

m3 <- lmer(mjp ~ 1 + perc_inequality + just_inequality + (1 | idencuesta),
                data = df1)

m4 <- lmer(mjp ~ 1 + perc_inequality + just_inequality + 
             merit_effort +  (1 | idencuesta),
                data = df1)

m5 <- lmer(mjp ~ 1 + perc_inequality + just_inequality + 
             merit_effort + merit_talent + (1 | idencuesta),
                data = df1)

m6 <- lmer(mjp ~ 1 + perc_inequality + just_inequality + 
             merit_effort + merit_talent + perc_inequality_mean +
             just_inequality_mean + merit_effort_mean + 
             merit_talent_mean + (1 | idencuesta),
                data = df1)

m7 <- lmer(mjp ~ 1 + perc_inequality + just_inequality + 
             merit_effort + merit_talent + perc_inequality_mean +
             just_inequality_mean + merit_effort_mean + 
             merit_talent_mean + age + sex + ess + ideo + educ + wave_f + (1 | idencuesta),
                data = df1)
```

```{r}
#| label: tbl-modelos
#| results: asis
#| tbl-cap: "Longitudinal multilevel models for market justice preferences"
#| tbl-cap-location: top

ccoef <- list(
  "(Intercept)" = "Intercept",
  perc_inequality = "Perception inequality (WE)",
  just_inequality = "Justification inequality (WE)",
  merit_effort = "Merit: Effort (WE)",
  merit_talent = "Merit: Talent (WE)",
  perc_inequality_mean = "Perception inequality (BE)",
  just_inequality_mean = "Justification inequality (BE)",
  merit_effort_mean = "Merit: Effort (BE)",
  merit_talent_mean = "Merit: Talent (BE)",
  "age30-49" = "30-49",
  "age50-64" = "50-64",
  "age65 or more" = "65 or more",
  sexFemale = "Female (Ref.= Male)",
  ess = "Subjective social status",
  ideoCenter = "Center",
  ideoRight = "Right",
  "ideoDoes not identify" = "Does not identify",
  educUniversitary = "Universitary (Ref.= Less than universitary)",
  wave_f2017 = "2017",
  wave_f2018 = "2018",
  wave_f2019 = "2019",
  wave_f2022 = "2022",
  wave_f2023 = "2023"
  )

texreg::htmlreg(list(m1, m2, m3, m4, m5, m6, m7),
               custom.model.names = c(paste0("Model ", seq(1:7))),
               caption = NULL,
               stars = c(0.05, 0.01, 0.001),
               custom.coef.map = ccoef,
               digits = 3,
               groups = list("Age (Ref. = 18-29)" = 10:12,"Political identification (Ref.= Left)" = 15:17, "Wave (Ref.= 2016)" = 19:23),
               custom.note = "\\item Note: Cells contain regression coefficients with standard errors in parentheses. %stars. \\\\ \\item Source: own elaboration with pooled data from ELSOC 2016-2023 (n = 16219).",
               threeparttable = T,
               leading.zero = T,
               float.pos = "h!",
               use.packages = F,
               booktabs = T,
               scalebox = 1)
```