### 2025-03-05 processing/01-proc-data.qmd
Update analysis with income and gaps

---
title: "Data preparation"
subtitle: "Perceptions of Inequality and Meritocracy: Their Interplay in Shaping Preferences for Market Justice in Chile (2016-2023)"
author: "Researcher"
date: today
lang: en
fontsize: 12pt
format:
  html:
    toc: true
    toc-location: right
    toc-depth: 2
    toc-expand: 2
    toc-title: Contents
    number-sections: true
    number-depth: 3
    theme:
      - cosmo
      - edumer_html.scss
    code-link: true
    title-block-banner: true
  pdf:
    number-sections: true
    number-depth: 3
editor_options: 
  chunk_output_type: console
---

# Presentation

This is the data preparation code for the paper "Perceptions of Inequality and Meritocracy: Their Interplay in Shaping Preferences for Market Justice in Chile (2016-2023)". The prepared dataset is `ELSOC_Long_2016_2023_1.00.RData`.

```{r}
#| label: set
#| echo: false
#| message: false
#| warning: false

library(knitr)
knitr::opts_chunk$set(echo = TRUE, include = TRUE, warning = FALSE, message = FALSE)

table_format <- if(is_html_output()) {
  "html"
} else if(is_latex_output()) {
  "latex"
}
table_format2 <- if(is_html_output()) {
  T
} else if(is_latex_output()) {
  F
}

options(kableExtra.html.bsTable = T)
options(knitr.kable.NA = "")
```

# Libraries

```{r}
#| label: packages
#| include: true

if (! require("pacman")) install.packages("pacman")

pacman::p_load(tidyverse,
               sjmisc, 
               here,
               sjlabelled,
               SciViews,
               naniar)

options(scipen=999)
rm(list = ls())
```

# Data

```{r}
#| label: data
#| echo: true
#| output: false

load(url("https://dataverse.harvard.edu/api/access/datafile/10797987"))

glimpse(elsoc_long_2016_2023)

```

# Processing

## Select

```{r}
db_proc <- elsoc_long_2016_2023 %>% 
  select(idencuesta, ola, 
         ponderador_long_total,segmento, estrato,
         just_educ = d02_02, just_pension = d02_01, 
         just_health = d02_03, merit_effort = c18_09,
         merit_talent = c18_10, perc_sal_gerente = d03_01,
         perc_sal_obrero = d03_02, just_sal_gerente = d04_01, 
         just_sal_obrero = d04_02, age = m0_edad, m01, 
         sex = m0_sexo, ess = d01_01, ideo = c15
         ) %>% 
  as_tibble()
```

## Filter

```{r}
#db_proc %>% 
#  group_by(ola) %>% 
#  count(just_educ,just_pension,just_health) %>% 
#  na.omit() %>% 
#  print(n = nrow(.))

db_proc <- db_proc %>% dplyr::filter(ola %in% c(1,2,3,4,6,7))

```

## Recode and transform

```{r}
# general na's
db_proc[db_proc ==-999] <- NA
db_proc[db_proc ==-888] <- NA
db_proc[db_proc ==-777] <- NA
db_proc[db_proc ==-666] <- NA

# mjp

frq(db_proc$just_health)
frq(db_proc$just_pension)
frq(db_proc$just_educ)

labels1 <- c("Strongly desagree" = 1, 
             "Desagree" = 2, 
             "Neither agree nor desagree" = 3, 
             "Agree" = 4, 
             "Strongly agree" = 5)

db_proc <- db_proc %>% 
  mutate(
    across(
      .cols = c(just_health, just_pension, just_educ),
      .fns = ~ sjlabelled::set_labels(., labels = labels1)
      )
    )

db_proc <- cbind(db_proc, "mjp" = rowMeans(db_proc %>% select(just_health, just_pension, just_educ), na.rm=TRUE))

summary(db_proc$mjp)

db_proc <- db_proc %>% 
  mutate(
    across(
      .cols = c(just_health, just_pension, just_educ),
      .fns = ~ car::recode(., recodes = c("1='Strongly desagree'; 2='Desagree';
                                          3='Neither agree nor desagree'; 4='Agree';
                                          5='Strongly agree'"), 
                           levels = c("Strongly desagree", "Desagree", "Neither agree nor desagree", "Agree", "Strongly agree"),
                           as.factor = T)
    )
  )

db_proc$just_health <- sjlabelled::set_label(db_proc$just_health, 
                        label = "Health distributive justice")

db_proc$just_pension <- sjlabelled::set_label(db_proc$just_pension, 
                        label = "Pension distributive justice")

db_proc$just_educ <- sjlabelled::set_label(db_proc$just_educ, 
                        label = "Education distributive justice")

db_proc$mjp <- sjlabelled::set_label(db_proc$mjp, 
                        label = "Market justice preferences")

# merit

frq(db_proc$merit_effort)
frq(db_proc$merit_talent)

db_proc <- db_proc %>% 
  mutate(
    across(
      .cols = c(merit_effort, merit_talent),
      .fns = ~ sjlabelled::set_labels(., labels = labels1)
      )
    )

db_proc <- db_proc %>% 
  mutate(
    across(
      .cols = c(merit_effort, merit_talent),
      .fns = ~ car::recode(., recodes = c("1='Strongly desagree'; 2='Desagree';
                                          3='Neither agree nor desagree'; 4='Agree';
                                          5='Strongly agree'"), 
                           levels = c("Strongly desagree", "Desagree", "Neither agree nor desagree", "Agree", "Strongly agree"),
                           as.factor = T)
    )
  )

db_proc$merit_effort <- sjlabelled::set_label(db_proc$merit_effort, 
                        label = "People are rewarded for their efforts")

db_proc$merit_talent <- sjlabelled::set_label(db_proc$merit_talent, 
                        label = "People are rewarded for their intelligence")

# perc inequality gap

frq(db_proc$perc_sal_gerente)
frq(db_proc$perc_sal_obrero)

db_proc <- db_proc %>% 
  mutate(perc_sal_obrero=replace(perc_sal_obrero, perc_sal_obrero <= 40000  | perc_sal_obrero>= 1000000, NA)) %>%
  mutate(perc_sal_gerente=replace(perc_sal_gerente, perc_sal_gerente <= 250000 | perc_sal_gerente>= 100000001, NA))

db_proc$perc_inequality <- SciViews::ln(db_proc$perc_sal_gerente/db_proc$perc_sal_obrero)

summary(db_proc$perc_inequality)

db_proc$perc_inequality <- sjlabelled::set_label(db_proc$perc_inequality, 
                        label = "Inequality gap perception")

# just inequality gap

frq(db_proc$just_sal_gerente)
frq(db_proc$just_sal_obrero)

db_proc <- db_proc %>% 
  mutate(just_sal_obrero=replace(just_sal_obrero, just_sal_obrero <= 40000  | just_sal_obrero>=999999999999999, NA)) %>%
  mutate(just_sal_gerente=replace(just_sal_gerente, just_sal_gerente <= 100000 | just_sal_gerente>=999999999999999, NA))

db_proc$just_inequality <- SciViews::ln(db_proc$just_sal_gerente/db_proc$just_sal_obrero)

summary(db_proc$just_inequality)

db_proc$just_inequality <- sjlabelled::set_label(db_proc$just_inequality, 
                        label = "Inequality gap justification")

# sex
db_proc$sex <- car::recode(db_proc$sex, 
                           recodes = c("1='Male'; 2='Female'"), 
                           levels = c("Male", "Female"),
                           as.factor = T)

db_proc$sex <- sjlabelled::set_label(db_proc$sex, 
                        label = "Gender")

# wave 
db_proc <- db_proc %>% 
  rename(wave = ola) %>% 
  mutate(wave = case_when(wave == 1 ~ 1,
                          wave == 2 ~ 2,
                          wave == 3 ~ 3,
                          wave == 4 ~ 4,
                          wave == 6 ~ 5,
                          wave == 7 ~ 6,
                          TRUE ~ NA_real_))

db_proc$wave <- sjlabelled::set_label(db_proc$wave, 
                        label = "Wave")

# age
frq(db_proc$age)

db_proc$age <- 
  factor(car::recode(db_proc$age, 
                     "18:29=1;30:49=2;50:64=3;65:150=4"),
         labels = c('18-29', '30-49', '50-64', '65 or more'))
db_proc$age <-
  sjlabelled::set_label(db_proc$age, 
                        label = c("Age groups"))

# political indentification

frq(db_proc$ideo)

db_proc$ideo<-
factor(
  car::recode(
    db_proc$ideo,
    "c(11,12,-888,-999)='Does not identify';c(0,1,2,3,4)='Left';
     c(5)='Center';c(6,7,8,9,10)='Right'"
  ),
  levels = c('Left', 'Center', 'Right', 'Does not identify')
)

db_proc$ideo<- factor(db_proc$ideo,levels = levels(db_proc$ideo))

db_proc$ideo <- 
sjlabelled::set_label(x = db_proc$ideo, 
                      label = "Political identification")

frq(db_proc$ideo)

# ess
sjmisc::frq(db_proc$ess)

db_proc$ess <- as.numeric(db_proc$ess)

db_proc$ess <-  sjlabelled::set_label(x = db_proc$ess,
                      label = "Subjective Social Status")

# educ
frq(db_proc$m01)

db_proc$educ <- 
  car::recode(db_proc$m01,
              "c(1,2,3,4,5,6,7)=1;c(8,9,10)=2; c(-888,-999)=NA")
db_proc$educ <-
  factor(db_proc$educ,
         labels = c("Less than Universitary","Universitary"))

db_proc$educ <- 
sjlabelled::set_label(x = db_proc$educ,
                      label = "Education")

#Recoding of education to years based on casen 2017.
db_proc$educyear<- as.numeric(
  car::recode(db_proc$m01, 
              "1=0;2=4.3;3=7.5;4=9.8;5=12.02;6=13.9;
               7=14.8;8=14.9;9=16.9;10=19.07;c(-888,-999)=NA", 
              as.numeric = T))

db_proc$educyear <- 
sjlabelled::set_label(x = db_proc$educyear,
                      label = "Education in years")

class(db_proc$educyear)
frq(db_proc$educyear)
```

```{r}
# Household income----------------------------------------------------------------

# Ltop is the lower limit of the top category, Ltop-1 is the lower limit of the category before the
# top one, ftop is the frequency in the top category, and ftop-1 is the frequency in the
# category before the top one.

sjmisc::frq( elsoc_long_2016_2023$m30)
Ltop_1<- 1850001
Ltop  <- 2700000

ftop_1 <- 42
ftop   <- 34

V  = (log(ftop_1 + ftop) - log(ftop)) / (log(Ltop) -log(Ltop_1))
M_top = 0.5* Ltop *(1+(V/(V-1)));M_top # = $3.897.232

#Impute midpoint of income ranges
elsoc_long_2016_2023$m30_rec <-
  as.numeric(car::recode( elsoc_long_2016_2023$m30,
                         "1=110000;2=251000;3=305000;4=355000;5=400000;
            6=445000;7=490000;8=535000;9=585000;10=640000;11=700000;12=765000;
            13=845000;14=935000;15=1040000;16=1180000;17=1375000;18=1670000;
            19=2275000;20=3897232;NA=NA;c(-888,-999)=NA"))

summary( elsoc_long_2016_2023$m30_rec)

#Impute midpoint of income ranges (2021)

frq(elsoc_long_2016_2023$m30b)
Ltop_1<- 450000
Ltop  <- 700000
ftop_1 <- 18
ftop   <- 27

V  = (log(ftop_1 + ftop) - log(ftop)) / (log(Ltop) -log(Ltop_1))
M_top = 0.5* Ltop *(1+(V/(V-1))) # = $2.941.412

 elsoc_long_2016_2023$m30b_rec <-
  as.numeric(car::recode( elsoc_long_2016_2023$m30b,
                         "1=125000;2=300000;3=400000;4=575000;5=2941412;NA=NA;c(-888,-999)=NA"))
summary( elsoc_long_2016_2023$m30b_rec)

# Combine m30_rec with m30b_rec
 elsoc_long_2016_2023$m30_rec <- 
  ifelse(is.na( elsoc_long_2016_2023$m30_rec),
         yes =  elsoc_long_2016_2023$m30b_rec,
         no =  elsoc_long_2016_2023$m30_rec)
summary( elsoc_long_2016_2023$m30_rec)

#Recode DK/DA of Income to NA
elsoc_long_2016_2023$m29_rec <-
  as.numeric(car::recode( elsoc_long_2016_2023$m29,"c(-888,-999)=NA"))

#replace NA of income with new imputed variable
elsoc_long_2016_2023$m29_imp <- 
  ifelse(test = !is.na( elsoc_long_2016_2023$m29_rec),
         yes =   elsoc_long_2016_2023$m29_rec,
         no =   elsoc_long_2016_2023$m30_rec)
summary( elsoc_long_2016_2023$m29_imp)

#  elsoc_long_2016_2023$m29_imp <- 
#   ifelse(test = is.na( elsoc_long_2016_2023$m29_imp),
#          yes =   elsoc_long_2016_2023$m30b_rec,
#          no =   elsoc_long_2016_2023$m29_imp)

summary( elsoc_long_2016_2023$m29_imp)

# N Household:
elsoc_long_2016_2023 <-
   elsoc_long_2016_2023 %>%
  mutate(n_hogar =
           dplyr::case_when(ola == 1 ~ nhogar1,
                            ola == 2 ~ m46_nhogar,
                            ola == 3 ~ m54,
                            ola == 4 ~ m54,
                            ola == 5 ~ m54,
                            ola == 6 ~ m54,
                            ola == 7 ~ m54))

summary( elsoc_long_2016_2023$n_hogar)
#imputar tamanio del hogar de ola 5 a ola 6
elsoc_long_2016_2023 <- 
   elsoc_long_2016_2023 %>%  
  group_by(idencuesta) %>%
  mutate(n_hogar = if_else(ola == 6 & is.na(n_hogar), last(n_hogar[ola == 5]), n_hogar)) %>% 
  ungroup()

sjmisc::frq( elsoc_long_2016_2023$n_hogar)
table( elsoc_long_2016_2023$n_hogar, elsoc_long_2016_2023$ola)

#Recode DK/DA to NA
 elsoc_long_2016_2023$n_hogar_r<-
  car::recode( elsoc_long_2016_2023$n_hogar,"c(-888,-999)=NA")

table( elsoc_long_2016_2023$m13, elsoc_long_2016_2023$n_hogar_r)

#ingresos individuales para hogares de 1 persona
 elsoc_long_2016_2023 %>% sjmisc::find_var('Ingreso')
 elsoc_long_2016_2023$inc_ind <- ifelse( elsoc_long_2016_2023$n_hogar_r==1, elsoc_long_2016_2023$m13,NA)
summary( elsoc_long_2016_2023$inc_ind)
# 
# c(40000,
# (40001 + 85000) / 2,
# (85001 + 125000) / 2,
# (125001 + 170000) / 2,
# (170001 + 210000) / 2,
# (210001 + 230000) / 2,
# (230001 + 280000) / 2,
# (280001 + 320000) / 2,
# (320001 + 360000) / 2,
# (360001 + 400000) / 2,
# (400001 + 465000) / 2,
# (465001 + 540000) / 2,
# (540001 + 665000) / 2,
# (665001 + 850000) / 2,
# (850001 + 1300000) / 2,
# 1300001)

frq( elsoc_long_2016_2023$m14)
Ltop_1<- 850001
Ltop  <- 1300001 
ftop_1 <- 65
ftop   <- 40

V  = (log(ftop_1 + ftop) - log(ftop)) / (log(Ltop) -log(Ltop_1))
M_top = 0.5* Ltop *(1+(V/(V-1))) # = $1.811.247

 elsoc_long_2016_2023$m14_rec<- 
  car::recode( elsoc_long_2016_2023$m14,
              "1=40000.0;2=62500.5;3=105000.5;4=147500.5;5=190000.5;6=220000.5;
              7=255000.5;8=300000.5;9=340000.5;10=380000.5;11=432500.5;12=502500.5;
              13=602500.5;14=757500.5;15=1075000.5;16=1811247"
  )

 elsoc_long_2016_2023$inc_ind <- ifelse( elsoc_long_2016_2023$n_hogar_r==1 & is.na( elsoc_long_2016_2023$inc_ind), elsoc_long_2016_2023$m14_rec, elsoc_long_2016_2023$inc_ind)
summary( elsoc_long_2016_2023$inc_ind)

# Ingreso del hogar pegarle ingreso del hogar unipersonal
summary( elsoc_long_2016_2023$m29_imp)
 elsoc_long_2016_2023$m29_imp <- 
  ifelse(is.na( elsoc_long_2016_2023$m29_imp) &  elsoc_long_2016_2023$n_hogar_r==1,
          elsoc_long_2016_2023$inc_ind,
          elsoc_long_2016_2023$m29_imp) #recupero 14 casos
summary( elsoc_long_2016_2023$m29_imp)

# Household Equivalent income 
# 1.0 to the first adult;
# 0.5 to the second and each subsequent person aged 14 and over;
# 0.3 to each child aged under 14.

 elsoc_long_2016_2023$n_hogar_r_eq <- ( elsoc_long_2016_2023$n_hogar_r + 1) /2
table( elsoc_long_2016_2023$n_hogar_r_eq)

frq( elsoc_long_2016_2023$n_hogar_r_eq)

# Per capita household income:
 elsoc_long_2016_2023$ing_pc <- 
  (elsoc_long_2016_2023$m29_imp/ elsoc_long_2016_2023$n_hogar_r)

# Per capita household income equivalized (OECD):
 elsoc_long_2016_2023$ing_pc_eq <- 
  ( elsoc_long_2016_2023$m29_imp/ elsoc_long_2016_2023$n_hogar_r_eq)

descr( elsoc_long_2016_2023$ing_pc_eq)
descr( elsoc_long_2016_2023$ing_pc)
 elsoc_long_2016_2023$ing_pc <-
  sjlabelled::set_label(x =  elsoc_long_2016_2023$ing_pc,
                        label = "Household income per capita")

sjmisc::descr( elsoc_long_2016_2023$ing_pc)

# Compute income quintiles
 elsoc_long_2016_2023 <-  elsoc_long_2016_2023 %>% 
  group_by(ola) %>% 
  mutate(quintil = ntile(-desc(ing_pc), 5)) %>% 
  ungroup()

 elsoc_long_2016_2023$quintil <- 
  factor( elsoc_long_2016_2023$quintil,
         levels = c(1, 2, 3, 4, 5),
         labels = c('Q1', 'Q2', 'Q3', 'Q4', 'Q5')) # Quintiles as factors

elsoc_long_2016_2023$quintil <- 
  sjlabelled::set_label(x =  elsoc_long_2016_2023$quintil,
                        label = "Household income quintile per capita")

sjmisc::frq( elsoc_long_2016_2023$quintil)

#include new quintile category with missing cases
 elsoc_long_2016_2023$quintil1<-
  car::recode( elsoc_long_2016_2023$quintil, 
              "'Q1'='Q1';'Q2'= 'Q2';'Q3'='Q3';'Q4'='Q4';'Q5'='Q5'; NA='QNA'")

# elsoc_long_2016_2023$quintil1 <- factor( elsoc_long_2016_2023$quintil1, c("Q1","Q2","Q3","Q4","Q5","QNA"))

 elsoc_long_2016_2023$quintil1 <- 
  sjlabelled::set_label(x =  elsoc_long_2016_2023$quintil1,
                        label = "Household income quintile per capita (NA)") 
sjmisc::frq( elsoc_long_2016_2023$quintil1)

```

```{r}
elsoc_income <- elsoc_long_2016_2023 %>% 
  dplyr::select(idencuesta, ola, ing_pc, quintil, quintil1) %>% 
  dplyr::filter(ola %in% c(1,2,3,4,6,7))

db_proc <- left_join(db_proc, elsoc_income, by = c("idencuesta", "wave" = "ola"))

```

## Missing values

```{r}
colSums(is.na(db_proc))

n_miss(db_proc)

prop_miss(db_proc)*100

miss_var_summary(db_proc)

miss_var_table(db_proc)

vis_miss(db_proc) + theme(axis.text.x = element_text(angle=80))

```

# Save and export

```{r}
db_proc <- db_proc %>% 
  select(-c(perc_sal_gerente, perc_sal_obrero, just_sal_gerente, just_sal_obrero, m01))

df1 <- db_proc %>% na.omit()

df1 <- sjlabelled::copy_labels(df_new = df1, df_origin = db_proc)

save(df1, file = here("input/data/proc/df1.RData"))
```