### 2025-04-11 03-methods.qmd
Update analysis and repo order

---
author: "Equipo EDUMER"
bibliography: "input/bib/refs.bib"
csl: "input/bib/apa6.csl"
---

# Data, Variables and Methods

## Data

## Variables

## Methods