### 2025-04-23 02-antecedents.qmd

---
author: "Equipo EDUMER"
bibliography: "input/bib/merit-edjust.bib"
csl: "input/bib/apa6.csl"
---

# Theoretical views on market justice, inequality perception and meritocracy

## Justification of market inequality

Market justice has been discussed in the literature as a normative principle that legitimates the distribution of economic rewards based on individual merit and effort. The most detailed theoretical account of this concept is provided by Lane [-@lane_market_1986], who defines market justice as a system of "earned deserts," whereby individuals are seen as entitled to their outcomes insofar as these are a product of personal productivity and competition. This view directly contrasts with political justice, which emphasizes redistributive mechanisms based on the principles of equality and need, and is often materialized through welfare policies. For Lane [-@lane_market_1986], market justice relies on the conception that markets are neutral and self-regulating arenas, where individuals are treated fairly because they face the same formal rules of engagement. What matters is not whether everyone ends up with the same, but that everyone has the chance to compete, and that success reflects effort and ability. Consequently, the legitimacy of market justice stems from the assumption that inequalities are not only inevitable but fair---so long as the rules are transparent and opportunities are open. These beliefs align with a broader ideological framework that values individual autonomy, private property, and limited government intervention. In this way, market justice provides a moral justification for inequality by framing it as a necessary outcome of personal responsibility and societal efficiency.

Empirical studies have developed different strategies to capture and operationalize market justice preferences. A common approach in the literature is to gauge attitudes toward the legitimacy of inequality in specific domains, especially when linked to income differences. Kluegel and Smith [-@kluegel_beliefs_1981] and later Kluegel et al. [-@kluegel_legitimation_1999] laid the groundwork for measuring beliefs in market justice through survey items that assess whether people support economic inequality when it reflects personal effort and skill. Over time, this approach has been extended beyond income to include other market-mediated outcomes, such as education, healthcare, or pensions. For example, Knesebeck et al. (2016) and Immergut and Schneider [-@immergut_it_2020] examine whether citizens consider it fair that individuals with higher incomes can access better healthcare, while Lee and Stacey [-@lee_fairness_2023] apply a similar method in the context of education in Australia. These studies usually rely on a survey item asking respondents to evaluate the fairness of income-based access to welfare services, allowing for comparing justice perceptions across different contexts. Some research, like that of Lindh [-@lindh_public_2015] and Svallfors [-@svallfors_political_2007], uses these indicators in combination to measure broader preferences for market-based distribution in welfare systems. More recently, Castillo et al. [-@castillo_socialization_2024] introduced a single-item composite measure of market justice to assess student attitudes toward income-based access to education, healthcare, and pensions in Chile. These different operationalizations all aim to capture the extent to which individuals accept inequality when it is framed as a reflection of market outcomes.

The study of market justice preferences has increasingly focused on how they are shaped by individuals' socioeconomic position, normative beliefs, and the institutional context in which they are embedded. Across the literature, there is strong empirical evidence suggesting that individuals in more advantaged socioeconomic positions---those with higher income, education, and occupational status---are more likely to endorse market-based distribution as fair and legitimate [@svallfors_political_2007; @koos_moral_2019]. This tendency reflects not only material self-interest but also a broader moral economy, in which winners of the market system internalize justifications for the status quo. At the same time, normative and ideological orientations---such as conservative, liberal, or neoliberal values---are consistently associated with greater support for meritocratic principles and skepticism toward redistribution. This is particularly visible in contexts such as Chile and Australia, where studies by Castillo et al. [-@castillo_socialization_2024] and Lee and Stacey [-@lee_fairness_2023] show that those with right-leaning political preferences tend to view income-based access to welfare services as legitimate. Beyond individual-level characteristics, country-level institutions also play a central role. In liberal welfare regimes like those of the United States or the United Kingdom, market justice preferences are more widespread, while in coordinated or social-democratic regimes---such as Sweden or Germany---citizens are generally more critical of market-based inequalities [@immergut_it_2020; @lindh_public_2015]. These findings suggest that market justice is not a fixed belief but a contextually shaped orientation, mediated by structural conditions, cultural values, and the historical development of welfare institutions.

## Perception of inequality

Perceptions of inequality refer to individuals' subjective evaluations of the extent, causes, and consequences of income and wealth disparities. Unlike objective measures such as the Gini index, perceived inequality captures how individuals make sense of distributive hierarchies in their everyday lives, shaped by reference groups, social comparisons, and information environments [@mijs_stratified_2016; @gimpelson_misperceiving_2018; @garcia-castro_perceiving_2020]. Scholars have proposed multiple dimensions of perceived inequality, including its magnitude (how significant are the gaps), vertical structure (between which groups), trend over time (increasing or decreasing), and legitimacy (whether it is just or not) [@engelhardt_what_2018; @garcia-sanchez_vicious_2019]. These dimensions encompass both cognitive and normative aspects of perceptions of inequality and can vary across societies and social groups, depending on exposure, ideology, and personal experience [@castillo_perception_2022; @garcia-sanchez_perceptions_2018].

Among the different approaches to measuring perceived inequality, one of the most widely used is the estimation of wage gaps between occupational extremes, such as between a CEO and a manual worker. This type of item provides a concrete frame that respondents can relate to more easily than abstract questions about national income distribution [@castillo_percepcion_2019; @easterbrook_social_2021; Willis et al., 2015]. While it enables the estimation of inequality using simple heuristics, this method is not without challenges. For instance, people often lack reliable knowledge about the earnings of those at the top of the income ladder, which leads to high variability in responses and the use of biased mental shortcuts [@knell_perceptions_2020]. Despite this, perceived wage gaps are strong predictors of political attitudes[@pedersen_attitudes_2019; @garcia-sanchez_perceptions_2018], making them a valuable tool for understanding public responses to economic disparities.

Another relevant critique of this approach lies in its conflation of different psychological constructs. Many surveys assess perceived inequality through Likert-type items that ask respondents to agree or disagree with statements, such as "income differences are too large," which captures general concern or discomfort rather than a specific perception [@Castillo2011; @garcia-sanchez_vicious_2019].These items mix cognitive estimations with affective evaluations, complicating the interpretation of what respondents perceive versus what they morally reject. As a result, the conceptual clarity between perceived inequality and inequality aversion remains blurred in many empirical studies. To address this limitation, recent work has emphasized the need to distinguish between absolute and comparative measures, as well as between ideal and actual estimates of economic gaps [@garcia-sanchez_creencias_2022; Auspurg et al., 2017].

Importantly, perceptions of inequality have been linked to attitudes about market justice---the belief that goods and services should be allocated according to individual contribution, competition, or willingness to pay [@lindh_public_2015; @kluegel_social_1995a]. Research indicates that lower perceived inequality can reinforce support for market-based distributive arrangements by suggesting that the system is fair and that outcomes reflect effort and ability [@kuhn_eye_2011]. In contrast, when inequality is perceived as excessive or structurally determined, individuals could question the legitimacy of market justice and become more supportive of redistributive policies [@castillo_perception_2022; @garcia-sanchez_vicious_2019].

## Perception of meritocracy

Meritocracy constitutes a central ideological framework for legitimizing social inequality. Rooted in the belief that rewards and positions should be allocated based on individual effort and talent, meritocracy operates as a normative ideal and a descriptive belief about how society functions. As initially conceptualized by Michael Young [-@young_rise_1962], the term was meant to critique a system in which merit-based stratification becomes a new form of inequality. However, over time, meritocracy has been widely adopted in many societies as a fair and desirable principle of distribution, particularly within liberal democracies and market-oriented economies [@mijs_paradox_2021; @sandel_tyranny_2020]. From a sociological perspective, the belief in meritocracy is more than a cognitive assessment; it reflects a moral lens through which individuals interpret inequality. People who believe that success results from hard work and talent are more likely to view social and economic disparities as legitimate [@castillo_percepcion_2019; @batruch_belief_2023]. Conversely, if they see outcomes as driven by luck, social origin, or systemic barriers, inequality is more likely to be perceived as unjust. This distinction becomes crucial in societies with persistent structural inequality, where public narratives often emphasize personal responsibility and merit while overlooking entrenched disadvantages.

We adopt a multidimensional perspective on meritocracy, distinguishing between two key dimensions: effort-based and talent-based perceptions. This distinction is essential, as it captures different pathways through which individuals justify inequality. Effort-based meritocracy emphasizes hard work and perseverance as the basis for social rewards, aligning closely with cultural narratives of personal responsibility. A talent-based meritocracy, by contrast, emphasizes intelligence and innate abilities, which are often perceived as less malleable and more unequally distributed. Both dimensions have been shown to correlate with acceptance of inequality, but they may carry distinct implications for how inequality is justified in specific domains [@castillo_multidimensional_2023]. The relevance of this distinction is supported by recent studies, which show that individuals respond differently to these dimensions. For instance, perceptions that effort is rewarded in society are more strongly associated with positive evaluations of fairness and acceptance of unequal outcomes [@batruch_belief_2023]. This may be because effort is seen as a controllable and morally virtuous trait, whereas talent is often perceived as a natural advantage. Consequently, effort-based meritocracy is likely more potent in legitimizing inequality, particularly in neoliberal contexts.

These dimensions of meritocracy reflect how respondents perceive society's distributive logic, regardless of whether they personally endorse meritocratic principles. This distinction aligns with recent findings indicating that individuals distinguish between how merit is perceived in society and how it should ideally operate, which in turn shapes their preferences for redistribution and justice (Tejero-Peregrina et al., 2025). Meritocratic beliefs serve as symbolic justifications for unequal outcomes, particularly when access is stratified by income or social opportunity. Prior studies have shown that individuals who perceive higher levels of meritocracy tend to express stronger support for unequal distributions that reflect market outcomes [@castillo_percepcion_2019; @castillo_socialization_2024].

In addition to influencing individual attitudes toward inequality, meritocratic beliefs can contribute to social division and the stigmatization of disadvantaged groups. Recent research has demonstrated that exposure to meritocratic narratives can reinforce the belief that poverty results from individual failings rather than systemic conditions, reducing support for redistributive measures and increasing the stigmatization of the poor (Hoyt & Burnette, 2023). This reinforces negative stereotypes and reduces empathy toward individuals from lower socioeconomic backgrounds. Moreover, Busemeyer, Abrassart, and Nezi (2019) argue that meritocratic narratives can serve as feedback mechanisms that shape public opinion and well-being by framing individuals' understanding of welfare outcomes as deserved or undeserved within existing institutional structures. This psychological mechanism highlights the normative power of meritocracy in stabilizing unequal systems by shaping political attitudes and personal perceptions of success and failure.

## This study (hypotheses)