### 2025-04-28 paper.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
]{article}

\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
  \setmainfont[]{Times New Roman}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=2cm]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage[noblocks]{authblk}
\renewcommand*{\Authsep}{, }
\renewcommand*{\Authand}{, }
\renewcommand*{\Authands}{, }
\renewcommand\Affilfont{\small}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Perceptions of Inequality and Meritocracy: Their Interplay in Shaping Preferences for Market Justice in Chile (2016-2023)},
  pdfauthor={Equipo EDUMER},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Perceptions of Inequality and Meritocracy: Their Interplay in
Shaping Preferences for Market Justice in Chile (2016-2023)}

  \author{Juan Carlos Castillo}
            \affil{%
                  Departamento de Sociología, Universidad de Chile
              }
          \affil{%
                  Centro de estudios del conflicto y cohesión social
                  (COES)
              }
          \affil{%
                  Núcleo milenio de desigualdades y oportunidades
                  digitales (NUDOS)
              }
        \author{Andreas Laffert}
            \affil{%
                  Instituto de Sociología, Pontificia Universidad
                  Católica de Chile
              }
        \author{Kevin Carrasco}
            \affil{%
                  Centro de estudios del conflicto y cohesión social
                  (COES)
              }
        \author{Julio Iturra}
            \affil{%
                  International Graduate School of Social Sciencies
                  (BIGSSS), University of Bremen, Germany
              }

\date{}
\begin{document}
\maketitle
\begin{abstract}
This study examines the interplay between perceptions of economic
inequality and meritocracy in shaping preferences for market justice in
Chile from 2016 to 2023. Economic inequality has intensified globally,
raising concerns about its social impacts and the legitimacy of
market-based distributions of resources. By utilizing longitudinal data
from the Chilean Longitudinal Social Survey (ELSOC), the research
investigates how subjective assessments of inequality influence
attitudes towards social justice, focusing on the extent to which
individuals believe access to essential services like healthcare,
education, and pensions should depend on personal merit. The findings
suggest that as perceptions of economic inequality rise, individuals
tend to favor redistributive policies, although those with strong
meritocratic beliefs are more likely to justify existing inequalities.
The analysis also highlights the effect of major social movements in
Chile during this period, which have shifted public discourse and
potentially altered perceptions of inequality and meritocracy. This
paper contributes to the understanding of how individual beliefs about
fairness and equity evolve within a socio-political context
characterized by high inequality and extensive free-market principles.
\newline \textbf{Keywords}: Economic inequality, meritocracy, market
justice, Chile, public preferences
\end{abstract}

\setstretch{1.15}
\section{Introduction}\label{introduction}

Since 1980, economic inequality and wealth concentration have
dramatically increased worldwide, becoming one of the main challenges
for the social sciences. Globally, in 2021, less than 50\% of the world
population owned only 2\% of the wealth, while the richest 10\%
concentrated 76\%, and the wealthiest 1\% captured nearly 38\% of total
assets (\citeproc{ref-chancel_world_2022}{Chancel et al., 2022}). This
context of economic disparity has sparked renewed interest in studying
not only the objective aspects of inequality, such as income and access
to resources, but also its subjective dimensions, including perceptions,
beliefs, and associated attitudes
(\citeproc{ref-janmaat_subjective_2013}{Janmaat, 2013}). This has led,
among other reasons, to widespread concern in the social sciences about
the political delegitimization that economic inequality can produce
(\citeproc{ref-castillo_perception_2022}{Castillo et al., 2022}).
Understanding how people perceive inequality is crucial, as these
perceptions can influence how societies comprehend and justify the
distribution of goods and services.

The perception of economic inequality can be understood as an
individual's subjective assessment of how resources are allocated among
members of a given society (\citeproc{ref-akyelken_urban_2020}{Akyelken,
2020}). Regardless of their measurement, various studies have shown that
this perception often underestimates the gap between the rich and the
poor, which could have implications for attitudes toward the
distribution of goods and services
(\citeproc{ref-castillo_perception_2022}{Castillo et al., 2022};
\citeproc{ref-schroder_income_2017}{Schröder, 2017}). Moreover, in
recent years, there has been a discussion about the subjective and
objective aspects of inequality, demonstrating inconsistencies between
the two (\citeproc{ref-trump_income_2018}{Trump, 2018}). This tension is
relevant because advancing the study of perceptions could help
understand how perceived inequality affects the distribution of
resources within a society (Becker, 2020).

Perception of economic inequality has been associated with important
outcomes such as redistributive preferences, justification of
inequality, and legitimacy of the economic system
(\citeproc{ref-castillo_perception_2022}{Castillo et al., 2022};
\citeproc{ref-garcia-sanchez_attitudes_2020}{García-Sánchez et al.,
2020}, \citeproc{ref-garcia-sanchez_perceptions_2018}{2018}). Research
shows that greater perceived inequality fosters stronger preferences for
redistribution, regardless of actual inequality levels
(\citeproc{ref-garcia-sanchez_vicious_2019}{García-Sánchez et al.,
2019}; \citeproc{ref-gimpelson_misperceiving_2018}{Gimpelson \&
Treisman, 2018}). However, other findings suggest that heightened
perceptions of income gaps between high- and low-paying occupations may
lead to greater justification of inequality
(\citeproc{ref-Castillo2011}{Castillo, 2011}). A less explored area in
this regard concerns market justice, which denotes the degree to which
individuals consider just that social goods and services (e.g., health,
pensions, education) are allocated according to individual contribution,
competition, and ability to pay
(\citeproc{ref-kluegel_legitimation_1999}{Kluegel et al., 1999};
\citeproc{ref-lane_market_1986}{Lane, 1986};
\citeproc{ref-lindh_public_2015}{Lindh, 2015}). Market justice attitudes
reflect the belief that the market promotes procedural
fairness---equality of opportunity---so that outcomes result from
individual achievement (\citeproc{ref-lane_market_1986}{Lane, 1986}).
Indeed, Lindh (\citeproc{ref-lindh_public_2015}{2015}) shows that market
justice preferences are stronger in countries with higher private
spending on services than in those with more comprehensive welfare
systems. From this perspective, perceptions of economic inequality may
be key, as lower perceived gaps can reinforce acceptance of free-market
systems and justify existing inequalities.

Meritocracy posits that inequality can be legitimized through
distributive criteria such as effort and talent
(\citeproc{ref-davis_principles_2001}{Davis \& Moore, 2001};
\citeproc{ref-young_rise_1962}{Young, 1962}). Previous studies have
demonstrated that people with stronger meritocratic beliefs tend to
perceive less inequality as they attribute economic differences to
individual achievements (\citeproc{ref-mijs_paradox_2021}{Mijs, 2021};
\citeproc{ref-wilson_role_2003}{Wilson, 2003}) and also to justify more
inequality as these beliefs are associated with attitudes that
legitimize status differences
(\citeproc{ref-batruch_belief_2023}{Batruch et al., 2023}). In a context
like Chile, where the distribution of goods and services is
predominantly governed by market logics strongly introduced during the
military dictatorship (1973-1989)
(\citeproc{ref-boccardo_30_2020}{Boccardo, 2020}), these beliefs can
play a crucial role in the acceptance of social inequalities.
Incorporating the perception of meritocracy in this literature allows
for an understanding of how attitudes toward inequality are shaped not
only by the perception of economic gaps, but also by beliefs about how
these gaps are justified. Furthermore, these beliefs are often
consolidated from an early age, reinforced by institutions promoting
values such as effort and individual skills to climb socially
(\citeproc{ref-castillo_socialization_2024}{Castillo et al., 2024};
\citeproc{ref-reynolds_perceptions_2014}{Reynolds \& Xian, 2014}).

The perception of meritocracy and the perception of economic inequality
would interact intricately in shaping preferences for market justice. On
the one hand, a stronger perception of merit as a criterion for
distribution could minimize the perception of economic gaps, thus
justifying an unequal system
(\citeproc{ref-castillo_percepcion_2019}{Castillo et al., 2012};
\citeproc{ref-mijs_paradox_2021}{Mijs, 2021}). On the other hand, the
perception of inequality could moderate the impact of meritocracy, as a
higher perception of economic gaps might challenge the idea that these
are solely based on merit.

The primary objective of this study is to analyze the interplay between
perceptions of inequality and meritocracy and their joint influence on
preferences for market justice in Chile from 2016 to 2023, using
longitudinal data from the ELSOC survey. This interaction is expected to
provide a more comprehensive explanation of attitudes toward market
justice in a country characterized by high inequality and strong
free-market influence (\citeproc{ref-boccardo_30_2020}{Boccardo, 2020};
\citeproc{ref-flores_top_2020}{Flores et al., 2020}). Additionally, this
analysis seeks to elucidate how political and social
contingencies---such as the 2019 and 2022 social movements---might have
moderated these relationships by prompting more critical reflection on
the commodification of social services. Recognizing the temporal
dimension in shaping market justice preferences is essential, given that
such preferences are not static but are influenced by historical and
contextual factors that challenge or reaffirm prevailing social norms.
In this regard, variations in perceived economic inequality and
meritocracy over time can affect how individuals endorse market-based
approaches. Consequently, public opinion may shift from supporting
market justice to embracing redistributive policies aimed at mitigating
inequalities.

\section{Theoretical views on market justice, inequality perception and
meritocracy}\label{theoretical-views-on-market-justice-inequality-perception-and-meritocracy}

\subsection{The justification of market
inequality}\label{the-justification-of-market-inequality}

Market justice has been discussed in the literature as a normative
principle that legitimates the distribution of economic rewards based on
individual merit and effort. The most detailed theoretical account of
this concept is provided by Lane
(\citeproc{ref-lane_market_1986}{1986}), who defines market justice as a
system of ``earned deserts,'' whereby individuals are seen as entitled
to their outcomes insofar as these are a product of personal
productivity and competition. This view directly contrasts with
political justice, which emphasizes redistributive mechanisms based on
the principles of equality and need, and is often materialized through
welfare policies. For Lane (\citeproc{ref-lane_market_1986}{1986}),
market justice relies on the conception that markets are neutral and
self-regulating arenas, where individuals are treated fairly because
they face the same formal rules of engagement. What matters is not
whether everyone ends up with the same, but that everyone has the chance
to compete, and that success reflects effort and ability. Consequently,
the legitimacy of market justice stems from the assumption that
inequalities are not only inevitable but fair---so long as the rules are
transparent and opportunities are open. These beliefs align with a
broader ideological framework that values individual autonomy, private
property, and limited government intervention. In this way, market
justice provides a moral justification for inequality by framing it as a
necessary outcome of personal responsibility and societal efficiency.

Empirical studies have developed different strategies to capture and
operationalize market justice preferences. A common approach in the
literature is to gauge attitudes toward the legitimacy of inequality in
specific domains, especially when linked to income differences. Kluegel
and Smith (\citeproc{ref-kluegel_beliefs_1981}{1981}) and later Kluegel
et al. (\citeproc{ref-kluegel_legitimation_1999}{1999}) laid the
groundwork for measuring beliefs in market justice through survey items
that assess whether people support economic inequality when it reflects
personal effort and skills. Over time, this approach has been extended
beyond income to include other market-mediated outcomes, such as
education, healthcare, and/or pensions. For example, Von Dem Knesebeck
et al. (\citeproc{ref-vondemknesebeck_are_2016}{2016}) and Immergut and
Schneider (\citeproc{ref-immergut_it_2020}{2020}) examine whether
citizens consider it fair that individuals with higher incomes can
access better healthcare, while Lee and Stacey
(\citeproc{ref-lee_fairness_2023}{2023}) apply a similar method in the
context of education in Australia. These studies usually rely on a
survey item asking respondents to evaluate the fairness of income-based
access to welfare services, allowing for comparing justice perceptions
across different contexts. Some research, like that of Lindh
(\citeproc{ref-lindh_public_2015}{2015}) and Svallfors
(\citeproc{ref-svallfors_political_2007}{2007}), uses these indicators
in combination to measure broader preferences for market-based
distribution in welfare systems. More recently, Castillo et al.
(\citeproc{ref-castillo_socialization_2024}{2024}) introduced a
single-item composite measure of market justice to assess student
attitudes toward income-based access to education, healthcare, and
pensions in Chile. These different operationalizations all aim to
capture the extent to which individuals accept inequality when it is
framed as a reflection of market outcomes.

The study of market justice preferences has increasingly focused on how
they are shaped by individuals' socioeconomic position, normative
beliefs, and the institutional context in which they are embedded.
Across the literature, there is strong empirical evidence suggesting
that individuals in more advantaged socioeconomic positions---those with
higher income, education, and occupational status---are more likely to
endorse market-based distribution as fair and legitimate
(\citeproc{ref-koos_moral_2019}{Koos \& Sachweh, 2019};
\citeproc{ref-svallfors_political_2007}{Svallfors, 2007}). This tendency
reflects not only material self-interest but also a broader moral
economy, in which winners of the market system internalize
justifications for the status quo. At the same time, normative and
ideological orientations---such as conservative, liberal, or neoliberal
values---are consistently associated with greater support for
meritocratic principles and skepticism toward redistribution. This is
particularly visible in contexts such as Chile and Australia, where
studies by Castillo et al.
(\citeproc{ref-castillo_socialization_2024}{2024}) and Lee and Stacey
(\citeproc{ref-lee_fairness_2023}{2023}) show that those with
right-leaning political preferences tend to view income-based access to
welfare services as legitimate. Beyond individual-level characteristics,
country-level institutions also play a central role. In liberal welfare
regimes like those of the United States or the United Kingdom, market
justice preferences are more widespread, while in coordinated or
social-democratic regimes---such as Sweden or Germany---citizens are
generally more critical of market-based inequalities
(\citeproc{ref-immergut_it_2020}{Immergut \& Schneider, 2020};
\citeproc{ref-lindh_public_2015}{Lindh, 2015}). These findings suggest
that market justice is not a fixed belief but a contextually shaped
orientation, mediated by structural conditions, cultural values, and the
historical development of welfare institutions.

\subsection{Perception of inequality}\label{perception-of-inequality}

Perceptions of inequality have been associated to attitudes about market
justice (\citeproc{ref-kluegel_social_1995a}{Kluegel et al., 1995};
\citeproc{ref-lindh_public_2015}{Lindh, 2015}). Research indicates that
lower perceived inequality can reinforce support for market-based
distributive arrangements by suggesting that the system is fair and that
outcomes reflect effort and ability (\citeproc{ref-kuhn_eye_2011}{Kuhn,
2011}). In contrast, when inequality is perceived as excessive or
structurally determined, individuals could question the legitimacy of
market justice and become more supportive of redistributive policies
(\citeproc{ref-castillo_perception_2022}{Castillo et al., 2022};
\citeproc{ref-garcia-sanchez_vicious_2019}{García-Sánchez et al.,
2019}).

Perceptions of inequality refer to individuals' subjective evaluations
of the extent, causes, and consequences of income and wealth
disparities. Unlike objective measures such as the Gini index, perceived
inequality captures how individuals make sense of distributive
hierarchies in their everyday lives, shaped by reference groups, social
comparisons, and information environments
(\citeproc{ref-garcia-castro_perceiving_2020}{García-Castro et al.,
2020}; \citeproc{ref-gimpelson_misperceiving_2018}{Gimpelson \&
Treisman, 2018}; \citeproc{ref-mijs_stratified_2016}{Mijs, 2016}).
Scholars have proposed multiple dimensions of perceived inequality,
including its magnitude (how significant are the gaps), vertical
structure (between which groups), trend over time (increasing or
decreasing), and legitimacy (whether it is just or not)
(\citeproc{ref-engelhardt_what_2018}{Engelhardt \& Wagener, 2018};
\citeproc{ref-garcia-sanchez_vicious_2019}{García-Sánchez et al.,
2019}). These dimensions encompass both cognitive and normative aspects
of perceptions of inequality and can vary across societies and social
groups, depending on exposure, ideology, and personal experience
(\citeproc{ref-castillo_perception_2022}{Castillo et al., 2022};
\citeproc{ref-garcia-sanchez_perceptions_2018}{García-Sánchez et al.,
2018}).

Among the different approaches to measuring perceived inequality, one of
the most widely used is the estimation of wage gaps between occupational
extremes, such as between a CEO and a manual worker. This type of item
provides a concrete frame that respondents can relate to more easily
than abstract questions about national income distribution
(\citeproc{ref-castillo_percepcion_2019}{Castillo et al., 2012};
\citeproc{ref-easterbrook_social_2021}{Easterbrook, 2021};
\citeproc{ref-willis_legitimacy_2015}{Willis et al., 2015}). While it
enables the estimation of inequality using simple heuristics, this
method is not without challenges. For instance, people often lack
reliable knowledge about the earnings of those at the top of the income
ladder, which leads to high variability in responses and the use of
biased mental shortcuts (\citeproc{ref-knell_perceptions_2020}{Knell \&
Stix, 2020}). Despite this, perceived wage gaps are strong predictors of
political attitudes
(\citeproc{ref-garcia-sanchez_perceptions_2018}{García-Sánchez et al.,
2018}; \citeproc{ref-pedersen_attitudes_2019}{Pedersen \& Mutz, 2019}),
making them a valuable tool for understanding public responses to
economic disparities.

Another relevant critique of this approach lies in its conflation of
different psychological constructs. Many surveys assess perceived
inequality through Likert-type items that ask respondents to agree or
disagree with statements, such as ``income differences are too large,''
which captures general concern or discomfort rather than a specific
perception (\citeproc{ref-Castillo2011}{Castillo, 2011};
\citeproc{ref-garcia-sanchez_vicious_2019}{García-Sánchez et al.,
2019}). These items mix cognitive estimations with affective
evaluations, complicating the interpretation of what respondents
perceive versus what they morally reject. As a result, the conceptual
clarity between perceived inequality and inequality aversion remains
blurred in many empirical studies. To address this limitation, recent
work has emphasized the need to distinguish between absolute and
comparative measures, as well as between ideal and actual estimates of
economic gaps (\citeproc{ref-auspurg_why_2017}{Auspurg et al., 2017};
\citeproc{ref-garcia-sanchez_creencias_2022}{García-Sánchez \& De
Carvalho, 2022}).

\subsection{Perception of meritocracy}\label{perception-of-meritocracy}

Meritocracy constitutes a central ideological framework for legitimizing
different types od social inequality, for instance through market
justice beliefs. Rooted in the belief that rewards and positions should
be allocated based on individual effort and talent, meritocracy operates
as a normative ideal and a descriptive belief about how society
functions. As initially conceptualized by Michael Young
(\citeproc{ref-young_rise_1962}{1962}), the term was meant to critique a
system in which merit-based stratification becomes a new form of
inequality. However, over time, meritocracy has been widely supported in
many societies as a fair and desirable principle of distribution,
particularly within liberal democracies and market-oriented economies
(\citeproc{ref-mijs_paradox_2021}{Mijs, 2021};
\citeproc{ref-sandel_tyranny_2020}{Sandel, 2020}). From a sociological
perspective, the belief in meritocracy is more than a cognitive
assessment; it reflects a moral lens through which individuals interpret
inequality. People who believe that success results from hard work and
talent are more likely to view social and economic disparities as
legitimate (\citeproc{ref-batruch_belief_2023}{Batruch et al., 2023};
\citeproc{ref-castillo_percepcion_2019}{Castillo et al., 2012}).
Conversely, if they see outcomes as driven by luck, social origin, or
systemic barriers, inequality is more likely to be perceived as unjust.
This distinction becomes crucial in societies with persistent structural
inequality, where public narratives often emphasize personal
responsibility and merit while overlooking entrenched disadvantages.

We adopt a multidimensional perspective on meritocracy, distinguishing
between two key dimensions: effort-based and talent-based perceptions.
This distinction is essential, as it captures different pathways through
which individuals justify inequality. Effort-based meritocracy
emphasizes hard work and perseverance as the basis for social rewards,
aligning closely with cultural narratives of personal responsibility. A
talent-based meritocracy, by contrast, emphasizes intelligence and
innate abilities, which are often perceived as less malleable and more
unequally distributed. Both dimensions have been shown to correlate with
acceptance of inequality, but they may carry distinct implications for
how inequality is justified in specific domains
(\citeproc{ref-castillo_multidimensional_2023}{Castillo et al., 2023}).
The relevance of this distinction is supported by recent studies, which
show that individuals respond differently to these dimensions. For
instance, perceptions that effort is rewarded in society are more
strongly associated with positive evaluations of fairness and acceptance
of unequal outcomes (\citeproc{ref-batruch_belief_2023}{Batruch et al.,
2023}). This may be because effort is seen as a controllable and morally
virtuous trait, whereas talent is often perceived as a natural
advantage. Consequently, effort-based meritocracy is likely more potent
in legitimizing inequality, particularly in neoliberal contexts.

These dimensions of meritocracy reflect how respondents perceive
society's distributive logic, regardless of whether they personally
endorse meritocratic principles. This distinction aligns with recent
findings indicating that individuals distinguish between how merit is
perceived in society and how it should ideally operate, which in turn
shapes their preferences for redistribution and justice
(\citeproc{ref-tejero-peregrina_perceived_2025}{Tejero-Peregrina et al.,
2025}) Meritocratic beliefs serve as symbolic justifications for unequal
outcomes, particularly when access is stratified by income or social
opportunity. Prior studies have shown that individuals who perceive
higher levels of meritocracy tend to express stronger support for
unequal distributions that reflect market outcomes
(\citeproc{ref-castillo_percepcion_2019}{Castillo et al., 2012};
\citeproc{ref-castillo_socialization_2024}{Castillo et al., 2024}).

In addition to influencing individual attitudes toward inequality,
meritocratic beliefs can contribute to social division and the
stigmatization of disadvantaged groups. Recent research has demonstrated
that exposure to meritocratic narratives can reinforce the belief that
poverty results from individual failings rather than systemic
conditions, reducing support for redistributive measures and increasing
the stigmatization of the poor (\citeproc{ref-hoyt_mindsets_2023}{Hoyt
et al., 2023}). This reinforces negative stereotypes and reduces empathy
toward individuals from lower socioeconomic backgrounds. Moreover,
Busemeyer et al. (\citeproc{ref-busemeyer_positive_2021}{2021}) argue
that meritocratic narratives can serve as feedback mechanisms that shape
public opinion and well-being by framing individuals' understanding of
welfare outcomes as deserved or undeserved within existing institutional
structures. This psychological mechanism highlights the normative power
of meritocracy in stabilizing unequal systems by shaping political
attitudes and personal perceptions of success and failure.

\subsection{This study}\label{this-study}

Building upon the previous literature, this study proposes that
attitudes toward market justice are shaped by a dynamic interplay
between individuals' perceptions of economic inequality and their
beliefs in meritocracy. Specifically, we argue that both perceptions
independently and interactively influence the extent to which
individuals endorse market-based distributions of social goods and
services in Chile.

First, consistent with previous findings, we expect that a lower
perception of economic inequality will be associated with stronger
market justice preferences. When individuals perceive smaller economic
gaps, they are more likely to view market mechanisms as fair and
legitimate, reinforcing acceptance of outcomes based on competition and
ability to pay. Conversely, a heightened perception of inequality may
erode confidence in market fairness, weakening support for market-based
distribution. This relationship is particularly relevant in the context
of Chile, where the neoliberal economic model has been a dominant force
shaping public attitudes toward inequality and justice.

Second, stronger meritocratic beliefs are expected to be positively
associated with market justice preferences. Individuals who believe that
effort and talent primarily determine success are more likely to justify
unequal outcomes and endorse the notion that markets allocate resources
fairly according to individual merit. This aligns with the idea that
meritocratic beliefs serve as a moral framework that legitimizes
market-based inequalities, as individuals perceive the system as just
when they believe that rewards are based on individual merit. This is
particularly relevant in contexts where neoliberal ideologies dominate,
as they often emphasize individual responsibility and competition as the
basis for social order.

Third, we propose that perceptions of meritocracy and perceptions of
economic inequality interact in shaping market justice attitudes.
Specifically, we argue that the legitimizing effect of meritocratic
beliefs on market justice preferences is moderated by perceptions of
economic inequality: when perceived inequality is low, the positive
association between meritocratic beliefs and market justice preferences
will be stronger. However, when perceived inequality is high, this
association will weaken, as greater awareness of large economic gaps may
challenge the view that outcomes are purely merit-based.

Additionally, this study examines whether these relationships vary over
time in response to significant social and political events. In
particular, we explore whether the political outburst of 2019 and the
subsequent constitutional processes in Chile have altered how
perceptions of inequality and meritocracy shape market justice
attitudes. We expect that after these critical events, perceptions of
inequality may have a stronger negative effect on market justice
preferences, reflecting increased societal questioning of market-based
allocation mechanisms.

Based on these arguments, we propose the following hypotheses:

\(H1\): Greater perceived economic inequality is associated with
stronger market justice preferences.

\(H2\): Stronger meritocratic beliefs are associated with stronger
market justice preferences.

\(H3\): The positive association between meritocratic beliefs and market
justice preferences is moderated by perceived economic inequality;
specifically, this association is weaker when perceived economic
inequality is high.

\(H4\): The effects described in H1--H3 are attenuated after major
social mobilizations (2019--2022), reflecting increased critical views
of the market's role in allocating social goods.

\section{Data, Variables and Methods}\label{data-variables-and-methods}

\subsection{Data}\label{data}

This study draws on data from the Chilean Longitudinal Social Survey
(ELSOC), a panel study collected annually from 2016 to 2023. The survey
evaluates how individuals think, feel, and behave regarding conflict and
social cohesion in Chile. ELSOC employs a probabilistic, stratified,
clustered, multistage sampling design encompassing major urban centers
(Santiago, Valparaíso, and Concepción) and smaller cities. The target
population includes women and men aged 18--75 who are habitual residents
of private dwellings. The first wave included 2,927 participants,
representing northern and southern regions, covering 77\% of the
country's total population and 93\% of its urban population, with a
response rate of 62.4\% (\citeproc{ref-elsoc_estudio_2022}{ELSOC,
2022}).

The survey has been conducted yearly since 2016, except in 2020, when it
was suspended due to the pandemic. Waves 2016, 2017, 2018, 2019, 2022,
and 2023 used computer-assisted personal interviewing (CAPI), while a
reduced wave in 2021 employed computer-assisted telephone interviews
(CATI). Wave 3 included a refreshment sample (N = 1,519) to counter
attrition, but these cases are not used here to capture longer response
trends. Between wave 1 and wave 7, attrition amounted to 40\%, achieving
a final sample of N = 1,741. Longitudinal weights adjust for both the
sampling design and potential biases arising from systematic
non-participation. Further details on sampling, attrition, and weighting
can be found at \url{https://coes.cl/encuesta-panel/}, and the dataset
is publicly available at
\url{https://dataverse.harvard.edu/dataverse/elsoc}.

\subsection{Variables}\label{variables}

\textbf{Market justice preferences}: The dependent variable in this
study is preferences for market justice. This construct is
operationalized through three items that capture how strongly
individuals justify conditioning access to social services---healthcare,
pensions, and education---on income. Specifically, the justification of
inequality in healthcare is assessed by the question: ``Is it fair in
Chile that people with higher incomes can access better healthcare than
people with lower incomes?'' The same question is posed for pensions and
education. In all cases, respondents indicate their level of agreement
on a five-point Likert scale ranging from 1 (``strongly disagree'') to 5
(``strongly agree''). Additionally, we include a composite measure of
``market justice preferences'', calculated as the average of these three
items (\(\alpha\) = 0.84). This index ranges from 1 to 5, with higher
values indicating stronger preferences for market justice (see
Table~\ref{tbl-summary1}).

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3367}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3367}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2143}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1122}}@{}}
\caption{Dependent variables for the first wave
(2016)}\label{tbl-summary1}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Label
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Stats / Values
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Freqs (\% of Valid)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Valid
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Label
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Stats / Values
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Freqs (\% of Valid)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Valid
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Health distributive justice &
\begin{minipage}[t]{\linewidth}\raggedright
1. Strongly desagree\\
2. Desagree\\
3. Neither agree nor desagre\\
4. Agree\\
5. Strongly agree\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
554 (37.5\%)\\
713 (48.3\%)\\
61 ( 4.1\%)\\
131 ( 8.9\%)\\
18 ( 1.2\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
1477\\
(100.0\%)\strut
\end{minipage} \\
Pension distributive justice &
\begin{minipage}[t]{\linewidth}\raggedright
1. Strongly desagree\\
2. Desagree\\
3. Neither agree nor desagre\\
4. Agree\\
5. Strongly agree\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
425 (28.8\%)\\
704 (47.7\%)\\
106 ( 7.2\%)\\
219 (14.8\%)\\
23 ( 1.6\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
1477\\
(100.0\%)\strut
\end{minipage} \\
Education distributive justice &
\begin{minipage}[t]{\linewidth}\raggedright
1. Strongly desagree\\
2. Desagree\\
3. Neither agree nor desagre\\
4. Agree\\
5. Strongly agree\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
519 (35.1\%)\\
764 (51.7\%)\\
71 ( 4.8\%)\\
112 ( 7.6\%)\\
11 ( 0.7\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
1477\\
(100.0\%)\strut
\end{minipage} \\
Market justice preferences & \begin{minipage}[t]{\linewidth}\raggedright
Mean (sd) : 2 (0.8)\\
min \textless{} med \textless{} max:\\
1 \textless{} 2 \textless{} 5\\
IQR (CV) : 1 (0.4)\strut
\end{minipage} & 12 distinct values &
\begin{minipage}[t]{\linewidth}\raggedright
1477\\
(100.0\%)\strut
\end{minipage} \\
\end{longtable}

\textbf{Perception of economic inequality}: The main independent
variable refers to the perception of economic inequality, measured
through the perceived wage gap.
(\citeproc{ref-castillo_cual_2009}{Castillo, 2009};
\citeproc{ref-gijsberts_thelegitimation_1999}{Gijsberts, 1999};
\citeproc{ref-hadler_why_2005}{Hadler, 2005}). This measure is derived
from the salary gap between the perceived salaries of jobs at opposite
ends of the occupational hierarchy. Specifically, it relies on the
division between the perceived salary of a large-company president and
that of an unskilled worker (\citeproc{ref-Castillo2011}{Castillo,
2011}). Higher values of this term indicate a greater perception of
economic inequality between occupations located at the extremes of the
status continuum. This measure includes a logarithmic term in order to
adjust income magnitudes (usually fewer cases with high income):

\[
\text{perceived wage gap} = \log_{10}\left(\frac{\text{perceived salary of a large-company president}}{\text{perceived salary of an unskilled worker}}\right)
\]

\textbf{Perception of Meritocracy}: this variable is operationalized
through two components, namely effort and talent
(\citeproc{ref-young_rise_1962}{Young, 1962}). The item used to gauge
effort is: ``In Chile, people are rewarded for their efforts,'' while
the item for talent is: ``In Chile, people are rewarded for their
intelligence and skills''. In both cases, respondents indicate their
level of agreement on a five-point Likert scale, ranging from 1
(``strongly disagree'') to 5 (``strongly agree'').

Table 2 shows the independent variables used, their response categories
and their frequencies.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3774}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3113}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2075}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1038}}@{}}
\caption{Independent variables for the first wave
(2016)}\label{tbl-summary2}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Label
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Stats / Values
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Freqs (\% of Valid)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Valid
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Label
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Stats / Values
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Freqs (\% of Valid)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Valid
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Inequality gap perception & \begin{minipage}[t]{\linewidth}\raggedright
Mean (sd) : 3.7 (1.1)\\
min \textless{} med \textless{} max:\\
0.4 \textless{} 3.7 \textless{} 6.9\\
IQR (CV) : 1.6 (0.3)\strut
\end{minipage} & 296 distinct values &
\begin{minipage}[t]{\linewidth}\raggedright
1501\\
(100.0\%)\strut
\end{minipage} \\
People are rewarded for their efforts &
\begin{minipage}[t]{\linewidth}\raggedright
1. Strongly desagree\\
2. Desagree\\
3. Neither agree nor desagre\\
4. Agree\\
5. Strongly agree\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
169 (11.3\%)\\
693 (46.2\%)\\
263 (17.5\%)\\
328 (21.9\%)\\
48 ( 3.2\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
1501\\
(100.0\%)\strut
\end{minipage} \\
People are rewarded for their intelligence &
\begin{minipage}[t]{\linewidth}\raggedright
1. Strongly desagree\\
2. Desagree\\
3. Neither agree nor desagre\\
4. Agree\\
5. Strongly agree\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
134 ( 8.9\%)\\
617 (41.1\%)\\
294 (19.6\%)\\
401 (26.7\%)\\
55 ( 3.7\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
1501\\
(100.0\%)\strut
\end{minipage} \\
\end{longtable}

\emph{Controls}

Sociodemographic and attitudinal variables are included to control for
potential composition effects in the population. In terms of
sociodemographic characteristics, we incorporate per capita household
income quintile, educational level (1=Less than Universitary,
2=Universitary), age (in years), and sex (1=Male, 2=Female), which have
been previously shown to influence market justice preferences
significantly (\citeproc{ref-castillo_socialization_2024}{Castillo et
al., 2024}; \citeproc{ref-lindh_public_2015}{Lindh, 2015}). Regarding
attitudinal variables, we include political identification (1=Left,
2=Center, 3=Right, 4=No identification) and subjective social status
(ranging from 1 to 10) as they may confound the relationship between
market justice preferences and perceptions of inequality and meritocracy
(\citeproc{ref-schneider_poverty_2015}{Schneider \& Castillo, 2015}).

\subsection{Methods}\label{methods}

Given the data's hierarchical structure, in which observations are
nested in survey waves, we employ longitudinal multilevel linear models
(\citeproc{ref-singer_applied_2009}{Singer \& Willett, 2009}). In a
panel-data framework, within-person effects capture how shifts in
individual-level variables across waves are associated with variations
in market justice preferences. By contrast, between-person effects focus
on differences among individuals, explaining how long-term (or average)
values relate to overall levels of market justice preferences.

To estimate within-person effects, we use group-mean centering, where
each respondent functions as the ``group'' (i.e., observations nested
within persons). Meanwhile, the between-person effects are derived from
each individual's average on these variables, calculated across the
waves of panel data.

All the analyses were conducted using R software and the \emph{lme4}
package (\citeproc{ref-bates_fitting_2015}{Bates et al., 2015}).

\section{Results}\label{results}

\subsection{Descriptive}\label{descriptive}

Figure~\ref{fig-alluvial} shows the annual frequencies of market justice
preferences for healthcare, pensions, and education from 2016 to 2023.
Each year presents stacked percentage frequencies, and the flows between
them reflect opinion changes among the same individuals from one year to
the next, given that we are using panel data. For instance, of the
40.8\% who strongly disagreed with justifying inequality in healthcare
in 2019, around 24.3\% maintained that position in 2022, while the
remaining 16.5\% shifted toward other response categories---primarily
moving into disagreement rather than strong disagreement. Overall, more
than half of the respondents exhibit a high level of disagreement
(disagree + strongly disagree) with inequality in these three social
service areas over time. Despite this general pattern, recent waves show
a slight decrease in disagreement and a corresponding rise in support
for inequality. Specifically, in healthcare and education, although
disagreement remains substantial, agreement (agree + strongly agree)
increased from 7.4\% and 7.2\% in 2019 to 13.1\% and 14.2\% in 2023,
respectively. This shift is most evident in pensions, where the combined
agree/strongly agree category grew by about 10 percentage points, from
16.9\% in 2016 to 28\% in 2023.

\begin{figure}[H]

\caption{\label{fig-alluvial}Change in the justification of inequality
in healthcare, pensions and education over time (2016-2023)}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-alluvial-1.pdf}

}

\end{figure}%

Regarding the main variables of this study, Figure~\ref{fig-meanchange}
depicts their average changes over the years. We observe an increase in
the average level of market justice preferences in the most recent
waves, which begins in 2019. The highest average consistently appears
for perceived economic inequality, although this variable shows a
downward trend of roughly one point over time. Interestingly, while
perceptions of economic inequality declined in the latest measurements
(2022--2023), market justice preferences increased. The meritocracy
measures remain stable, though the perception that individuals are
rewarded for intelligence is slightly higher than the perception that
they are rewarded for their effort.

\begin{figure}[H]

\caption{\label{fig-meanchange}Change in the mean of market justice
preferences, economic inequality perception, and meritocracy
(2016-2023)}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{paper_files/figure-pdf/fig-meanchange-1.pdf}

}

\end{figure}%

Figure~\ref{fig-matrix} presents the correlation matrix for the main
variables in the latest wave (2023). Overall, the coefficients range
from low to moderate. The association between market justice preferences
and economic inequality perception is negative but small and
statistically significant (\(r\) = -0.05, \(p\) \textless{} .05). By
contrast, market justice preferences positively and significantly
correlate with the two meritocracy variables (\(r\) = 0.15, \(p\)
\textless{} .01; \(r\) = 0.13, \(p\) \textless{} .01). Perceived
economic inequality has negative and nonsignificant correlations with
both meritocracy perceptions (\(r\) = -0.04, \(p\) \textgreater{} .05;
\(r\) = -0.04, \(p\) \textgreater{} .05). Finally, the two meritocracy
variables exhibit a strong positive association with each other (\(r\) =
0.69, \(p\) \textless{} .01).

\begin{figure}[H]

\caption{\label{fig-matrix}Correlation matrix of the main variables for
the last wave (2023)}

\centering{

\includegraphics[width=0.9\textwidth,height=\textheight]{paper_files/figure-pdf/fig-matrix-1.pdf}

}

\end{figure}%

\subsection{Multilevel models}\label{multilevel-models}

Table~\ref{tbl-modelos} presents the results of the multilevel models
estimated for market justice preferences, examining both individuals
(within) and group-level (between) effects. The intraclass correlation
(\citeproc{ref-hoxMultilevelAnalysisTechniques2017a}{Hox et al., 2017})
from the empty model (see Supplementary Material), which decomposes the
variance of market justice preferences, is 0.23, indicating that
approximately 23\% of the variation is attributable to differences
between individuals. Complementary, 77\% of the variation corresponds to
within-individual differences over time.

\begin{table}

\caption{\label{tbl-modelos}Longitudinal multilevel models for market
justice preferences}

\centering{

\begin{center}
\scalebox{0.79}{
\begin{tabular}{l c c c c c c c}
\hline
 & Model 1 & Model 2 & Model 3 & Model 4 & Model 5 & Model 6 & Model 7 \\
\hline
Intercept                            & $1.968^{***}$  & $1.973^{***}$  & $1.985^{***}$  & $1.982^{***}$  & $2.122^{***}$  & $1.408^{***}$  & $1.499^{***}$  \\
                                     & $(0.021)$      & $(0.032)$      & $(0.032)$      & $(0.032)$      & $(0.069)$      & $(0.096)$      & $(0.114)$      \\
Wave (Ref.= 2016)                    &                &                &                &                &                &                &                \\
                                     &                &                &                &                &                &                &                \\
\quad Wave 2017                      & $-0.167^{***}$ &                &                &                &                &                &                \\
                                     & $(0.027)$      &                &                &                &                &                &                \\
\quad Wave 2018                      & $-0.025$       &                &                &                &                &                &                \\
                                     & $(0.027)$      &                &                &                &                &                &                \\
\quad Wave 2019                      & $-0.046$       &                &                &                &                &                &                \\
                                     & $(0.027)$      &                &                &                &                &                &                \\
\quad Wave 2022                      & $0.285^{***}$  &                &                &                &                &                &                \\
                                     & $(0.028)$      &                &                &                &                &                &                \\
\quad Wave 2023                      & $0.284^{***}$  &                &                &                &                &                &                \\
                                     & $(0.027)$      &                &                &                &                &                &                \\
Wave                                 &                & $-0.075^{***}$ & $-0.078^{***}$ & $-0.079^{***}$ & $-0.079^{***}$ & $-0.079^{***}$ & $-0.079^{***}$ \\
                                     &                & $(0.018)$      & $(0.018)$      & $(0.018)$      & $(0.018)$      & $(0.018)$      & $(0.018)$      \\
Wave$^2$                             &                & $0.017^{***}$  & $0.018^{***}$  & $0.018^{***}$  & $0.018^{***}$  & $0.018^{***}$  & $0.018^{***}$  \\
                                     &                & $(0.002)$      & $(0.002)$      & $(0.002)$      & $(0.002)$      & $(0.002)$      & $(0.002)$      \\
Perception inequality (WE)           &                &                & $-0.034^{***}$ & $-0.029^{**}$  & $-0.029^{**}$  & $-0.029^{**}$  & $-0.029^{**}$  \\
                                     &                &                & $(0.010)$      & $(0.010)$      & $(0.010)$      & $(0.010)$      & $(0.010)$      \\
Merit: Effort (WE)                   &                &                &                & $0.071^{***}$  & $0.071^{***}$  & $0.070^{***}$  & $0.070^{***}$  \\
                                     &                &                &                & $(0.012)$      & $(0.012)$      & $(0.012)$      & $(0.012)$      \\
Merit: Talent (WE)                   &                &                &                & $0.028^{*}$    & $0.028^{*}$    & $0.028^{*}$    & $0.027^{*}$    \\
                                     &                &                &                & $(0.012)$      & $(0.012)$      & $(0.012)$      & $(0.012)$      \\
Perception inequality (BE)           &                &                &                &                & $-0.042^{*}$   & $-0.005$       & $-0.032$       \\
                                     &                &                &                &                & $(0.018)$      & $(0.018)$      & $(0.018)$      \\
Merit: Effort (BE)                   &                &                &                &                &                & $0.186^{***}$  & $0.178^{***}$  \\
                                     &                &                &                &                &                & $(0.032)$      & $(0.031)$      \\
Merit: Talent (BE)                   &                &                &                &                &                & $0.040$        & $0.019$        \\
                                     &                &                &                &                &                & $(0.032)$      & $(0.031)$      \\
\hline
Controls                             & No             & No             & No             & No             & No             & No             & Yes            \\
BIC                                  & $20780.594$    & $20757.596$    & $20761.847$    & $20704.882$    & $20714.843$    & $20633.143$    & $20727.029$    \\
Num. obs.                            & $8643$         & $8643$         & $8643$         & $8643$         & $8643$         & $8643$         & $8643$         \\
Num. groups: idencuesta              & $1687$         & $1687$         & $1687$         & $1687$         & $1687$         & $1687$         & $1687$         \\
Var: idencuesta (Intercept)          & $0.172$        & $0.207$        & $0.204$        & $0.198$        & $0.196$        & $0.172$        & $0.160$        \\
Var: Residual                        & $0.530$        & $0.497$        & $0.497$        & $0.493$        & $0.493$        & $0.492$        & $0.493$        \\
Var: idencuesta ola\_num             & $$             & $0.007$        & $0.007$        & $0.006$        & $0.006$        & $0.006$        & $0.006$        \\
Cov: idencuesta (Intercept) ola\_num & $$             & $-0.017$       & $-0.017$       & $-0.014$       & $-0.014$       & $-0.013$       & $-0.013$       \\
\hline
\multicolumn{8}{l}{\scriptsize{Note: Cells contain regression coefficients with standard errors in parentheses. $^{***}p<0.001$; $^{**}p<0.01$; $^{*}p<0.05$.}}
\end{tabular}
}
\caption{}
\label{table:coefficients}
\end{center}

}

\end{table}%

According to Model 1, which includes the survey waves to capture
intertemporal variations in the dependent variable, there is a decrease
in 2017 (\(\beta\) = -0.167, \(p\) \textless{} .001) relative to 2016,
and similarly in 2018 (\(\beta\) = -0.025, \(p\) \textgreater{} .05) and
2019 (\(\beta\) = -0.046, \(p\) \textgreater{} .05), although the latter
effects are not statistically significant. In contrast, in the more
recent waves of 2022 and 2023, there is a statistically significant
increase in market justice preferences (\(\beta\) = 0.285, \(p\)
\textless{} .001; \(\beta\) = 0.284, \(p\) \textless{} .001), suggesting
a non-linear effect. To model this trajectory over time, Model 2
incorporates time (survey waves) as a continuous variable, along with
its quadratic term, representing the non-linear association initially
observed in Model 1. While the linear term (survey wave) shows a
negative association, reflecting an overall decline in market
preferences over time, the positive quadratic term indicates a reversal
of this pattern in the final measurement points.

Models 3 and 4 incorporate the within-group effects (WE) of the primary
independent variables, capturing how individual changes in these
variables over time shape the dependent variable. The results in Model 3
suggest that the within effect of perceived economic inequality is
negative and statistically significant (\(p\) \textless{} .001).
Specifically, each one-point increase in an individual's perception of
economic inequality between waves is associated with a 0.034 point
decrease in market justice preferences. Model 4 shows that meritocratic
beliefs operate in the opposite direction. An upward shift in the
perception that effort is rewarded exerts a positive within effect
(\(\beta\) = 0.071, \(p\) \textless{} .001), and a parallel increase in
the perception that intelligence and ability are rewarded is likewise
associated with higher market-justice preferences (\(\beta\) = 0.028,
\(p\) \textless{} .05). Taken together, these results suggest that
individuals who increasingly perceive meritocracy---whether through
effort or talent---tend to hold stronger market justice preferences.

When examining the between-group effects (BE) in Model 5 and 6, which
capture differences between individuals in the average of the main
variables, a similar pattern emerges. Individuals who perceive higher
levels of economic inequality tend to prefer less market justice
(\(\beta\) = -0.042, \(p\) \textless{} .05). In Model 6, the
meritocratic belief that effort is rewarded is positively associated
with market justice preferences (\(\beta\) = 0.186, \(p\) \textless{}
.001). , whereas the belief that talent is rewarded shows a positive but
non-significant coefficient (\(\beta\) = 0.040). Notably, once the
meritocratic variables are included, the negative coefficient for
perceived economic inequality remains but is no longer statistically
significant.

Model 7 adds the control variables. With the exception of the
between-effect for perceived economic inequality---which becomes
nonsignificant---the within- and between-effects of the principal
predictors retain both their direction and statistical significance,
confirming the robustness of the associations (see Supplementary
Material for effects of control variables).

\begin{table}

\caption{\label{tbl-interactions1}Interactions for meritocracy,
perceived economic inequality and market justice preferences}

\centering{

\begin{center}
\scalebox{0.95}{
\begin{tabular}{l c c}
\hline
 & Model 8 & Model 9 \\
\hline
Intercept                                                & $1.511^{***}$ & $2.309^{***}$ \\
                                                         & $(0.114)$     & $(0.287)$     \\
Merit: Effort (WE) x Perception inequality (WE)          & $0.018$       &               \\
                                                         & $(0.016)$     &               \\
Merit: Talent (WE) x Perception inequality (WE)          & $-0.002$      &               \\
                                                         & $(0.016)$     &               \\
Merit: Effort (BE) x Perception inequality (BE)          &               & $0.016$       \\
                                                         &               & $(0.045)$     \\
Merit: Talent (BE) x Perception inequality (BE)          &               & $0.069$       \\
                                                         &               & $(0.045)$     \\
\hline
Controls                                                 & Yes           & Yes           \\
BIC                                                      & $20858.742$   & $20799.038$   \\
Num. obs.                                                & $8643$        & $8643$        \\
Num. groups: idencuesta                                  & $1687$        & $1687$        \\
Var: idencuesta (Intercept)                              & $0.147$       & $0.141$       \\
Var: idencuesta merit\_effort\_cwc                       & $0.017$       & $$            \\
Var: idencuesta merit\_talent\_cwc                       & $0.018$       & $$            \\
Var: idencuesta perc\_inequality\_cwc                    & $0.000$       & $$            \\
Cov: idencuesta (Intercept) merit\_effort\_cwc           & $-0.005$      & $$            \\
Cov: idencuesta (Intercept) merit\_talent\_cwc           & $0.015$       & $$            \\
Cov: idencuesta (Intercept) perc\_inequality\_cwc        & $-0.006$      & $$            \\
Cov: idencuesta merit\_effort\_cwc merit\_talent\_cwc    & $-0.003$      & $$            \\
Cov: idencuesta merit\_effort\_cwc perc\_inequality\_cwc & $0.000$       & $$            \\
Cov: idencuesta merit\_talent\_cwc perc\_inequality\_cwc & $0.001$       & $$            \\
Var: Residual                                            & $0.501$       & $0.527$       \\
\hline
\multicolumn{3}{l}{\scriptsize{Note: Cells contain regression coefficients with standard errors in parentheses. $^{***}p<0.001$; $^{**}p<0.01$; $^{*}p<0.05$.}}
\end{tabular}
}
\caption{}
\label{table:coefficients}
\end{center}

}

\end{table}%

Table~\ref{tbl-interactions1} reports whether perceived economic
inequality moderates the impact of meritocratic beliefs on
market-justice preferences. In neither the within-person specification
(Model 8) nor the between-person specification (Model 9) do the
interaction terms reach statistical significance, indicating that higher
perceptions of economic inequality do not meaningfully alter the effect
of meritocracy on support for market allocation of social services.

With regard to how the effects of perceived economic inequality,
meritocratic beliefs, and their interaction evolve over time\ldots{}

Valores predichos o marginales??

\section{Discussion}\label{discussion}

\section{Conclusion}\label{conclusion}

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-akyelken_urban_2020}
Akyelken, N. (2020). Urban conceptions of economic inequalities.
\emph{Regional Studies}, \emph{54}(6), 863--872.
\url{https://doi.org/10.1080/00343404.2020.1732902}

\bibitem[\citeproctext]{ref-auspurg_why_2017}
Auspurg, K., Hinz, T., \& Sauer, C. (2017). Why {Should Women Get Less}?
{Evidence} on the {Gender Pay Gap} from {Multifactorial Survey
Experiments}. \emph{American Sociological Review}, \emph{82}(1),
179--210. \url{https://doi.org/10.1177/0003122416683393}

\bibitem[\citeproctext]{ref-bates_fitting_2015}
Bates, D., Mächler, M., Bolker, B., \& Walker, S. (2015). Fitting linear
mixed-effects models using {lme4}. \emph{Journal of Statistical
Software}, \emph{67}(1), 1--48.
\url{https://doi.org/10.18637/jss.v067.i01}

\bibitem[\citeproctext]{ref-batruch_belief_2023}
Batruch, A., Jetten, J., Van De Werfhorst, H., Darnon, C., \& Butera, F.
(2023). Belief in {School Meritocracy} and the {Legitimization} of
{Social} and {Income Inequality}. \emph{Social Psychological and
Personality Science}, \emph{14}(5), 621--635.
\url{https://doi.org/10.1177/19485506221111017}

\bibitem[\citeproctext]{ref-boccardo_30_2020}
Boccardo, G. (2020). \emph{30 a{ñ}os de privatizaciones en {Chile}: Lo
que la pandemia revel{ó}} (Nodo XXI). Santiago.

\bibitem[\citeproctext]{ref-busemeyer_positive_2021}
Busemeyer, M. R., Abrassart, A., \& Nezi, R. (2021). Beyond {Positive}
and {Negative}: {New Perspectives} on {Feedback Effects} in {Public
Opinion} on the {Welfare State}. \emph{British Journal of Political
Science}, \emph{51}(1), 137--162.
\url{https://doi.org/10.1017/S0007123418000534}

\bibitem[\citeproctext]{ref-castillo_cual_2009}
Castillo, J. C. (2009). {{\textquestiondown}Cu{á}l es la brecha salarial
justa? Opini{ó}n p{ú}blica y legitimaci{ó}n de la desigualdad en Chile}.
\emph{Estudios P{ú}blicos}, (113).

\bibitem[\citeproctext]{ref-Castillo2011}
Castillo, J. C. (2011). Legitimacy of {Inequality} in a {Highly Unequal
Context}: {Evidence} from the {Chilean Case}. \emph{Social Justice
Research}, \emph{24}(4), 314--340.
\url{https://doi.org/10.1007/s11211-011-0144-5}

\bibitem[\citeproctext]{ref-castillo_perception_2022}
Castillo, J. C., García-Castro, J.-D., \& Venegas, M. (2022). Perception
of economic inequality: Concepts, associated factors and prospects of a
burgeoning research agenda. \emph{International Journal of Social
Psychology}, \emph{37}(1), 180--207.
\url{https://doi.org/10.1080/02134748.2021.2009275}

\bibitem[\citeproctext]{ref-castillo_multidimensional_2023}
Castillo, J. C., Iturra, J., Maldonado, L., Atria, J., \& Meneses, F.
(2023). A {Multidimensional Approach} for {Measuring Meritocratic
Beliefs}: {Advantages}, {Limitations} and {Alternatives} to the {ISSP
Social Inequality Survey}. \emph{International Journal of Sociology},
1--25. \url{https://doi.org/10.1080/00207659.2023.2274712}

\bibitem[\citeproctext]{ref-castillo_percepcion_2019}
Castillo, J. C., Miranda, D., \& Carrasco, D. (2012). Percepci{ó}n de
{Desigualdad Econ{ó}mica} en {Chile}: {Medici{ó}n}, {Diferencias} y
{Determinantes}. \emph{Psykhe (Santiago)}, \emph{21}(1), 99--114.
\url{https://doi.org/10.4067/S0718-22282012000100007}

\bibitem[\citeproctext]{ref-castillo_socialization_2024}
Castillo, J. C., Salgado, M., Carrasco, K., \& Laffert, A. (2024). The
{Socialization} of {Meritocracy} and {Market Justice Preferences} at
{School}. \emph{Societies}, \emph{14}(11), 214.
\url{https://doi.org/10.3390/soc14110214}

\bibitem[\citeproctext]{ref-chancel_world_2022}
Chancel, L., Piketty, T., Saez, E., \& Zucman, G. (2022). World
inequality report 2022.
https://bibliotecadigital.ccb.org.co/handle/11520/27510.

\bibitem[\citeproctext]{ref-davis_principles_2001}
Davis, K., \& Moore, W. E. (2001). Some {Principles} of
{Stratification}. In \emph{Social {Stratification}, {Class}, {Race}, and
{Gender} in {Sociological Perspective}, {Second Edition}} (2nd ed.).
Routledge.

\bibitem[\citeproctext]{ref-easterbrook_social_2021}
Easterbrook, M. J. (2021). \emph{The social psychology of economic
inequality} (43rd ed., Vol. 2021). UNU-WIDER.
\url{https://doi.org/10.35188/UNU-WIDER/2021/981-5}

\bibitem[\citeproctext]{ref-elsoc_estudio_2022}
ELSOC, S. T. (2022). Estudio {Longitudinal Social} de {Chile}. Harvard
Dataverse. \url{https://doi.org/10.7910/dvn/0kirbj}

\bibitem[\citeproctext]{ref-engelhardt_what_2018}
Engelhardt, C., \& Wagener, A. (2018). What do {Germans} think and know
about income inequality? {A} survey experiment. \emph{Socio-Economic
Review}, \emph{16}(4), 743--767.
\url{https://doi.org/10.1093/ser/mwx036}

\bibitem[\citeproctext]{ref-flores_top_2020}
Flores, I., Sanhueza, C., Atria, J., \& Mayer, R. (2020). Top {Incomes}
in {Chile}: {A Historical Perspective} on {Income Inequality},
1964--2017. \emph{Review of Income and Wealth}, \emph{66}(4), 850--874.
\url{https://doi.org/10.1111/roiw.12441}

\bibitem[\citeproctext]{ref-garcia-castro_perceiving_2020}
García-Castro, J. D., Rodríguez-Bailón, R., \& Willis, G. B. (2020).
Perceiving economic inequality in everyday life decreases tolerance to
inequality. \emph{Journal of Experimental Social Psychology}, \emph{90},
104019. \url{https://doi.org/10.1016/j.jesp.2020.104019}

\bibitem[\citeproctext]{ref-garcia-sanchez_creencias_2022}
García-Sánchez, E., \& De Carvalho, S. (2022). Las creencias que
justifican la desigualdad moderan la relaci{ó}n entre el estatus
socioecon{ó}mico y el apoyo a la redistribuci{ó}n. \emph{Revista
Internacional de Sociolog{í}a}, \emph{80}(3), e210.
\url{https://doi.org/10.3989/ris.2022.80.3.21.29}

\bibitem[\citeproctext]{ref-garcia-sanchez_attitudes_2020}
García-Sánchez, E., Osborne, D., Willis, G. B., \& Rodríguez-Bailón, R.
(2020). Attitudes towards redistribution and the interplay between
perceptions and beliefs about inequality. \emph{British Journal of
Social Psychology}, \emph{59}(1), 111--136.
\url{https://doi.org/10.1111/bjso.12326}

\bibitem[\citeproctext]{ref-garcia-sanchez_vicious_2019}
García-Sánchez, E., Van Der Toorn, J., Rodríguez-Bailón, R., \& Willis,
G. B. (2019). The {Vicious Cycle} of {Economic Inequality}: {The Role}
of {Ideology} in {Shaping} the {Relationship Between} {``{What Is}''}
and {``{What Ought} to {Be}''} in 41 {Countries}. \emph{Social
Psychological and Personality Science}, \emph{10}(8), 991--1001.
\url{https://doi.org/10.1177/1948550618811500}

\bibitem[\citeproctext]{ref-garcia-sanchez_perceptions_2018}
García-Sánchez, E., Willis, G. B., Rodríguez-Bailón, R., Palacio Sañudo,
J., David Polo, J., \& Rentería Pérez, E. (2018). Perceptions of
{Economic Inequality} and {Support} for {Redistribution}: {The} role of
{Existential} and {Utopian Standards}. \emph{Social Justice Research},
\emph{31}(4), 335--354. \url{https://doi.org/10.1007/s11211-018-0317-6}

\bibitem[\citeproctext]{ref-gijsberts_thelegitimation_1999}
Gijsberts, M. (1999). \emph{{Thelegitimation of inequality in state-
socialist and market societies, 1987 - 1996}}. Amsterdam: Thela Thesis.

\bibitem[\citeproctext]{ref-gimpelson_misperceiving_2018}
Gimpelson, V., \& Treisman, D. (2018). Misperceiving inequality.
\emph{Economics \& Politics}, \emph{30}(1), 27--54.
\url{https://doi.org/10.1111/ecpo.12103}

\bibitem[\citeproctext]{ref-hadler_why_2005}
Hadler, M. (2005). Why {Do People Accept Different Income Ratios}?: {A
Multi-level Comparison} of {Thirty Countries}. \emph{Acta Sociologica},
\emph{48}(2), 131--154. \url{https://doi.org/10.1177/0001699305053768}

\bibitem[\citeproctext]{ref-hoxMultilevelAnalysisTechniques2017a}
Hox, J. J., Moerbeek, M., \& Van de Schoot, R. (2017). \emph{Multilevel
{Analysis}: {Techniques} and {Applications}}.

\bibitem[\citeproctext]{ref-hoyt_mindsets_2023}
Hoyt, C. L., Burnette, J. L., Billingsley, J., Becker, W., \& Babij, A.
D. (2023). Mindsets of poverty: {Implications} for redistributive policy
support. \emph{Analyses of Social Issues and Public Policy},
\emph{23}(3), 668--693. \url{https://doi.org/10.1111/asap.12367}

\bibitem[\citeproctext]{ref-immergut_it_2020}
Immergut, E. M., \& Schneider, S. M. (2020). Is it unfair for the
affluent to be able to purchase {``better''} healthcare? {Existential}
standards and institutional norms in healthcare attitudes across 28
countries. \emph{Social Science \& Medicine}, \emph{267}, 113146.
\url{https://doi.org/10.1016/j.socscimed.2020.113146}

\bibitem[\citeproctext]{ref-janmaat_subjective_2013}
Janmaat, J. G. (2013). Subjective inequality: {A} review of
international comparative studies on people's views about inequality.
\emph{Archives Europeennes de Sociologie}, \emph{54}(3), 357--389.
\url{https://doi.org/10.1017/S0003975613000209}

\bibitem[\citeproctext]{ref-kluegel_social_1995a}
Kluegel, J. R., Mason, D. S., \& Wegener, B. (Eds.). (1995).
\emph{Social {Justice} and {Political Change}: {Public Opinion} in
{Capitalist} and {Post-Communist States}} (1st ed.). Routledge.

\bibitem[\citeproctext]{ref-kluegel_legitimation_1999}
Kluegel, J. R., Mason, D. S., \& Wegener, B. (1999). The {Legitimation}
of {Capitalism} in the {Postcommunist Transition}: {Public Opinion}
about {Market Justice}, 1991-1996. \emph{European Sociological Review},
\emph{15}(3), 251--283. Retrieved from
\url{https://www.jstor.org/stable/522731}

\bibitem[\citeproctext]{ref-kluegel_beliefs_1981}
Kluegel, J. R., \& Smith, E. R. (1981). Beliefs {About Stratification}.
\emph{Annual Review of Sociology}, 29--56.

\bibitem[\citeproctext]{ref-knell_perceptions_2020}
Knell, M., \& Stix, H. (2020). Perceptions of inequality. \emph{European
Journal of Political Economy}, \emph{65}, 101927.
\url{https://doi.org/10.1016/j.ejpoleco.2020.101927}

\bibitem[\citeproctext]{ref-koos_moral_2019}
Koos, S., \& Sachweh, P. (2019). The moral economies of market
societies: Popular attitudes towards market competition, redistribution
and reciprocity in comparative perspective. \emph{Socio-Economic
Review}, \emph{17}(4), 793--821.
\url{https://doi.org/10.1093/ser/mwx045}

\bibitem[\citeproctext]{ref-kuhn_eye_2011}
Kuhn, A. (2011). In the eye of the beholder: {Subjective} inequality
measures and individuals' assessment of market justice. \emph{European
Journal of Political Economy}, \emph{27}(4), 625--641.
\url{https://doi.org/10.1016/j.ejpoleco.2011.06.002}

\bibitem[\citeproctext]{ref-lane_market_1986}
Lane, R. E. (1986). Market {Justice}, {Political Justice}.
\emph{American Political Science Review}, \emph{80}(2), 383--402.
\url{https://doi.org/10.2307/1958264}

\bibitem[\citeproctext]{ref-lee_fairness_2023}
Lee, J.-S., \& Stacey, M. (2023). Fairness perceptions of educational
inequality: The effects of self-interest and neoliberal orientations.
\emph{The Australian Educational Researcher}.
\url{https://doi.org/10.1007/s13384-023-00636-6}

\bibitem[\citeproctext]{ref-lindh_public_2015}
Lindh, A. (2015). Public {Opinion} against {Markets}? {Attitudes}
towards {Market Distribution} of {Social Services} -- {A Comparison} of
17 {Countries}. \emph{Social Policy \& Administration}, \emph{49}(7),
887--910. \url{https://doi.org/10.1111/spol.12105}

\bibitem[\citeproctext]{ref-mijs_stratified_2016}
Mijs, J. (2016). Stratified {Failure}: {Educational Stratification} and
{Students}' {Attributions} of {Their Mathematics Performance} in 24
{Countries}. \emph{Sociology of Education}, \emph{89}(2), 137--153.
\url{https://doi.org/10.1177/0038040716636434}

\bibitem[\citeproctext]{ref-mijs_paradox_2021}
Mijs, J. (2021). The paradox of inequality: Income inequality and belief
in meritocracy go hand in hand. \emph{Socio-Economic Review},
\emph{19}(1), 7--35. \url{https://doi.org/10.1093/ser/mwy051}

\bibitem[\citeproctext]{ref-pedersen_attitudes_2019}
Pedersen, R. T., \& Mutz, D. C. (2019). Attitudes {Toward Economic
Inequality}: {The Illusory Agreement}. \emph{Political Science Research
and Methods}, \emph{7}(04), 835--851.
\url{https://doi.org/10.1017/psrm.2018.18}

\bibitem[\citeproctext]{ref-reynolds_perceptions_2014}
Reynolds, J., \& Xian, H. (2014). Perceptions of meritocracy in the land
of opportunity. \emph{Research in Social Stratification and Mobility},
\emph{36}, 121--137. \url{https://doi.org/10.1016/j.rssm.2014.03.001}

\bibitem[\citeproctext]{ref-sandel_tyranny_2020}
Sandel, M. J. (2020). \emph{The tyranny of merit: {What}'s become of the
common good?} (First edition). New York: {Farrar, Straus and Giroux}.

\bibitem[\citeproctext]{ref-schneider_poverty_2015}
Schneider, S. M., \& Castillo, J. C. (2015). Poverty {Attributions} and
the {Perceived Justice} of {Income Inequality} : {A Comparison} of
{East} and {West Germany}.
\url{https://doi.org/10.1177/0190272515589298}

\bibitem[\citeproctext]{ref-schroder_income_2017}
Schröder, M. (2017). Is {Income Inequality Related} to {Tolerance} for
{Inequality}? \emph{Social Justice Research}, \emph{30}(1), 23--47.
\url{https://doi.org/10.1007/s11211-016-0276-8}

\bibitem[\citeproctext]{ref-singer_applied_2009}
Singer, J. D., \& Willett, J. B. (2009). \emph{Applied longitudinal data
analysis: Modeling change and event occurence}. New York: Oxford
University Press, Incorporated.

\bibitem[\citeproctext]{ref-svallfors_political_2007}
Svallfors, S. (Ed.). (2007). \emph{The {Political Sociology} of the
{Welfare State}: {Institutions}, {Social Cleavages}, and {Orientations}}
(1st ed.). Stanford University Press.
\url{https://doi.org/10.2307/j.ctvr0qv0q}

\bibitem[\citeproctext]{ref-tejero-peregrina_perceived_2025}
Tejero-Peregrina, L., Willis, G., Sánchez-Rodríguez, Á., \&
Rodríguez-Bailón, R. (2025). From {Perceived Economic Inequality} to
{Support} for {Redistribution}: {The Role} of {Meritocracy Perception}.
\emph{International Review of Social Psychology}, \emph{38}(1), 4.
\url{https://doi.org/10.5334/irsp.1013}

\bibitem[\citeproctext]{ref-trump_income_2018}
Trump, K.-S. (2018). Income {Inequality Influences Perceptions} of
{Legitimate Income Differences}. \emph{British Journal of Political
Science}, \emph{48}(4), 929--952.
\url{https://doi.org/10.1017/S0007123416000326}

\bibitem[\citeproctext]{ref-vondemknesebeck_are_2016}
Von Dem Knesebeck, O., Vonneilich, N., \& Kim, T. J. (2016). Are health
care inequalities unfair? {A} study on public attitudes in 23 countries.
\emph{International Journal for Equity in Health}, \emph{15}(1), 61.
\url{https://doi.org/10.1186/s12939-016-0350-8}

\bibitem[\citeproctext]{ref-willis_legitimacy_2015}
Willis, G. B., Rodríguez-Bailón, R., López-Rodríguez, L., \&
García-Sánchez, E. (2015). Legitimacy {Moderates} the {Relation Between
Perceived} and {Ideal Economic Inequalities}. \emph{Social Justice
Research}, \emph{28}(4), 493--508.
\url{https://doi.org/10.1007/s11211-015-0253-7}

\bibitem[\citeproctext]{ref-wilson_role_2003}
Wilson, C. (2003). The {Role} of a {Merit Principle} in {Distributive
Justice}. \emph{The Journal of Ethics}, \emph{7}(3), 277--314.
\url{https://doi.org/10.1023/A:1024667228488}

\bibitem[\citeproctext]{ref-young_rise_1962}
Young, M. (1962). \emph{The rise of the meritocracy}. Baltimore: Penguin
Books.

\end{CSLReferences}

\end{document}