### 2025-05-21 05-discussion.qmd

---
author: "Equipo EDUMER"
bibliography: "input/bib/merit-edjust.bib"
csl: "input/bib/apa6.csl"
---

# Discussion

This study tests several key hypotheses concerning the relationships between perceptions of economic inequality, meritocratic beliefs, and attitudes toward market justice in Chile. It posits that higher perceived inequality and stronger meritocratic beliefs both independently contribute to greater support for market-based distributions of social goods. Furthermore, the study examines how perceived economic inequality moderates the influence of meritocratic beliefs on market justice preferences. Additionally, it considers how major social and political mobilizations, such as those occurring around 2019 and subsequent years, may have weakened these relationships, reflecting increased critical perspectives on systemic inequalities and market mechanisms. The subsequent discussion interprets the empirical findings in light of these hypotheses, considering their implications for understanding attitudes toward inequality and social justice in a context marked by profound social change.

The first hypothesis suggested that perceptions of economic inequality were strongly linked to attitudes toward market justice. The findings demonstrated that individuals perceiving higher income disparities showed stronger support for the distribution of social services based on individual income, aligning with theories emphasizing that awareness of inequality could fuel a critical view of such policies (Castillo et al., 2012; Mijs, 2021). This relationship reflected a moral economy where perceptions of systemic unfairness threatened the legitimacy of existing distributions, prompting calls for greater equity. Interestingly, the negative within-effects observed over time implied that as perceptions of inequality became more salient, some individuals experienced a decline in support for market justice, possibly due to increasing distrust in market mechanisms or disillusionment with the fairness of the system. This indicated that while perception heightened awareness, prolonged exposure or social mobilization could also foster skepticism about whether redistribution effectively addressed inequality or simply perpetuated systemic flaws.

Regarding the second hypothesis, the results reinforced that meritocratic perceptions ---particularly those emphasizing individual effort and talent--- were associated with stronger support for market-based distribution systems. Individuals who viewed success as primarily earned through effort were more inclined to justify existing inequalities, seeing them as outcomes of individual merit rather than systemic injustice. This aligned with research indicating that meritocratic narratives served as moral justifications that reinforced societal stratification (Castillo et al., 2012; Hoyt et al., 2023). These beliefs functioned symbolically to legitimize structural disparities, reducing support for redistributive measures by framing inequality as fair and deserved. The findings suggested that these perceptions bolstered the social hierarchy by fostering acceptance of inequality as a reflection of individual virtue rather than structural failure, which was especially pertinent in neoliberal contexts like Chile's, where market logic heavily influenced social attitudes.

The third hypothesis concerned the moderating role of perceptions of inequality on the relationship between meritocracy and market justice attitudes. The analysis indicated that the positive association between meritocratic beliefs and support for market-based distribution tended to weaken when perceptions of economic inequality were particularly high. This pattern suggested that high awareness of disparities could challenge the moral high ground of meritocracy, as individuals might question whether success was truly based on effort or whether systemic biases distorted outcomes. Although the moderation was not consistently statistically significant across all models (i.e.within subjects effects), the observed trend aligned with theoretical expectations that perceptions of inequality undermined the legitimacy of meritocratic narratives and fragmented support for existing economic arrangements. This interaction revealed the complex psychological landscape in which deeply held beliefs about effort and reward could be destabilized by heightened awareness of systemic disparities, especially during times of social upheaval or mobilization.

Finally, the fourth hypothesis addressed the impact of recent social movements, particularly from 2019 onwards, on the relationship between meritocratic beliefs, perceptions of inequality, and attitudes toward market justice. The results suggested that these movements contributed to a more critical stance toward the fairness of existing distributional mechanisms. Post-mobilization, the positive relationship between meritocratic beliefs and support for inequality appeared attenuated, indicating that collective action and public discourse during this period fostered questioning of the legitimacy of systemic inequalities justified solely through meritocratic narratives. These social mobilizations challenged dominant narratives, emphasizing structural barriers and systemic injustices, which in turn weakened the association between individual effort and deservingness. This is aligned with literature emphasizing that social protests act as catalysts for transforming societal attitudes by disrupting normative beliefs and encouraging more critical evaluations of institutional legitimacy [@tilly_contentious_2015]. The findings underscored the importance of socio-political context in shaping and reshaping attitudes toward inequality and justice, highlighting how collective action could temporarily or permanently alter the normative foundations that sustained support for market justice.