### 2025-05-22 Title-Page.Rmd
Up title page

---
title: "Perceptions of Inequality and Meritocracy: Their Interplay in Shaping Preferences for Market Justice in Chile (2016-2023)"
linestretch: '1.5'
output:
  pdf_document:
    toc: false
    number_sections: false
    keep-tex: true
geometry: margin=0.78in
header-includes:
  - \usepackage{times}           # Times New Roman
  - \usepackage{caption}
  - \captionsetup[figure, table]{labelfont={bf},labelformat={default},labelsep=period}
  - \usepackage[figurename=Fig.]{caption}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
fontsize: 12pt

---

## Authors Identification

Juan Carlos Castillo, Department of Sociology, Universidad de Chile, Santiago de Chile[^*].  
ORCID: 0000-0003-1265-7854

Andreas Laffert, Institute of Sociology, Pontificia Universidad Católica de Chile.  
ORCID: 0000-0002-9008-2454

Kevin Carrasco, Center for Social Conflict and Cohesion Studies.  
ORCID: 0000-0002-9348-8361

Julio Iturra, Bremen International Graduate School of Social Sciences, Bremen, Germany.
ORCID: 0000-0003-1993-3479

## Exclusive Submission Statement

The authors confirm that this article has been submitted exclusively to this journal.

## Acknowledgments

The authors recognize the support of the National Research and Development Agency (ANID) through the grant from the National Fund for Scientific and Technological Development (FONDECYT) N° 1210847; from the Center for Conflict and Social Cohesion Studies (COES) grant Financing Fund for Research Centers in Priority Areas (FONDAP) N° 1523A0005; and Millenium Nucleus on Digital Inequalities and Opportunities (NUDOS) NCS2022_046.

## Authors' Contributions

Author roles were classified using the Contributor Role Taxonomy [CRediT](https://credit.niso.org/) as follows: Juan Carlos Castillo: Conceptualization, Supervision, Methodology, Writing - review and editing; Andreas Laffert: Conceptualization, Methodology, Formal analysis, Software, Visualization, Writing - review and editing; Kevin Carrasco: Methodology, Formal analysis, Writing - review and editing. Julio Iturra: Conceptualization, Writing - review and editing.

## Conflict of Interest

The authors declare no conflict of interest.

## Funding

This research received funding from ANID/FONDECYT grant 1210847 and the Center for Conflict and Social Cohesion Studies (COES) through ANID/FONDAP grant 1513009.

## Data Avaibility Statement

The original data presented in the study are openly available in <https://dataverse.harvard.edu/api/access/datafile/10797987>

Access to the data, code, and analysis presented in the study are openly available in <https://github.com/justicia-distributiva/market-justice-panel-elsoc>.

[^*]: Corresponding author (juancastillov@uchile.cl - [jc-castillo.com](https://juancarloscastillo.github.io/jc-castillo/) - Ignacio Carrera Pinto 1045 Ñuñoa, Santiago de Chile).