### 2025-05-22 cover_letter.qmd

---
title: "Perceptions of Inequality and Meritocracy: Their Interplay in Shaping Preferences for Market Justice in Chile (2016-2023)"
format:
  html: 
    code-fold: true
    html-math-method: katex
    number-sections: true
  pdf: 
    geometry: margin=2cm
    template-partials: 
      - title.tex
    keep-tex: true
    number-sections: true
editor: source
link-citations: true
linestretch: 1.15       
mainfont: Times New Roman
fontsize: 12pt          
colorlinks: true
fig-height: 4
fig-width: 7.5
---

## Authors Identification

Juan Carlos Castillo, Department of Sociology, Universidad de Chile, Santiago de Chile[^*].  
ORCID: 0000-0003-1265-7854

Andreas Laffert, Institute of Sociology, Pontificia Universidad Católica de Chile.  
ORCID: 0000-0002-9008-2454

Kevin Carrasco, Center for Social Conflict and Cohesion Studies.  
ORCID: 0000-0002-9348-8361

Julio Iturra, Bremen International Graduate School of Social Sciences, Bremen, Germany.
ORCID: 0000-0003-1993-3479

## Exclusive Submission Statement

The authors confirm that this article has been submitted exclusively to this journal.

## Acknowledgments

The authors recognize the support of the National Research and Development Agency (ANID) through the grant from the National Fund for Scientific and Technological Development (FONDECYT) N° 1210847; from the Center for Conflict and Social Cohesion Studies (COES) grant Financing Fund for Research Centers in Priority Areas (FONDAP) N° 1523A0005; and Millenium Nucleus on Digital Inequalities and Opportunities (NUDOS) NCS2022_046.

## Authors' Contributions

Author roles were classified using the Contributor Role Taxonomy [CRediT](https://credit.niso.org/) as follows: Juan Carlos Castillo: Conceptualization, Supervision, Methodology, Writing - review and editing; Andreas Laffert: Conceptualization, Methodology, Formal analysis, Software, Visualization, Writing - review and editing; Kevin Carrasco: Methodology, Formal analysis, Writing - review and editing. Julio Iturra: Conceptualization, Writing - review and editing.

## Conflict of Interest

The authors declare no conflict of interest.

## Funding

This research received funding from ANID/FONDECYT grant 1210847 and the Center for Conflict and Social Cohesion Studies (COES) through ANID/FONDAP grant 1513009.

## Data Avaibility Statement

The original data presented in the study are openly available in <https://dataverse.harvard.edu/api/access/datafile/10797987>

Access to the data, code, and analysis presented in the study are openly available in <https://github.com/justicia-distributiva/market-justice-panel-elsoc>.

[^*]: Corresponding author (juancastillov@uchile.cl - [jc-castillo.com](https://juancarloscastillo.github.io/jc-castillo/) - Ignacio Carrera Pinto 1045 Ñuñoa, Santiago de Chile).