### 2025-05-22 cover_letter.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
]{article}

\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
  \setmainfont[]{Times New Roman}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=2cm]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Perceptions of Inequality and Meritocracy: Their Interplay in Shaping Preferences for Market Justice in Chile (2016-2023)},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Perceptions of Inequality and Meritocracy: Their Interplay in
Shaping Preferences for Market Justice in Chile (2016-2023)}

\date{}
\begin{document}
\maketitle

\setstretch{1.15}
\section{Authors Identification}\label{authors-identification}

Juan Carlos Castillo, Department of Sociology, Universidad de Chile,
Santiago de Chile\footnote{Corresponding author (juancastillov@uchile.cl
  -
  \href{https://juancarloscastillo.github.io/jc-castillo/}{jc-castillo.com}
  - Ignacio Carrera Pinto 1045 Ñuñoa, Santiago de Chile).}.\\
ORCID: 0000-0003-1265-7854

Andreas Laffert, Institute of Sociology, Pontificia Universidad Católica
de Chile.\\
ORCID: 0000-0002-9008-2454

Kevin Carrasco, Center for Social Conflict and Cohesion Studies.\\
ORCID: 0000-0002-9348-8361

Julio Iturra, Bremen International Graduate School of Social Sciences,
Bremen, Germany. ORCID: 0000-0003-1993-3479

\section{Exclusive Submission
Statement}\label{exclusive-submission-statement}

The authors confirm that this article has been submitted exclusively to
this journal.

\section{Acknowledgments}\label{acknowledgments}

The authors recognize the support of the National Research and
Development Agency (ANID) through the grant from the National Fund for
Scientific and Technological Development (FONDECYT) N° 1210847; from the
Center for Conflict and Social Cohesion Studies (COES) grant Financing
Fund for Research Centers in Priority Areas (FONDAP) N° 1523A0005; and
Millenium Nucleus on Digital Inequalities and Opportunities (NUDOS)
NCS2022\_046.

\section{Authors' Contributions}\label{authors-contributions}

Author roles were classified using the Contributor Role Taxonomy
\href{https://credit.niso.org/}{CRediT} as follows: Juan Carlos
Castillo: Conceptualization, Supervision, Methodology, Writing - review
and editing; Andreas Laffert: Conceptualization, Methodology, Formal
analysis, Software, Visualization, Writing - review and editing; Kevin
Carrasco: Methodology, Formal analysis, Writing - review and editing.
Julio Iturra: Conceptualization, Writing - review and editing.

\section{Conflict of Interest}\label{conflict-of-interest}

The authors declare no conflict of interest.

\section{Funding}\label{funding}

This research received funding from ANID/FONDECYT grant 1210847 and the
Center for Conflict and Social Cohesion Studies (COES) through
ANID/FONDAP grant 1513009.

\section{Data Avaibility Statement}\label{data-avaibility-statement}

The original data presented in the study are openly available in
\url{https://dataverse.harvard.edu/api/access/datafile/10797987}

Access to the data, code, and analysis presented in the study are openly
available in
\url{https://github.com/justicia-distributiva/market-justice-panel-elsoc}.

\end{document}