### 2025-07-11 06-conclusion.qmd

---
author: "Equipo EDUMER"
bibliography: "input/bib/merit-edjust.bib"
csl: "input/bib/apa6.csl"
---

# Conclusions

This study examined the complex interplay between perceptions of economic inequality, meritocratic perceptions, and attitudes toward market justice in Chile from 2016 to 2023, drawing on longitudinal data from the ELSOC survey. By exploring how subjective assessments and social contexts influence support for redistribution and market-based resource allocation, the research offers different elements that contribute to the understanding of the normative foundations underpinning social justice attitudes in a highly unequal, neoliberal environment.

The findings confirm that higher perceptions of economic inequality are associated with less support for market justice attitudes, this is, the belief that it is fine that those with higher income have better social services such as education, pensions and health. Meritocratic perceptions strongly reinforce support for market justice, serving as moral justifications for structural inequalities. Additionally, perceptions of inequality moderate the influence of meritocratic perceptions, such that as the awareness of disparities are attenuated (i.e. less negative) the role of meritocratic perceptions on existing market-based allocations become stronger. Notably, the social mobilizations of 2019 and subsequent periods seem to have contributed to shifting attitudes, fostering greater skepticism of systemic fairness and reducing the legitimizing power of meritocratic narratives.

This research advances the extant literature by integrating subjective perceptions with social and political contexts to explain attitudes toward economic inequality and distributional justice. While previous studies primarily focused on objective measures or individual characteristics, this work emphasizes the dynamic and interactional nature of perceptions and beliefs over time. It highlights the importance of socio-political upheavals in reshaping normative attitudes, underscoring the role of collective action in challenging entrenched narratives of meritocracy and fairness. The longitudinal approach provides a deeper temporal perspective on how societal events could influence individual perceptions and preferences.

Despite its contributions, the study has several limitations. First, the reliance on self-reported perceptions and attitudes may be subject to social desirability bias. Second, the focus on Chile, while offering valuable insights into a specific context, limits the generalizability of results to other countries with different institutional and cultural backgrounds. Third, the observational nature of the data prevents causal inferences, and unobserved confounding factors may influence the observed relationships. Finally, the potential influence of media, political discourse, and peer effects on perceptions was not directly assessed. Future research could explore the causal mechanisms linking perceptions and attitudes, possibly through experimental or mixed-method designs. Comparative studies across different national contexts would help determine the universality or specificity of these dynamics. Additionally, investigating the role of media, political communication, and education in shaping perceptions of inequality and meritocracy would deepen understanding of the normative foundations of social justice attitudes. Finally, examining how these perceptions influence behavioral outcomes, such as political participation or support for social movements, would provide valuable insights into the pathways from beliefs to collective action and policy change.