### 2025-07-28 resubmission/blinded-paper-resubmit.tex
Update varios ajustes

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
]{article}

\usepackage{amsmath,amssymb}
\usepackage{setspace}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
  \setmainfont[]{Times New Roman}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=2cm]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage[noblocks]{authblk}
\renewcommand*{\Authsep}{, }
\renewcommand*{\Authand}{, }
\renewcommand*{\Authands}{, }
\renewcommand\Affilfont{\small}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Perceptions of Inequality and Meritocracy: Their Interplay in Shaping Preferences for Market Justice in Chile (2016-2023)},
  pdfauthor={Equipo EDUMER},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Perceptions of Inequality and Meritocracy: Their Interplay in
Shaping Preferences for Market Justice in Chile (2016-2023)}

\date{}
\begin{document}
\maketitle
\begin{abstract}
Several countries have experienced a shift toward the privatization and
commodification of public goods, welfare policies, and social services.
In Latin America, Chile stands out as a paradigmatic case where this
trend has led to the extensive marketization of essential services. From
a moral economy perspective, the extent to which individuals consider it
fair for access to such services to depend on market criteria has been
conceptualized as market justice preferences. This study investigates
the relationship between perceptions of economic inequality,
meritocratic beliefs, and market justice preferences in Chile between
2016 and 2023. Using six waves of panel data from the Chilean
Longitudinal Social Survey -- ELSOC (\(N_{observations}\) = 8,643;
\(N_{individuals}\) = 1,687), the analysis examines how subjective
assessments of inequality shape attitudes toward the role of merit in
access to key social services such as healthcare, education, and
pensions. The findings show that greater perceived inequality is
associated with lower market justice preferences. However, individuals
who believe that effort is rewarded are more likely to legitimize
existing disparities. In contrast, the perception that talent is
rewarded shows a negative effect on market justice preferences; an
effect that intensifies as perceived inequality increases over time. The
study also considers the influence of major social movements during this
period, which appear to have reshaped public discourse on justice and
fairness. These findings contribute to a deeper understanding of how
distributive beliefs evolve in contexts marked by persistent inequality
and entrenched neoliberal frameworks \newline \textbf{Keywords}:
Economic inequality, meritocracy, market justice, Chile, public
preferences, inequality perception
\end{abstract}

\setstretch{1.15}
\section{Introduction}\label{introduction}

Since 1980, economic inequality and wealth concentration have
dramatically increased worldwide, becoming one of the main challenges
for the social sciences. Globally, in 2021, less than 50\% of the world
population owned only 2\% of the wealth, while the richest 10\%
concentrated 76\%. Futhermore, the wealthiest 1\% captured nearly 38\%
of total assets (\citeproc{ref-chancel_world_2022}{Chancel et al.,
2022}). This context of economic disparity has sparked renewed interest
in studying not only the objective aspects of inequality, such as income
and access to resources, but also its subjective dimensions, including
perceptions, beliefs, and associated attitudes
(\citeproc{ref-janmaat_subjective_2013}{Janmaat, 2013}). Such dimensions
are deemed as key to understand the political delegitimation that
economic inequality can elicit
(\citeproc{ref-castillo_perception_2022}{Castillo et al., 2022}).
Therefore, understanding how people perceive inequality is crucial, as
these perceptions can influence how societies comprehend and justify (or
not) inequality in the distribution of goods and services, as well as
their capacity to funnel political discontent and eventually opposition
to the current economic distribution.

The justification of economic inequality has a long tradition in
sociology, with seminal works such as the ones from Kluegel
(\citeproc{ref-kluegel_social_1995a}{1995};
\citeproc{ref-kluegel_beliefs_1986}{1986}) and Lane
(\citeproc{ref-lane_political_1962}{1962};
\citeproc{ref-lane_market_1986}{1986}). Lane's main aim was to ``help
explain the tenacious hold of market capitalism on the public mind''
(\citeproc{ref-lane_market_1986}{1986, p. 384}), proposing the
distinction between market and political justice. Whereas political
justice refer to government actions aligned with justice principles such
equality or need, market justice involves the allocation linked to
earned desserts, particularly through individual contributions based on
as effort and ability. Sociological survey research in this area has
considered questions related to the extent to which it is considered
just to get better access to social services (such as pensions, health
and education) based on payment capacity
(\citeproc{ref-busemeyer_skills_2014}{Busemeyer, 2014};
\citeproc{ref-castillo_socialization_2024}{Castillo et al., 2024};
\citeproc{ref-immergut_it_2020}{Immergut \& Schneider, 2020};
\citeproc{ref-koos_moral_2019}{Koos \& Sachweh, 2019};
\citeproc{ref-lindh_public_2015}{Lindh, 2015};
\citeproc{ref-lindh_bringing_2023}{Lindh \& McCall, 2023}). Although
these services could be considered basic rights of citizens, evidence to
date suggests that a significant portion of the population justifies
inequality in these areas (\citeproc{ref-koos_moral_2019}{Koos \&
Sachweh, 2019}) and that such preferences vary according to objective
socioeconomic factors and the contextual characteristics of countries
(\citeproc{ref-busemeyer_skills_2014}{Busemeyer, 2014};
\citeproc{ref-lindh_public_2015}{Lindh, 2015}). In the present paper, we
argue that there are two other relevant factors to take into account
when analyzing differences between individuals and groups to this
regard: the perception of economic inequality and the perception of
meritocracy.

The perception of economic inequality can be understood as an
individual's subjective assessment of how resources are allocated among
members of a given society (\citeproc{ref-akyelken_urban_2020}{Akyelken,
2020}). Regardless of their measurement, various studies have shown that
this perception often underestimates the gap between the rich and the
poor, revealing inconsistencies between objective and subjective aspects
of inequality (\citeproc{ref-trump_income_2018}{Trump, 2018}). Moreover,
there is consistent evidence that perceived inequality, even under
control of objective inequality, could have implications for attitudes
toward the distribution of goods and services
(\citeproc{ref-castillo_perception_2022}{Castillo et al., 2022};
\citeproc{ref-schroder_income_2017}{Schröder, 2017}), redistributive
preferences, justification of inequality, and legitimacy of the economic
system (\citeproc{ref-castillo_perception_2022}{Castillo et al., 2022};
\citeproc{ref-garcia-sanchez_attitudes_2020}{García-Sánchez et al.,
2020}, \citeproc{ref-garcia-sanchez_perceptions_2018}{2018}). From this
perspective, in this paper we argue that perceptions of economic
inequality may play a relevant aspect for understanding market justice
preferences, as a lower perceived inequality could reinforce the idea
that the differences in access to social services could be justified ---
a link that becomes difficult to hold when high inequality is perceived.

Along with perception of inequality, a second focus of this paper is the
perception of meritocracy. Meritocracy posits that inequality can be
legitimized through distributive criteria such as effort and talent
(\citeproc{ref-davis_principles_2001}{Davis \& Moore, 2001};
\citeproc{ref-young_rise_1962}{Young, 1962}). Previous studies have
demonstrated that people with stronger meritocratic beliefs tend to
perceive less inequality as they attribute economic differences to
individual achievements (\citeproc{ref-mijs_paradox_2021}{Mijs, 2021};
\citeproc{ref-wilson_role_2003}{Wilson, 2003}) and also to justify more
inequality as these beliefs are associated with attitudes that
legitimize status differences
(\citeproc{ref-batruch_belief_2023}{Batruch et al., 2023}). In a context
as Chile, where the distribution of goods and services is predominantly
governed by market logics strongly introduced during the military
dictatorship (1973-1989) (\citeproc{ref-boccardo_30_2020}{Boccardo,
2020}; \citeproc{ref-madariaga_three_2020}{Madariaga, 2020}),
meritocratic perceptions have the potential to affect the acceptance of
social inequalities. Therefore, incorporating the perception of
meritocracy in this literature allows for an understanding of how
attitudes toward inequality are shaped not only by the perception of
economic gaps, but also by beliefs about how these gaps are justified.
Furthermore, these beliefs are often consolidated from an early age,
reinforced by institutions promoting values such as effort and
individual skills to climb socially
(\citeproc{ref-castillo_socialization_2024}{Castillo et al., 2024};
\citeproc{ref-reynolds_perceptions_2014}{Reynolds \& Xian, 2014}).

At this point we can propose the following research questions: To what
extent the preferences for market justice are associated with
perceptions of inequality and meritocracy? How stable are these
associations (within subjects) along time? The primary objective of this
study is to analyze the interplay between perceptions of inequality and
meritocracy and their joint influence on preferences for market justice
along time in Chile using the panel survey ELSOC (Social Chilean
Longitudinal Study, 2016-2023). The chilean case appears as a particular
case for studying market justice preferences given their story of high
inequality and strong free-market policies introduced in the
dictatorship period (\citeproc{ref-flores_top_2020}{Flores et al.,
2020}; \citeproc{ref-madariaga_three_2020}{Madariaga, 2020}).
Additionally, this analysis seeks to explore how political and social
contingencies---such as the 2019 and 2022 social movements---might have
moderated these relationships by prompting more critical reflection on
the commodification of social services
(\citeproc{ref-somma_no_2021}{Somma et al., 2021}). Recognizing the
temporal dimension in shaping market justice preferences is essential,
given that historical and contextual factors that could challenge or
reaffirm the way in which individuals persceive inequality and
meritocracy, and consequantely --- following our argument --- the
preferences for market justice.

\section{Theoretical views on market justice, inequality perception, and
meritocracy}\label{theoretical-views-on-market-justice-inequality-perception-and-meritocracy}

\subsection{The justification of market
inequality}\label{the-justification-of-market-inequality}

Conceptually, \emph{market justice} has been discussed in the literature
as a normative principle that legitimates the distribution of economic
rewards based on individual merit. It is possible to trace the concept
to the understanding of Lane (\citeproc{ref-lane_market_1986}{1986}),
who makes a contrast between market justice and political justice. The
author defines market justice as a system of ``earned deserts'', whereby
individuals are seen as deserving of a determined distributive outcome
due to their effort and skills. In contrast, political justice
emphasizes principles of equality and need, which are often represented
by the welfare state action through social policies. An important remark
is that the principles of market justice rely on the assumption that
markets are neutral and self-regulating arenas, where individuals are
treated fairly because they face the same formal rules of engagement
(\citeproc{ref-lane_market_1986}{Robert E. Lane, 1986}). Consequently,
the legitimacy of market justice stems from the assumption that
inequalities are not only inevitable but fair---so long as the rules are
transparent and opportunities are open. In this way, market justice
provides a moral justification for inequality by framing it as a
necessary outcome of individual responsibility
(\citeproc{ref-kluegel_legitimation_1999}{Kluegel et al., 1999};
\citeproc{ref-lindh_public_2015}{Lindh, 2015}).

Empirical studies have shown different strategies for the study of
market justice preferences. A common approach in the literature is to
gauge attitudes toward the legitimacy of inequality in specific domains,
especially when linked to income differences. This can be traced to the
seminal work of Kluegel and Smith
(\citeproc{ref-kluegel_beliefs_1981}{1981}) who assessed the normative
foundations that explain public support for economic inequality. Over
time, this approach has been extended beyond income to include other
market-mediated outcomes, such as education, healthcare, and/or
pensions. For example, Von Dem Knesebeck et al.
(\citeproc{ref-vondemknesebeck_are_2016}{2016}) and Immergut and
Schneider (\citeproc{ref-immergut_it_2020}{2020}) examine whether
citizens consider it fair that individuals with higher incomes can
access better healthcare, while Lee and Stacey
(\citeproc{ref-lee_fairness_2023}{2023}) apply a similar method in the
context of education in Australia. These studies usually rely on a
survey item asking respondents to evaluate the fairness of income-based
access to welfare services, allowing for comparing justice perceptions
across different contexts. Similarly, comparative studies have also
considered these indicators to study the support for market-based
distribution in welfare systems (\citeproc{ref-lindh_public_2015}{Lindh,
2015}; \citeproc{ref-svallfors_political_2007}{Svallfors, 2007}). More
recently, Castillo et al.
(\citeproc{ref-castillo_socialization_2024}{2024}) introduced a
single-item composite measure of market justice to assess student
attitudes toward income-based access to education, healthcare, and
pensions in Chile. These empirical strategies all aim to capture the
extent to which individuals accept inequality when framed as a
reflection of market outcomes.

The study of market justice preferences has increasingly focused on how
they are shaped by individuals' socioeconomic position, normative
beliefs, and the institutional context in which they are embedded.
Across the literature, there is empirical evidence suggesting that
individuals in more advantaged socioeconomic positions---those with
higher occupational class, income, and education---are more likely to
support market justice principles (\citeproc{ref-koos_moral_2019}{Koos
\& Sachweh, 2019}; \citeproc{ref-lindh_public_2015}{Lindh, 2015};
\citeproc{ref-svallfors_political_2007}{Svallfors, 2007}). This tendency
reflects not only material self-interest but also a broader moral
economy, in which winners of the market system internalize
justifications for the status quo. At the same time, political ideology
also plays a role ---such as economically conservative values --- where
more right-wing individuals show higher support for meritocracy and more
skepticism towards redistribution. This is particularly salient in
countries with more restricted public provision of social services. For
example, Castillo et al.
(\citeproc{ref-castillo_socialization_2024}{2024}) in Chile and, Lee and
Stacey (\citeproc{ref-lee_fairness_2023}{2023}) in Australia show that
right-wing individuals are more supportive of market-based distribution
of welfare. Beyond individual characteristics, country-level
institutions also play a central role. In liberal welfare regimes like
those of the United States or the United Kingdom, market justice
preferences are more widespread, while in coordinated or
social-democratic regimes---such as Sweden or Germany---citizens are
generally more critical of market-based inequalities
(\citeproc{ref-immergut_it_2020}{Immergut \& Schneider, 2020};
\citeproc{ref-lindh_public_2015}{Lindh, 2015}). These findings suggest
that market justice is also shaped contextually, which is represented by
the development of welfare institutions, in line with the
policy-feedbacks literature
(\citeproc{ref-campbell_policy_2012}{Campbell, 2012};
\citeproc{ref-pierson_when_1993}{Pierson, 1993}).

\subsection{Perception of inequality}\label{perception-of-inequality}

Perceptions of inequality have been associated with a range of
distributive attitudes (\citeproc{ref-castillo_perception_2022}{Castillo
et al., 2022}; \citeproc{ref-garcia-sanchez_vicious_2019}{García-Sánchez
et al., 2019}; \citeproc{ref-mccall_exposure_2017}{McCall et al., 2017};
\citeproc{ref-trump_income_2018}{Trump, 2018}). Research indicates that
perceptions of inequality influence redistributive preferences to a
greater extent than objective or actual economic inequality
(\citeproc{ref-gimpelson_misperceiving_2018}{Gimpelson \& Treisman,
2018}). Others have shown that the perception of inequality, through the
perceived wage gap, is related to the justification of greater
inequality (\citeproc{ref-Castillo2011}{Castillo, 2011}). In that sense,
lower perceived inequality can reinforce support for market-based
distributive arrangements by suggesting that the system is fair and that
outcomes reflect effort and ability (\citeproc{ref-kuhn_eye_2011}{Kuhn,
2011}). In contrast, when inequality is perceived as excessive or
structurally determined, individuals could question the legitimacy of
market justice and become more supportive of redistributive policies
(\citeproc{ref-garcia-sanchez_vicious_2019}{García-Sánchez et al.,
2019}).

Perceptions of inequality refer to individuals' subjective evaluations
of the extent, causes, and consequences of income and wealth disparities
(\citeproc{ref-akyelken_urban_2020}{Akyelken, 2020}). Unlike objective
measures such as the Gini index, perceived inequality captures how
individuals make sense of distributive hierarchies in their everyday
lives, shaped by reference groups, social comparisons, and information
environments (\citeproc{ref-garcia-castro_perceiving_2020}{García-Castro
et al., 2020}; \citeproc{ref-gimpelson_misperceiving_2018}{Gimpelson \&
Treisman, 2018}; \citeproc{ref-mijs_stratified_2016}{Mijs, 2016}).
Scholars have proposed multiple dimensions of perceived inequality,
including its magnitude (how significant are the gaps), vertical
structure (between which groups), the trend over time (increasing or
decreasing), and legitimacy (whether it is just or not)
(\citeproc{ref-engelhardt_what_2018}{Engelhardt \& Wagener, 2018};
\citeproc{ref-garcia-sanchez_vicious_2019}{García-Sánchez et al.,
2019}). These dimensions encompass both cognitive and normative aspects
of perceptions of inequality and can vary across societies and social
groups, depending on exposure, ideology, and personal experience
(\citeproc{ref-castillo_perception_2022}{Castillo et al., 2022};
\citeproc{ref-garcia-sanchez_perceptions_2018}{García-Sánchez et al.,
2018}).

Among the different approaches to measuring perceived inequality, one of
the most widely used is the estimation of wage gaps between occupational
extremes, such as between a CEO and a manual worker. This type of item
provides a concrete frame that respondents can relate to more easily
than abstract questions about national income distribution
(\citeproc{ref-castillo_percepcion_2019}{Castillo et al., 2012};
\citeproc{ref-easterbrook_social_2021}{Easterbrook, 2021};
\citeproc{ref-willis_legitimacy_2015}{Willis et al., 2015}). While it
enables the estimation of inequality using simple heuristics, this
method is not without challenges. For instance, people often lack
reliable knowledge about the earnings of those at the top of the income
ladder, which leads to high variability in responses and the use of
biased mental shortcuts (\citeproc{ref-knell_perceptions_2020}{Knell \&
Stix, 2020}). Despite this, perceived wage gaps are strong predictors of
political attitudes
(\citeproc{ref-garcia-sanchez_perceptions_2018}{García-Sánchez et al.,
2018}; \citeproc{ref-pedersen_attitudes_2019}{Pedersen \& Mutz, 2019}),
making them a valuable tool for understanding public responses to
economic disparities.

Another relevant critique of this approach lies in its conflation of
different psychological constructs. Many surveys assess perceived
inequality through Likert-type items that ask respondents to agree or
disagree with statements, such as ``income differences are too large,''
which captures general concern or discomfort rather than a specific
perception (\citeproc{ref-Castillo2011}{Castillo, 2011};
\citeproc{ref-garcia-sanchez_vicious_2019}{García-Sánchez et al.,
2019}). These items mix cognitive estimations with affective
evaluations, complicating the interpretation of what respondents
perceive versus what they morally reject. As a result, the conceptual
clarity between perceived inequality and inequality aversion remains
blurred in many empirical studies. To address this limitation, recent
work has emphasized the need to distinguish between absolute and
comparative measures, as well as between ideal and actual estimates of
economic gaps (\citeproc{ref-auspurg_why_2017}{Auspurg et al., 2017};
\citeproc{ref-garcia-sanchez_creencias_2022}{García-Sánchez \& De
Carvalho, 2022}).

\subsection{Perception of meritocracy}\label{perception-of-meritocracy}

Meritocracy constitutes a central ideological framework for legitimizing
different types of social inequality, for instance through market
justice beliefs. Rooted in the belief that rewards and positions should
be allocated based on individual effort and talent, meritocracy operates
as a normative ideal and a descriptive belief about how society
functions. As initially conceptualized by Michael Young
(\citeproc{ref-young_rise_1962}{1962}), the term was meant to critique a
system in which merit-based stratification becomes a new form of
inequality. However, over time, meritocracy has been widely supported in
many societies as a fair and desirable principle of distribution,
particularly within liberal democracies and market-oriented economies
(\citeproc{ref-mijs_paradox_2021}{Mijs, 2021};
\citeproc{ref-sandel_tyranny_2020}{Sandel, 2020}). From a sociological
perspective, the belief in meritocracy is more than a cognitive
assessment; it reflects a moral lens through which individuals interpret
inequality. People who believe that success results from hard work and
talent are more likely to view social and economic disparities as
legitimate (\citeproc{ref-batruch_belief_2023}{Batruch et al., 2023};
\citeproc{ref-castillo_percepcion_2019}{Castillo et al., 2012}).
Conversely, if they see outcomes as driven by luck, social origin, or
systemic barriers, inequality is more likely to be perceived as unjust.
This distinction becomes crucial in societies with persistent structural
inequality, where public narratives often emphasize personal
responsibility and merit while overlooking entrenched disadvantages.

We adopt a multidimensional perspective on meritocracy, distinguishing
between two key dimensions: effort-based and talent-based perceptions.
This distinction is essential, as it captures different pathways through
which individuals justify inequality. Effort-based meritocracy
emphasizes hard work and perseverance as the basis for social rewards,
aligning closely with cultural narratives of personal responsibility. A
talent-based meritocracy, by contrast, emphasizes intelligence and
innate abilities, which are often perceived as less malleable and more
unequally distributed. Both dimensions have been shown to correlate with
acceptance of inequality, but they may carry distinct implications for
how inequality is justified in specific domains
(\citeproc{ref-castillo_multidimensional_2023}{Castillo et al., 2023}).
The relevance of this distinction is supported by recent studies, which
show that individuals respond differently to these dimensions. For
instance, perceptions that effort is rewarded in society are more
strongly associated with positive evaluations of fairness and acceptance
of unequal outcomes (\citeproc{ref-batruch_belief_2023}{Batruch et al.,
2023}). This may be because effort is seen as a controllable and morally
virtuous trait, whereas talent is often perceived as a natural
advantage. Consequently, effort-based meritocracy is likely more potent
in legitimizing inequality, particularly in neoliberal contexts.

These dimensions of meritocracy reflect how respondents perceive
society's distributive logic, regardless of whether they endorse
meritocratic principles. This distinction aligns with recent findings
indicating that individuals distinguish between how merit is perceived
in society and how it should ideally operate, which in turn shapes their
preferences for redistribution and justice
(\citeproc{ref-tejero-peregrina_perceived_2025}{Tejero-Peregrina et al.,
2025}). Meritocratic beliefs serve as symbolic justifications for
unequal outcomes, particularly when access is stratified by income or
social opportunity. Prior studies have shown that individuals who
perceive higher levels of meritocracy tend to express stronger support
for unequal distributions that reflect market outcomes
(\citeproc{ref-castillo_percepcion_2019}{Castillo et al., 2012};
\citeproc{ref-castillo_socialization_2024}{Castillo et al., 2024}).

In addition to influencing individual attitudes toward inequality,
meritocratic perceptions can contribute to social division and the
stigmatization of disadvantaged groups. Recent research has demonstrated
that exposure to meritocratic narratives can reinforce the belief that
poverty results from individual failings rather than systemic
conditions, reducing support for redistributive measures and increasing
the stigmatization of the poor (\citeproc{ref-hoyt_mindsets_2023}{Hoyt
et al., 2023}). This reinforces negative stereotypes and reduces empathy
toward individuals from lower socioeconomic backgrounds. Moreover,
Busemeyer et al. (\citeproc{ref-busemeyer_positive_2021}{2021}) argues
that meritocratic narratives can serve as feedback mechanisms that shape
public opinion and well-being by framing individuals' understanding of
welfare outcomes as deserved or undeserved within existing institutional
structures. This psychological mechanism highlights the normative power
of meritocracy in stabilizing unequal systems by shaping political
attitudes and personal perceptions of success and failure.

\subsection{The Chilean context}\label{the-chilean-context}

Chile offers an interesting case for examining how public attitudes
toward the distribution of social services evolve amid declining poverty
and persistently high income inequality within a residual welfare
framework (\citeproc{ref-ferre_welfare_2023}{Ferre, 2023}). Following
the neoliberal reforms of the 1980s, the Chilean welfare state has
become increasingly reliant on private provision, with essential
services often commodified and largely reserved for those with
sufficient purchasing power (\citeproc{ref-boccardo_30_2020}{Boccardo,
2020}; \citeproc{ref-madariaga_three_2020}{Madariaga, 2020}). This model
has favored the affluent, while lower-income groups are left dependent
on limited and underfunded public alternatives. Research on social
stratification characterizes Chile as a society where mobility into
intermediate classes is feasible, but access to elite strata remains
largely restricted
(\citeproc{ref-lopez-roldan_comparative_2021}{López-Roldán \& Fachelli,
2021}; \citeproc{ref-torche_intergenerational_2014a}{Torche, 2014}).
While some indicators---such as education and occupation---suggest a
relatively fluid class structure, income mobility remains constrained
(\citeproc{ref-espinoza_movilidad_2014}{Espinoza \& Núñez, 2014}).

Despite steady economic expansion, Chile continues to rank among the
most unequal nations in the OECD, marked by a high Gini coefficient and
substantial wealth concentration at the top
(\citeproc{ref-flores_top_2020}{Flores et al., 2020}). In this context,
Chile experienced a wave of mass protests from October 2019 to March
2020. Initially led by high school students, the demonstrations quickly
gained momentum as broader segments of society mobilized around demands
for more equality of opportunities, particularly regarding precarious
access to welfare services. The political system interpreted this unrest
as a call to reform existing political institutions, with a special
focus on equalizing opportunities and providing more egalitarian access
to welfare by reducing reliance on out-of-pocket mechanisms
(\citeproc{ref-somma_no_2021}{Somma et al., 2021}).

Concerning public opinion, the ELSOC survey shows that Chile has a low
level of agreement with the idea that it is fair for those with higher
incomes to buy better welfare in the areas of education, healthcare, and
pensions, with average support of 8.5\% between 2016 and 2023. At the
same time, around 55\% and 47\% of the public between 2016 and 2023
\emph{disagreed} that effort and talent are rewarded accordingly. This
coincides with 90\% of citizens \emph{agreeing} that income differences
in Chile are \emph{too large}
(\citeproc{ref-castillo_changes_2025}{Castillo et al., 2025}).

In sum, Chile represents an illustrative case of a particular form of
\emph{moral economy} (\citeproc{ref-koos_moral_2019}{Koos \& Sachweh,
2019}), in which national socioeconomic characteristics, institutional
arrangements, and public attitudes coexist. This suggests that the
country combines persistently high levels of economic and social
inequality with a strong institutionalization of market-based mechanisms
within the welfare system
(\citeproc{ref-madariaga_three_2020}{Madariaga, 2020}). At the same
time, perceptions of inequality are high, and there is widespread public
criticism of the commodification of social services
(\citeproc{ref-somma_no_2021}{Somma et al., 2021}). For the above
reasons, we argue that the Chilean case represents a relevant context to
study the role of perceived income inequality and meritocracy in shaping
public preferences regarding the commodification of welfare services,
particularly from the perspective of individual change.

\subsection{This study}\label{this-study}

Building upon the previous literature, this study proposes that
attitudes toward market justice are shaped by a dynamic interplay
between individuals' perceptions of economic inequality and their
perceptions of meritocracy. Specifically, we argue that both perceptions
independently and interactively influence the extent to which
individuals endorse market-based distributions of social goods and
services in Chile.

First, consistent with previous findings about redistributive
preferences (\citeproc{ref-castillo_perception_2022}{Castillo et al.,
2022}), we expect that a higher perception of economic inequality will
be associated with lower market justice preferences. When individuals
perceive smaller income gaps, they are more likely to view market
mechanisms as fair and legitimate, reinforcing the acceptance of
outcomes based on competition and ability to pay. Conversely, a
heightened perception of inequality may erode confidence in market
fairness, weakening support for market-based distribution. This
relationship is particularly relevant in the context of Chile, where the
neoliberal economic model has been a dominant force in shaping public
attitudes toward inequality and justice
(\citeproc{ref-canalesceron_sujeto_2021}{Canales Cerón et al., 2021};
\citeproc{ref-mac-clure_justicia_2024}{Mac-Clure et al., 2024}).

Second, higher perceived meritocracy is expected to be positively
associated with market justice preferences. Individuals who believe that
effort and talent primarily determine success are more likely to justify
unequal outcomes and endorse the notion that markets allocate resources
fairly according to individual merit
(\citeproc{ref-castillo_socialization_2024}{Castillo et al., 2024}).
This aligns with the idea that meritocratic perceptions serve as a moral
framework that legitimizes market-based inequalities, as individuals
perceive the system as just when they believe that rewards are based on
individual merit. This is particularly relevant in contexts where
neoliberal ideologies dominate, as they often emphasize individual
responsibility and competition as the basis for social order.

Third, we propose that perceptions of meritocracy and perceptions of
economic inequality interact in shaping market justice attitudes.
Specifically, we argue that the legitimizing effect of perceived
meritocracy on market justice preferences is moderated by perceived
economic inequality: when perceived inequality is low, the positive
association between meritocratic perceptions and market justice
preferences will be stronger. However, when perceived inequality is
high, this association will weaken, as greater awareness of large
economic gaps may challenge the view that outcomes are purely
merit-based.

Additionally, this study examines whether these relationships vary over
time, particularly in light of broader social and political
transformations. We are especially interested in whether changes in
individuals' perceptions of inequality and meritocracy are associated
with shifts in their market justice attitudes across different time
points. The primary focus of this study is thus to scrutinize how
individual-level changes in perceptions relate to changes in preferences
over time. While we acknowledge the relevance of contextual events ---
such as the 2019 political outburst or the subsequent constitutional
process --- for public opinion and political orientation
(\citeproc{ref-disipavlic_justification_2025}{Disi Pavlic et al.,
2025}), identifying causal effects of such events (e.g., mass protest
exposure) on individual attitudes lies outside the scope of this
research. Instead, we focus on how intra-individual trajectories in
perceived inequality and merit shape evolving attitudes toward the
commodification of social services.

Based on these arguments, we propose the following hypotheses:

\(H1_{a} \; (\text{Between-person}):\) Individuals who perceive higher
levels of economic inequality will report lower support for market
justice preferences, on average, compared to those who perceive lower
levels of inequality.

\(H1_{b} \; (\text{Within-person}):\) Increases in an individual's
perception of economic inequality over time will be associated with
decreases in their support for market justice preferences.

\(H2_{a} \; (\text{Between-person}):\) Individuals who endorse stronger
meritocratic perceptions will report higher support for market justice
preferences, on average, compared to those with weaker meritocratic
perceptions

\(H2_{b} \; (\text{Within-person}):\) Increases in an individual's
meritocratic perceptions over time will be associated with increases in
their support for market justice preferences.

\(H3_{a} \; (\text{Between-person}):\) The positive association between
meritocratic perceptions and market justice preferences will be weaker
among individuals who perceive higher levels of economic inequality,
compared to those who perceive lower inequality.

\(H3_{b} \; (\text{Within-person}):\) The positive association between
meritocratic perceptions and market justice preferences will be weaker
at times when individuals perceive higher levels of inequality, relative
to times when they perceive lower inequality.

\(H4_{a} \; (\text{Growth moderation – Inequality}):\) The longitudinal
association between perceived economic inequality and market justice
preferences will strengthen over time, both within and between
individuals.

\(H4_{b} \; (\text{Growth moderation – Meritocracy}):\) The longitudinal
association between meritocratic perceptions and market justice
preferences will weaken over time, both within and between individuals.

\section{Data, Variables and Methods}\label{data-variables-and-methods}

\subsection{Data}\label{data}

This study draws on data from the Chilean Longitudinal Social Survey
(ELSOC), a nationally representative panel study of the urban adult
population in Chile, conducted annually between 2016 and 2023. Designed
to examine individuals' attitudes, emotions, and behaviors regarding
social conflict and cohesion, ELSOC employs a probabilistic, stratified,
clustered, and multistage sampling design covering both major urban
centers and smaller cities. The sampling frame was proportionally
stratified into six categories of urban population size (e.g., large and
small cities), followed by the random selection of households within
1,067 city blocks. The target population includes men and women aged 18
to 75 who are habitual residents of private dwellings.

The survey has been conducted every year since 2016, except in 2020,
when it was suspended due to the COVID-19 pandemic. This study uses six
waves: 2016, 2017, 2018, 2019, 2022, and 2023. The 2021 wave was
excluded because a reduced version of the questionnaire omitted key
variables of interest. Between waves 1 and 6, panel attrition reached
40\%, resulting in a final two-level sample comprising N = 8,643
observations nested within N = 1,687 individuals. Longitudinal weights
are applied to adjust for both the sampling design and potential biases
from systematic non-response. Further details on sampling, attrition,
and weighting procedures are available at
https://coes.cl/encuesta-panel/, and the dataset is publicly accessible
at https://dataverse.harvard.edu/dataverse/elsoc.

\subsection{Variables}\label{variables}

\textbf{Market justice preferences}: the dependent variable in this
study is preferences for market justice. This construct is
operationalized through three items that capture how strongly
individuals justify conditioning access to core services---healthcare,
pensions, and education--- basen on individual income. Specifically, the
justification of inequality in healthcare is assessed by the question:
``Is it fair in Chile that people with higher incomes can access better
healthcare than people with lower incomes?'' The same question is posed
for pensions and education. In all cases, respondents indicate their
level of agreement on a five-point Likert scale ranging from 1
(``strongly disagree'') to 5 (``strongly agree''). Additionally, we
include a composite measure of ``market justice preferences'',
calculated as the average of these three items (\(\alpha\) = 0.84). This
index ranges from 1 to 5, with higher values indicating stronger
preferences for market justice (see Table~\ref{tbl-summary1}).

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3367}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3367}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2143}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1122}}@{}}
\caption{Dependent variables for the first wave
(2016)}\label{tbl-summary1}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Label
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Stats / Values
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Freqs (\% of Valid)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Valid
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Label
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Stats / Values
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Freqs (\% of Valid)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Valid
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Health distributive justice &
\begin{minipage}[t]{\linewidth}\raggedright
1. Strongly desagree\\
2. Desagree\\
3. Neither agree nor desagre\\
4. Agree\\
5. Strongly agree\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
558 (37.2\%)\\
729 (48.6\%)\\
63 ( 4.2\%)\\
133 ( 8.9\%)\\
18 ( 1.2\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
1501\\
(100.0\%)\strut
\end{minipage} \\
Pension distributive justice &
\begin{minipage}[t]{\linewidth}\raggedright
1. Strongly desagree\\
2. Desagree\\
3. Neither agree nor desagre\\
4. Agree\\
5. Strongly agree\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
426 (28.4\%)\\
718 (47.8\%)\\
108 ( 7.2\%)\\
226 (15.1\%)\\
23 ( 1.5\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
1501\\
(100.0\%)\strut
\end{minipage} \\
Education distributive justice &
\begin{minipage}[t]{\linewidth}\raggedright
1. Strongly desagree\\
2. Desagree\\
3. Neither agree nor desagre\\
4. Agree\\
5. Strongly agree\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
521 (34.7\%)\\
783 (52.2\%)\\
73 ( 4.9\%)\\
113 ( 7.5\%)\\
11 ( 0.7\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
1501\\
(100.0\%)\strut
\end{minipage} \\
Market justice preferences & \begin{minipage}[t]{\linewidth}\raggedright
Mean (sd) : 2 (0.8)\\
min \textless{} med \textless{} max:\\
1 \textless{} 2 \textless{} 5\\
IQR (CV) : 0.7 (0.4)\strut
\end{minipage} & 12 distinct values &
\begin{minipage}[t]{\linewidth}\raggedright
1501\\
(100.0\%)\strut
\end{minipage} \\
\end{longtable}

\textbf{Perception of economic inequality}: this variable is measured
through the perceived wage gap
(\citeproc{ref-castillo_cual_2009}{Castillo, 2009};
\citeproc{ref-gijsberts_thelegitimation_1999}{Gijsberts, 1999};
\citeproc{ref-hadler_why_2005}{Hadler, 2005}). This measure is derived
from the salary gap between the perceived salaries of jobs at opposite
ends of the occupational hierarchy. Specifically, it relies on the
division between the perceived salary of a large-company president and
that of an unskilled worker (\citeproc{ref-Castillo2011}{Castillo,
2011}). Higher values of this term indicate a greater perception of
economic inequality between occupations located at the extremes of the
status continuum. This measure includes a logarithmic term in order to
adjust income magnitudes (usually fewer cases with high income):

\[
\text{perceived wage gap} = \log_{10}\left(\frac{\text{perceived salary of a large-company president}}{\text{perceived salary of an unskilled worker}}\right)
\]

\textbf{Perception of Meritocracy}: this variable is operationalized
through two components, namely effort and talent
(\citeproc{ref-young_rise_1962}{Young, 1962}). The item used to gauge
effort is: ``In Chile, people are rewarded for their efforts,'' while
the item for talent is: ``In Chile, people are rewarded for their
intelligence and skills''. In both cases, respondents indicate their
level of agreement on a five-point Likert scale, ranging from 1
(``strongly disagree'') to 5 (``strongly agree'').

Table 2 shows the independent variables used, their response categories
and their frequencies.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3774}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3113}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2075}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1038}}@{}}
\caption{Independent variables ELSOC survey (descriptives for first wave
2016)}\label{tbl-summary2}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Label
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Stats / Values
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Freqs (\% of Valid)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Valid
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Label
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Stats / Values
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Freqs (\% of Valid)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Valid
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Inequality gap perception & \begin{minipage}[t]{\linewidth}\raggedright
Mean (sd) : 3.7 (1.1)\\
min \textless{} med \textless{} max:\\
0.4 \textless{} 3.7 \textless{} 6.9\\
IQR (CV) : 1.6 (0.3)\strut
\end{minipage} & 296 distinct values &
\begin{minipage}[t]{\linewidth}\raggedright
1501\\
(100.0\%)\strut
\end{minipage} \\
People are rewarded for their efforts &
\begin{minipage}[t]{\linewidth}\raggedright
1. Strongly desagree\\
2. Desagree\\
3. Neither agree nor desagre\\
4. Agree\\
5. Strongly agree\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
169 (11.3\%)\\
693 (46.2\%)\\
263 (17.5\%)\\
328 (21.9\%)\\
48 ( 3.2\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
1501\\
(100.0\%)\strut
\end{minipage} \\
People are rewarded for their intelligence &
\begin{minipage}[t]{\linewidth}\raggedright
1. Strongly desagree\\
2. Desagree\\
3. Neither agree nor desagre\\
4. Agree\\
5. Strongly agree\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
134 ( 8.9\%)\\
617 (41.1\%)\\
294 (19.6\%)\\
401 (26.7\%)\\
55 ( 3.7\%)\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
1501\\
(100.0\%)\strut
\end{minipage} \\
\end{longtable}

\emph{Controls}

Sociodemographic and attitudinal variables are included to control for
potential composition effects in the population. In terms of
sociodemographic characteristics, we incorporate per capita household
income quintile, educational level (1=Less than Universitary,
2=Universitary), age (in ranges), and sex (1=Male, 2=Female), which have
been previously shown to influence market justice preferences
significantly (\citeproc{ref-castillo_socialization_2024}{Castillo et
al., 2024}; \citeproc{ref-lindh_public_2015}{Lindh, 2015}). Regarding
attitudinal variables, we include political identification (1=Left,
2=Center, 3=Right, 4=No identification) and subjective social status
(measured through a scale from 1 to 10) as they may affect the
relationship between market justice preferences, perceptions of
inequality, and meritocracy
(\citeproc{ref-schneider_poverty_2015}{Schneider \& Castillo, 2015}).
Descriptive statistics for the control variables can be found in the
supplementary material.

\subsection{Methods}\label{methods}

Given the data's hierarchical structure, in which observations are
nested in survey waves, we employ longitudinal multilevel linear models
(\citeproc{ref-singer_applied_2009}{Singer \& Willett, 2009}). In a
panel-data framework, within-person effects capture how shifts in
individual-level variables across waves are associated with variations
in market justice preferences. By contrast, between-person effects focus
on differences among individuals, explaining how long-term (or average)
values relate to overall levels of market justice preferences.

To estimate within-person effects, we use group-mean centering, where
each respondent functions as the ``group'' (i.e., observations nested
within persons). Meanwhile, the between-person effects are derived from
each individual's average on these variables, calculated across the
waves of panel data.

All the analyses were conducted using R software and the \emph{lme4}
package (\citeproc{ref-bates_fitting_2015}{Bates et al., 2015}).

\section{Results}\label{results}

\subsection{Descriptive statistics}\label{descriptive-statistics}

Figure~\ref{fig-alluvial} shows the annual frequencies of market justice
preferences for healthcare, pensions, and education from 2016 to 2023.
Each year presents stacked percentage frequencies, and the flows between
them reflect opinion changes among the same individuals from one year to
the next, given that we are using panel data. For instance, of the
40.8\% who strongly disagreed with justifying inequality in healthcare
in 2019, around 24.3\% maintained that position in 2022, while the
remaining 16.5\% shifted toward other response categories---primarily
moving into disagreement rather than strong disagreement. Overall, more
than half of the respondents exhibit a high level of disagreement
(disagree + strongly disagree) with inequality in these three social
service areas over time. Despite this general pattern, recent waves show
a slight decrease in disagreement and a corresponding rise in support
for market-justice inequality. Specifically, in healthcare and
education, although disagreement remains substantial, agreement (agree +
strongly agree) increased from 7.4\% and 7.2\% in 2019 to 13.1\% and
14.2\% in 2023, respectively. This shift is most evident in pensions,
where the combined agree/strongly agree category grew by about 10
percentage points, from 16.9\% in 2016 to 28\% in 2023.

\begin{figure}[H]

\caption{\label{fig-alluvial}Change in the justification of inequality
in healthcare, pensions and education over time (2016-2023)}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{blinded-paper-resubmit_files/figure-pdf/fig-alluvial-1.pdf}

}

\end{figure}%

Regarding the main dependent and independent variables of this study,
Figure~\ref{fig-meanchange} depicts their standardized average values
across survey waves. The results show a notable upward trend in market
justice preferences, particularly in the most recent waves. Perceived
economic inequality exhibited the highest standardized mean in 2016;
however, this variable shows a general downward trajectory over time,
with a temporary increase in 2019, possibly reflecting the effects of
the social uprising that year. Interestingly, while perceptions of
inequality declined in the latest waves (2022--2023), market justice
preferences continued to rise. Meritocratic beliefs, in contrast, remain
relatively stable overall. Nevertheless, perceptions that individuals
are rewarded based on talent tend to be slightly higher than those based
on effort. Both meritocracy-related measures follow a similar temporal
pattern: an increase from 2016 to 2018, a decline between 2019 and
2022---potentially associated with the social unrest and the
consequences of the COVID-19 pandemic---and a subsequent rise in 2023,
returning to levels comparable to those observed in 2016.

\begin{figure}[H]

\caption{\label{fig-meanchange}Change in the standarized mean of market
justice preferences, economic inequality perception, and meritocracy
(2016-2023)}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{blinded-paper-resubmit_files/figure-pdf/fig-meanchange-1.pdf}

}

\end{figure}%

\subsection{Multilevel models}\label{multilevel-models}

Table~\ref{tbl-modelos} presents the results of the multilevel models
estimated for market justice preferences, examining both individuals
(within) and group-level (between) effects. The intraclass correlation
(\citeproc{ref-hoxMultilevelAnalysisTechniques2017a}{Hox et al., 2017})
from the empty model (see Supplementary Material), which decomposes the
variance of market justice preferences, is 0.31, indicating that
approximately 31\% of the variation is attributable to differences
between individuals. Complementary, 69\% of the variation corresponds to
within-individual differences over time.

\begin{table}

\caption{\label{tbl-modelos}Longitudinal multilevel models for market
justice preferences}

\centering{

\caption{}
\begin{center}
\scalebox{0.8}{
\begin{tabular}{l c c c c c c c}
\hline
 & Model 1 & Model 2 & Model 3 & Model 4 & Model 5 & Model 6 & Model 7 \\
\hline
Intercept                          & $1.938^{***}$  & $1.948^{***}$  & $1.965^{***}$  & $1.967^{***}$  & $1.974^{***}$  & $1.186^{***}$  & $1.250^{***}$  \\
                                   & $(0.023)$      & $(0.037)$      & $(0.037)$      & $(0.037)$      & $(0.087)$      & $(0.124)$      & $(0.144)$      \\
Wave (Ref.= 2016)                  &                &                &                &                &                &                &                \\
                                   &                &                &                &                &                &                &                \\
\quad Wave 2017                    & $-0.183^{***}$ &                &                &                &                &                &                \\
                                   & $(0.025)$      &                &                &                &                &                &                \\
\quad Wave 2018                    & $-0.009$       &                &                &                &                &                &                \\
                                   & $(0.025)$      &                &                &                &                &                &                \\
\quad Wave 2019                    & $-0.009$       &                &                &                &                &                &                \\
                                   & $(0.025)$      &                &                &                &                &                &                \\
\quad Wave 2022                    & $0.300^{***}$  &                &                &                &                &                &                \\
                                   & $(0.025)$      &                &                &                &                &                &                \\
\quad Wave 2023                    & $0.320^{***}$  &                &                &                &                &                &                \\
                                   & $(0.025)$      &                &                &                &                &                &                \\
Wave                               &                & $-0.088^{***}$ & $-0.095^{***}$ & $-0.096^{***}$ & $-0.096^{***}$ & $-0.096^{***}$ & $-0.096^{***}$ \\
                                   &                & $(0.020)$      & $(0.020)$      & $(0.020)$      & $(0.020)$      & $(0.020)$      & $(0.020)$      \\
Wave$^2$                           &                & $0.024^{***}$  & $0.024^{***}$  & $0.025^{***}$  & $0.025^{***}$  & $0.025^{***}$  & $0.025^{***}$  \\
                                   &                & $(0.003)$      & $(0.003)$      & $(0.003)$      & $(0.003)$      & $(0.003)$      & $(0.003)$      \\
Perception inequality (WE)         &                &                & $-0.027^{**}$  & $-0.025^{**}$  & $-0.025^{**}$  & $-0.025^{**}$  & $-0.025^{**}$  \\
                                   &                &                & $(0.009)$      & $(0.009)$      & $(0.009)$      & $(0.009)$      & $(0.009)$      \\
Merit: Effort (WE)                 &                &                &                & $0.070^{***}$  & $0.070^{***}$  & $0.070^{***}$  & $0.070^{***}$  \\
                                   &                &                &                & $(0.011)$      & $(0.011)$      & $(0.011)$      & $(0.011)$      \\
Merit: Talent (WE)                 &                &                &                & $-0.027^{*}$   & $-0.027^{*}$   & $-0.027^{*}$   & $-0.027^{*}$   \\
                                   &                &                &                & $(0.011)$      & $(0.011)$      & $(0.011)$      & $(0.011)$      \\
Perception inequality (BE)         &                &                &                &                & $-0.002$       & $0.043$        & $0.008$        \\
                                   &                &                &                &                & $(0.023)$      & $(0.023)$      & $(0.024)$      \\
Merit: Effort (BE)                 &                &                &                &                &                & $0.206^{***}$  & $0.191^{***}$  \\
                                   &                &                &                &                &                & $(0.041)$      & $(0.040)$      \\
Merit: Talent (BE)                 &                &                &                &                &                & $0.036$        & $0.021$        \\
                                   &                &                &                &                &                & $(0.040)$      & $(0.040)$      \\
\hline
Controls                           & No             & No             & No             & No             & No             & No             & Yes            \\
BIC                                & $32146.711$    & $31406.958$    & $31414.699$    & $31404.308$    & $31419.062$    & $31366.239$    & $31473.850$    \\
Numb. obs.                         & $8643$         & $8643$         & $8643$         & $8643$         & $8643$         & $8643$         & $8643$         \\
Num. groups: individuals           & $1687$         & $1687$         & $1687$         & $1687$         & $1687$         & $1687$         & $1687$         \\
Var: individuals (Intercept)       & $0.205$        & $0.370$        & $0.366$        & $0.363$        & $0.364$        & $0.336$        & $0.326$        \\
Var: Residual                      & $0.416$        & $0.345$        & $0.345$        & $0.343$        & $0.343$        & $0.343$        & $0.343$        \\
Var: individuals, wave             & $$             & $0.022$        & $0.021$        & $0.021$        & $0.021$        & $0.021$        & $0.021$        \\
Cov: individuals (Intercept), wave & $$             & $-0.061$       & $-0.060$       & $-0.059$       & $-0.059$       & $-0.058$       & $-0.059$       \\
\hline
\multicolumn{8}{l}{\scriptsize{Note: Cells contain regression coefficients with standard errors in parentheses. $^{***}p<0.001$; $^{**}p<0.01$; $^{*}p<0.05$.}}
\end{tabular}
}
\label{table:coefficients}
\end{center}

}

\end{table}%

According to Model 1, which includes the survey waves to capture
intertemporal variations in the dependent variable, there is a decrease
in 2017 (\(\beta\) = -0.183, \(p\) \textless{} .001) relative to 2016,
and similarly in 2018 (\(\beta\) = -0.009, \(p\) \textgreater{} .05) and
2019 (\(\beta\) = -0.009, \(p\) \textgreater{} .05), although the latter
effects are not statistically significant. In contrast, in the more
recent waves of 2022 and 2023, there is a statistically significant
increase in market justice preferences (\(\beta\) = 0.300, \(p\)
\textless{} .001; \(\beta\) = 0.320, \(p\) \textless{} .001), suggesting
a non-linear effect. To model this trajectory over time, Model 2
incorporates time (survey waves) as a continuous variable, along with
its quadratic term, representing the non-linear association initially
observed in Model 1. While the linear term (survey wave) shows a
negative association, reflecting an overall decline in market
preferences over time, the positive quadratic term indicates a reversal
of this pattern in the final measurement points.

Models 3 and 4 incorporate the within-group effects (WE) of the primary
independent variables, capturing how individual changes in these
variables over time shape the dependent variable. The results in Model 3
suggest that the within effect of perceived economic inequality is
negative and statistically significant (\(p\) \textless{} .001).
Specifically, each one-point increase in an individual's perception of
economic inequality between waves is associated with a 0.027 point
decrease in market justice preferences. Model 4 shows that meritocratic
perceptions operate in distinct directions. An upward shift in the
perception that effort is rewarded exerts a positive within effect
(\(\beta\) = 0.070, \(p\) \textless{} .001), while a parallel increase
in the perception that intelligence and ability are rewarded is likewise
associated with lower market-justice preferences (\(\beta\) = -0.027,
\(p\) \textless{} .05). Taken together, these results suggest that
people who increasingly perceive meritocracy based on effort tend to
have stronger preferences for market justice, and that the opposite is
true for those who increasingly perceive meritocracy based on talent.

When examining the between-group effects (BE) in Model 5 and 6, which
capture differences between individuals in the average of the main
variables, a similar pattern emerges. Individuals who perceive higher
levels of economic inequality tend to prefer less market justice
(\(\beta\) = -0.002, \(p\) \textgreater{} .05). However, this effects is
no longer statistically significant. In Model 6, the meritocratic
perception that effort is rewarded is positively associated with market
justice preferences (\(\beta\) = 0.206, \(p\) \textless{} .001), whereas
the perception that talent is rewarded shows a positive but
non-significant coefficient (\(\beta\) = 0.036).

Model 7 adds the control variables. The within- and between-effects of
the principal predictors retain both their direction and statistical
significance, confirming the robustness of the associations (see
Supplementary Material for effects of control variables).

\begin{table}

\caption{\label{tbl-interactions1}Interactions for meritocracy,
perceived economic inequality and market justice preferences}

\centering{

\caption{}
\begin{center}
\scalebox{0.8}{
\begin{tabular}{l c c c c}
\hline
 & Model 8 & Model 9 & Model 10 & Model 11 \\
\hline
Intercept                                                     & $1.328^{***}$ & $1.346^{***}$ & $1.956^{***}$  & $2.281^{***}$  \\
                                                              & $(0.143)$     & $(0.143)$     & $(0.350)$      & $(0.369)$      \\
Perception inequality (WE)                                    & $-0.037^{*}$  & $-0.036^{*}$  & $-0.040^{***}$ & $-0.040^{***}$ \\
                                                              & $(0.015)$     & $(0.015)$     & $(0.009)$      & $(0.009)$      \\
Merit: Effort (WE)                                            & $0.075^{***}$ & $0.085^{***}$ & $0.081^{***}$  & $0.081^{***}$  \\
                                                              & $(0.018)$     & $(0.012)$     & $(0.011)$      & $(0.011)$      \\
Merit: Talent (WE)                                            & $-0.021$      & $-0.017$      & $-0.026^{*}$   & $-0.026^{*}$   \\
                                                              & $(0.011)$     & $(0.016)$     & $(0.011)$      & $(0.011)$      \\
Perception inequality (BE)                                    & $-0.003$      & $-0.003$      & $-0.176$       & $-0.268^{**}$  \\
                                                              & $(0.023)$     & $(0.023)$     & $(0.096)$      & $(0.101)$      \\
Merit: Effort (BE)                                            & $0.180^{***}$ & $0.183^{***}$ & $-0.045$       & $0.191^{***}$  \\
                                                              & $(0.040)$     & $(0.040)$     & $(0.127)$      & $(0.041)$      \\
Merit: Talent (BE)                                            & $0.022$       & $0.018$       & $0.017$        & $-0.320^{*}$   \\
                                                              & $(0.039)$     & $(0.039)$     & $(0.041)$      & $(0.129)$      \\
Merit: Effort (WE) x Perception inequality (WE)               & $-0.014$      &               &                &                \\
                                                              & $(0.013)$     &               &                &                \\
Merit: Talent (WE) x Perception inequality (WE)               &               & $-0.032^{*}$  &                &                \\
                                                              &               & $(0.013)$     &                &                \\
Merit: Effort (BE) x Perception inequality (BE)               &               &               & $0.070$        &                \\
                                                              &               &               & $(0.036)$      &                \\
Merit: Talent (BE) x Perception inequality (BE)               &               &               &                & $0.099^{**}$   \\
                                                              &               &               &                & $(0.036)$      \\
\hline
Controls                                                      & Yes           & Yes           & Yes            & Yes            \\
BIC                                                           & $31201.714$   & $31254.484$   & $32281.698$    & $32277.809$    \\
Numb. obs.                                                    & $8643$        & $8643$        & $8643$         & $8643$         \\
Num. groups: individuals                                      & $1687$        & $1687$        & $1687$         & $1687$         \\
Var: individuals (Intercept)                                  & $0.178$       & $0.178$       & $0.170$        & $0.169$        \\
Var: individuals, merit effort cwc                            & $0.127$       & $$            & $$             & $$             \\
Var: individuals, perception inequality cwc                   & $0.090$       & $0.088$       & $$             & $$             \\
Cov: individuals (Intercept), merit effort cwc                & $-0.009$      & $$            & $$             & $$             \\
Cov: individuals (Intercept), perception inequality cwc       & $-0.015$      & $-0.013$      & $$             & $$             \\
Cov: individuals, merit effort cwc, perception inequality cwc & $0.003$       & $$            & $$             & $$             \\
Var: Residuals                                                & $0.296$       & $0.303$       & $0.417$        & $0.417$        \\
Var: individuals, merit talent cwc                            & $$            & $0.097$       & $$             & $$             \\
Cov: individuals (Intercept), merit talent cwc                & $$            & $-0.011$      & $$             & $$             \\
Cov: individuals, merit talent cwc, perception inequality cwc & $$            & $-0.004$      & $$             & $$             \\
\hline
\multicolumn{5}{l}{\scriptsize{Note: Cells contain regression coefficients with standard errors in parentheses. $^{***}p<0.001$; $^{**}p<0.01$; $^{*}p<0.05$. CWC = centered within group.}}
\end{tabular}
}
\label{table:coefficients}
\end{center}

}

\end{table}%

\begin{figure}[H]

\caption{\label{fig-interact}Predicted values of market justice
preferences by perceptions of meritocraticy and economic inequality}

\centering{

\includegraphics[width=1\textwidth,height=\textheight]{blinded-paper-resubmit_files/figure-pdf/fig-interact-1.pdf}

}

\end{figure}%

Table~\ref{tbl-interactions1} examines whether perceived economic
inequality moderates the effect of meritocratic perceptions on market
justice preferences. Contrary to our expectations, the interaction terms
in the within-person specification of Model 9 indicate that the negative
effect of talent-based meritocratic perceptions becomes stronger as
perceived inequality increases (\(\beta\) = -0.032, \(p\) \textless{}
.05). In contrast, in the between-person specification (Model 11), the
interaction term is positive (\(\beta\) = 0.099, \(p\) \textless{} .01),
suggesting that perceptions of economic inequality significantly shape
the influence of meritocratic perceptions on support for market-based
allocation of social services. Specifically, these positive effects
indicate that, as perceptions of inequality increase, the association
between talent-based meritocratic perceptions and market justice
preferences shifts in a positive direction.

The effects of effort-based meritocratic perceptions are not
statistically significant in either the within- or between-person
specifications (Models 8 and 10). However, the direction of the
coefficients is negative in the former and positive in the latter. As
shown in Figure~\ref{fig-interact}, Panel A, the between-person effect
of effort-based meritocratic perceptions on market justice preferences
strengthens as perceived inequality increases across individuals,
whereas this effect weakens among those who perceive lower levels of
inequality. Panel B reveals a similar pattern for the between-person
effect of talent-based meritocratic perceptions, although the effect is
substantially weaker at low levels of perceived inequality.

Substantively, these findings suggest that the legitimizing function of
meritocratic perceptions is amplified in contexts perceived as highly
unequal. In such contexts, meritocratic narratives may reinforce
individualistic understandings of distributive justice and intensify
support for market-based access to essential social services.

\begin{table}

\caption{\label{tbl-interactions2}Interactions for meritocracy,
perceived economic inequality and market justice preferences}

\centering{

\caption{}
\begin{center}
\scalebox{0.7}{
\begin{tabular}{l c c c c c c}
\hline
 & Model 12 & Model 13 & Model 14 & Model 15 & Model 16 & Model 17 \\
\hline
Intercept                                                & $1.263^{***}$ & $1.265^{***}$ & $1.246^{***}$ & $1.128^{***}$ & $1.099^{***}$ & $1.090^{***}$ \\
                                                         & $(0.143)$     & $(0.141)$     & $(0.142)$     & $(0.176)$     & $(0.169)$     & $(0.170)$     \\
Wave x Perception inequality (WE)                        & $0.003$       &               &               &               &               &               \\
                                                         & $(0.006)$     &               &               &               &               &               \\
Wave x Merit: effort (WE)                                &               & $0.012$       &               &               &               &               \\
                                                         &               & $(0.006)$     &               &               &               &               \\
Wave x Merit: talent (WE)                                &               &               & $0.007$       &               &               &               \\
                                                         &               &               & $(0.006)$     &               &               &               \\
Wave x Perception inequality (BE)                        &               &               &               & $-0.011$      &               &               \\
                                                         &               &               &               & $(0.009)$     &               &               \\
Wave x Merit: effort (BE)                                &               &               &               &               & $-0.020$      &               \\
                                                         &               &               &               &               & $(0.011)$     &               \\
Wave x Merit: talent (BE)                                &               &               &               &               &               & $-0.016$      \\
                                                         &               &               &               &               &               & $(0.010)$     \\
\hline
Controls                                                 & Yes           & Yes           & Yes           & Yes           & Yes           & Yes           \\
BIC                                                      & $31010.853$   & $30888.601$   & $30785.269$   & $31515.978$   & $31493.536$   & $31501.405$   \\
Numb. obs.                                               & $8643$        & $8643$        & $8643$        & $8643$        & $8643$        & $8643$        \\
Num. groups: individuals                                 & $1687$        & $1687$        & $1687$        & $1687$        & $1687$        & $1687$        \\
Var: individuals (Intercept)                             & $0.388$       & $0.334$       & $0.379$       & $0.274$       & $0.532$       & $0.372$       \\
Var: individuals, perception inequality cwc              & $0.100$       & $$            & $$            & $$            & $$            & $$            \\
Var: individuals, wave                                   & $0.025$       & $0.023$       & $0.024$       & $0.021$       & $0.021$       & $0.021$       \\
Cov: individuals (Intercept), perception inequality cwc  & $-0.081$      & $$            & $$            & $$            & $$            & $$            \\
Cov: individuals (Intercept), wave                       & $-0.074$      & $-0.062$      & $-0.072$      & $-0.054$      & $-0.056$      & $-0.049$      \\
Cov: individuals, perception inequality cwc, wave        & $0.021$       & $$            & $$            & $$            & $$            & $$            \\
Var: Residuals                                           & $0.291$       & $0.280$       & $0.275$       & $0.343$       & $0.342$       & $0.343$       \\
Var: individuals, merit effort cwc                       & $$            & $0.122$       & $$            & $$            & $$            & $$            \\
Cov: individuals (Intercept), merit effort cwc           & $$            & $-0.009$      & $$            & $$            & $$            & $$            \\
Cov: individuals, merit effort cwc, wave                 & $$            & $0.000$       & $$            & $$            & $$            & $$            \\
Var: individuals, merit talent cwc                       & $$            & $$            & $0.112$       & $$            & $$            & $$            \\
Cov: individuals (Intercept), merit talent cwc           & $$            & $$            & $0.003$       & $$            & $$            & $$            \\
Cov: individuals, merit talent cwc, wave                 & $$            & $$            & $-0.003$      & $$            & $$            & $$            \\
Var: individuals, perception inequality mean             & $$            & $$            & $$            & $0.000$       & $$            & $$            \\
Cov: individuals (Intercept), perception inequality mean & $$            & $$            & $$            & $0.007$       & $$            & $$            \\
Cov: individuals, perception inequality mean, wave       & $$            & $$            & $$            & $-0.001$      & $$            & $$            \\
Var: individuals, merit effort mean                      & $$            & $$            & $$            & $$            & $0.051$       & $$            \\
Cov: individuals (Intercept), merit effort mean          & $$            & $$            & $$            & $$            & $-0.110$      & $$            \\
Cov: individuals, merit effort mean, wave                & $$            & $$            & $$            & $$            & $-0.001$      & $$            \\
Var: individuals, merit talent mean                      & $$            & $$            & $$            & $$            & $$            & $0.031$       \\
Cov: individuals (Intercept), merit talent mean          & $$            & $$            & $$            & $$            & $$            & $-0.053$      \\
Cov: individuals, merit talent mean, wave                & $$            & $$            & $$            & $$            & $$            & $-0.003$      \\
\hline
\multicolumn{7}{l}{\scriptsize{Note: Cells contain regression coefficients with standard errors in parentheses. $^{***}p<0.001$; $^{**}p<0.01$; $^{*}p<0.05$. CWC = centered within group.}}
\end{tabular}
}
\label{table:coefficients}
\end{center}

}

\end{table}%

With respect to the temporal dynamics of the key predictors,
Table~\ref{tbl-interactions2} shows that time has no statistically
significant effects on either the within- or between-person effects of
perceived inequality and meritocratic perceptions. The within-person
interaction effects are positive, suggesting that in later waves, the
influence of perceived inequality and meritocratic beliefs on market
justice preferences tends to increase. Conversely, in the between-person
specification, the interaction effects are negative, indicating that
over time, the impact of perceived inequality and meritocratic beliefs
on support for market-based justice tends to diminish. However, these
trends are not statistically significant, and thus no definitive
conclusions can be drawn.

\section{Discussion}\label{discussion}

This study tests several key hypotheses concerning the relationships
between perceptions of economic inequality, meritocratic perceptions,
and attitudes toward market justice in Chile. It posits that higher
perceived inequality and stronger meritocratic perceptions both
independently contribute to greater support for market-based
distributions of social goods. Furthermore, the study examines how
perceived economic inequality moderates the influence of meritocratic
perceptions on market justice preferences. Additionally, it considers
how major social and political mobilizations, such as those occurring
around 2019 and subsequent years, may have weakened these relationships,
reflecting increased critical perspectives on systemic inequalities and
market mechanisms. The subsequent discussion interprets the empirical
findings in light of these hypotheses, considering their implications
for understanding attitudes toward inequality and social justice in a
context marked by profound social change.

The first hypothesis suggested that perceptions of economic inequality
were strongly linked to attitudes toward market justice. The findings
demonstrated that individuals perceiving higher income disparities
showed stronger support for the distribution of social services based on
individual income, aligning with theories emphasizing that awareness of
inequality could fuel a critical view of such policies
(\citeproc{ref-Castillo2012a_justice}{Castillo, 2012};
\citeproc{ref-mijs_paradox_2021}{Mijs, 2021}). This relationship
reflected a moral economy where perceptions of systemic unfairness
threatened the legitimacy of existing distributions, prompting calls for
greater equity. Interestingly, the negative within-effects observed over
time implied that as perceptions of inequality became more salient, some
individuals experienced a decline in support for market justice,
possibly due to increasing distrust in market mechanisms or
disillusionment with the fairness of the system. This indicated that
while perception heightened awareness, prolonged exposure or social
mobilization could also foster skepticism about whether redistribution
effectively addressed inequality or simply perpetuated systemic flaws.

Regarding the second hypothesis, the results reinforced that
meritocratic perceptions ---particularly those emphasizing individual
effort and talent--- were associated with stronger support for
market-based distribution systems. Individuals who viewed success as
primarily earned through effort were more inclined to justify existing
inequalities, seeing them as outcomes of individual merit rather than
systemic injustice. This aligned with research indicating that
meritocratic narratives served as moral justifications that reinforced
societal stratification (\citeproc{ref-Castillo2012a_justice}{Castillo,
2012}; \citeproc{ref-hoyt_mindsets_2023}{Hoyt et al., 2023}). These
perceptions functioned symbolically to legitimize structural
disparities, reducing support for redistributive measures by framing
inequality as fair and deserved. The findings suggested that these
perceptions bolstered the social hierarchy by fostering acceptance of
inequality as a reflection of individual virtue rather than structural
failure, which was especially pertinent in neoliberal contexts like
Chile's, where market logic heavily influenced social attitudes
(\citeproc{ref-canalesceron_sujeto_2021}{Canales Cerón et al., 2021}).

The third hypothesis concerned the moderating role of perceptions of
inequality on the relationship between meritocracy and market justice
attitudes. The analysis indicated that the positive association between
meritocratic perceptions and support for market-based distribution
tended to weaken when perceptions of economic inequality were
particularly high. Although this moderation is not significant for
within-subject effects (i.e.~over time), when taking into account
between-subjects differences we found a positive effect. This appears as
contrary to our initial hypothesis, as it would mean that effect of
meritocratic perceptions becomes stronger as perception of inequality
increases. Nevertheless, we need to be cautious about this
interpretation, as the main effect of inequality perception on market
justice is negative. Therefore, as perceived inequality become less
negative (i.e.~more positive), or closer to zero, then the meritocratic
perception has an stronger positive relationship on market justice
preferences. This could mean that meritocratic perceptions play an
stronger role in justification of market justice inequalities for those
who are less capable or less prone to perceive inequality.

Finally, the fourth hypothesis addressed the impact of recent social
movements, particularly from 2019 onwards, on the relationship between
meritocratic perceptions, perceptions of inequality, and attitudes
toward market justice. The results suggested that these movements
contributed to a more critical stance toward the fairness of existing
distributional mechanisms. Post-mobilization, the positive relationship
between meritocratic perceptions and support for inequality appeared
attenuated, indicating that collective action and public discourse
during this period fostered questioning of the legitimacy of systemic
inequalities justified solely through meritocratic narratives. These
social mobilizations challenged dominant narratives, emphasizing
structural barriers and systemic injustices, which in turn weakened the
association between individual effort and deservingness. This is aligned
with literature emphasizing that social protests act as catalysts for
transforming societal attitudes by disrupting normative beliefs and
encouraging more critical evaluations of institutional legitimacy
(\citeproc{ref-tilly_contentious_2015}{Tilly \& Tarrow, 2015}). The
findings underscored the importance of socio-political context in
shaping and reshaping attitudes toward inequality and justice,
highlighting how collective action could temporarily or permanently
alter the normative foundations that sustained support for market
justice.

\section{Conclusions}\label{conclusions}

This study examined the complex interplay between perceptions of
economic inequality, meritocratic perceptions, and attitudes toward
market justice in Chile from 2016 to 2023, drawing on longitudinal data
from the ELSOC survey. By exploring how subjective assessments and
social contexts influence support for redistribution and market-based
resource allocation, the research offers different elements that
contribute to the understanding of the normative foundations
underpinning social justice attitudes in a highly unequal, neoliberal
environment.

The findings confirm that higher perceptions of economic inequality are
associated with less support for market justice attitudes, this is, the
belief that it is fine that those with higher income have better social
services such as education, pensions and health. Meritocratic
perceptions strongly reinforce support for market justice, serving as
moral justifications for structural inequalities. Additionally,
perceptions of inequality moderate the influence of meritocratic
perceptions, such that as the awareness of disparities are attenuated
(i.e.~less negative) the role of meritocratic perceptions on existing
market-based allocations become stronger. Notably, the social
mobilizations of 2019 and subsequent periods seem to have contributed to
shifting attitudes, fostering greater skepticism of systemic fairness
and reducing the legitimizing power of meritocratic narratives.

This research advances the extant literature by integrating subjective
perceptions with social and political contexts to explain attitudes
toward economic inequality and distributional justice. While previous
studies primarily focused on objective measures or individual
characteristics, this work emphasizes the dynamic and interactional
nature of perceptions and beliefs over time. It highlights the
importance of socio-political upheavals in reshaping normative
attitudes, underscoring the role of collective action in challenging
entrenched narratives of meritocracy and fairness. The longitudinal
approach provides a deeper temporal perspective on how societal events
could influence individual perceptions and preferences.

Despite its contributions, the study has several limitations. First, the
reliance on self-reported perceptions and attitudes may be subject to
social desirability bias. Second, the focus on Chile, while offering
valuable insights into a specific context, limits the generalizability
of results to other countries with different institutional and cultural
backgrounds. Third, the observational nature of the data prevents causal
inferences, and unobserved confounding factors may influence the
observed relationships. Finally, the potential influence of media,
political discourse, and peer effects on perceptions was not directly
assessed. Future research could explore the causal mechanisms linking
perceptions and attitudes, possibly through experimental or mixed-method
designs. Comparative studies across different national contexts would
help determine the universality or specificity of these dynamics.
Additionally, investigating the role of media, political communication,
and education in shaping perceptions of inequality and meritocracy would
deepen understanding of the normative foundations of social justice
attitudes. Finally, examining how these perceptions influence behavioral
outcomes, such as political participation or support for social
movements, would provide valuable insights into the pathways from
beliefs to collective action and policy change.

\section{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-akyelken_urban_2020}
Akyelken, N. (2020). Urban conceptions of economic inequalities.
\emph{Regional Studies}, \emph{54}(6), 863--872.
\url{https://doi.org/10.1080/00343404.2020.1732902}

\bibitem[\citeproctext]{ref-auspurg_why_2017}
Auspurg, K., Hinz, T., \& Sauer, C. (2017). Why {Should Women Get Less}?
{Evidence} on the {Gender Pay Gap} from {Multifactorial Survey
Experiments}. \emph{American Sociological Review}, \emph{82}(1),
179--210. \url{https://doi.org/10.1177/0003122416683393}

\bibitem[\citeproctext]{ref-bates_fitting_2015}
Bates, D., Mächler, M., Bolker, B., \& Walker, S. (2015). Fitting linear
mixed-effects models using {lme4}. \emph{Journal of Statistical
Software}, \emph{67}(1), 1--48.
\url{https://doi.org/10.18637/jss.v067.i01}

\bibitem[\citeproctext]{ref-batruch_belief_2023}
Batruch, A., Jetten, J., Van De Werfhorst, H., Darnon, C., \& Butera, F.
(2023). Belief in {School Meritocracy} and the {Legitimization} of
{Social} and {Income Inequality}. \emph{Social Psychological and
Personality Science}, \emph{14}(5), 621--635.
\url{https://doi.org/10.1177/19485506221111017}

\bibitem[\citeproctext]{ref-boccardo_30_2020}
Boccardo, G. (2020). \emph{30 a{ñ}os de privatizaciones en {Chile}: {Lo}
que la pandemia revel{ó}} (Nodo XXI). Santiago.

\bibitem[\citeproctext]{ref-busemeyer_skills_2014}
Busemeyer, M. (2014). \emph{Skills and {Inequality}: {Partisan Politics}
and the {Political Economy} of {Education Reforms} in {Western Welfare
States}}. Cambridge University Press.

\bibitem[\citeproctext]{ref-busemeyer_positive_2021}
Busemeyer, M., Abrassart, A., \& Nezi, R. (2021). Beyond {Positive} and
{Negative}: {New Perspectives} on {Feedback Effects} in {Public Opinion}
on the {Welfare State}. \emph{British Journal of Political Science},
\emph{51}(1), 137--162. \url{https://doi.org/10.1017/S0007123418000534}

\bibitem[\citeproctext]{ref-campbell_policy_2012}
Campbell, A. L. (2012). Policy {Makes Mass Politics}. \emph{Annual
Review of Political Science}, \emph{15}(Volume 15, 2012), 333--351.
\url{https://doi.org/10.1146/annurev-polisci-012610-135202}

\bibitem[\citeproctext]{ref-canalesceron_sujeto_2021}
Canales Cerón, M., Orellana Calderón, V. S., \& Guajardo Mañán, F.
(2021). Sujeto y cotidiano en la era neoliberal: El caso de la
educaci{ó}n chilena. \emph{Revista Mexicana de Ciencias Pol{í}ticas y
Sociales}, \emph{67}(244).
\url{https://doi.org/10.22201/fcpys.2448492xe.2022.244.70386}

\bibitem[\citeproctext]{ref-castillo_cual_2009}
Castillo, J. C. (2009). {{\textquestiondown}Cu{á}l es la brecha salarial
justa? Opini{ó}n p{ú}blica y legitimaci{ó}n de la desigualdad en Chile}.
\emph{Estudios P{ú}blicos}, (113).

\bibitem[\citeproctext]{ref-Castillo2011}
Castillo, J. C. (2011). Legitimacy of {Inequality} in a {Highly Unequal
Context}: {Evidence} from the {Chilean Case}. \emph{Social Justice
Research}, \emph{24}(4), 314--340.
\url{https://doi.org/10.1007/s11211-011-0144-5}

\bibitem[\citeproctext]{ref-Castillo2012a_justice}
Castillo, J. C. (2012). Is {Inequality Becoming Just}? {Changes} in
{Public Opinion} about {Economic Distribution} in {Chile}.
\emph{Bulletin of Latin American Research}, \emph{31}(1), 1--18.
\url{https://doi.org/10.1111/j.1470-9856.2011.00605.x}

\bibitem[\citeproctext]{ref-castillo_perception_2022}
Castillo, J. C., García-Castro, J.-D., \& Venegas, M. (2022). Perception
of economic inequality: Concepts, associated factors and prospects of a
burgeoning research agenda. \emph{International Journal of Social
Psychology}, \emph{37}(1), 180--207.
\url{https://doi.org/10.1080/02134748.2021.2009275}

\bibitem[\citeproctext]{ref-castillo_changes_2025}
Castillo, J. C., Iturra, J., \& Carrasco, K. (2025). Changes in the
{Justification} of {Educational Inequalities}: {The Role} of
{Perceptions} of {Inequality} and {Meritocracy During} the {COVID
Pandemic}. \emph{Social Justice Research}.
\url{https://doi.org/10.1007/s11211-025-00458-0}

\bibitem[\citeproctext]{ref-castillo_multidimensional_2023}
Castillo, J. C., Iturra, J., Maldonado, L., Atria, J., \& Meneses, F.
(2023). A {Multidimensional Approach} for {Measuring Meritocratic
Beliefs}: {Advantages}, {Limitations} and {Alternatives} to the {ISSP
Social Inequality Survey}. \emph{International Journal of Sociology},
1--25. \url{https://doi.org/10.1080/00207659.2023.2274712}

\bibitem[\citeproctext]{ref-castillo_percepcion_2019}
Castillo, J. C., Miranda, D., \& Carrasco, D. (2012). Percepci{ó}n de
{Desigualdad Econ{ó}mica} en {Chile}: {Medici{ó}n}, {Diferencias} y
{Determinantes}. \emph{Psykhe (Santiago)}, \emph{21}(1), 99--114.
\url{https://doi.org/10.4067/S0718-22282012000100007}

\bibitem[\citeproctext]{ref-castillo_socialization_2024}
Castillo, J. C., Salgado, M., Carrasco, K., \& Laffert, A. (2024). The
{Socialization} of {Meritocracy} and {Market Justice Preferences} at
{School}. \emph{Societies}, \emph{14}(11), 214.
\url{https://doi.org/10.3390/soc14110214}

\bibitem[\citeproctext]{ref-chancel_world_2022}
Chancel, L., Piketty, T., Saez, E., \& Zucman, G. (2022). World
inequality report 2022.
https://bibliotecadigital.ccb.org.co/handle/11520/27510.

\bibitem[\citeproctext]{ref-davis_principles_2001}
Davis, K., \& Moore, W. E. (2001). Some {Principles} of
{Stratification}. In \emph{Social {Stratification}, {Class}, {Race}, and
{Gender} in {Sociological Perspective}, {Second Edition}} (2nd ed.).
Routledge.

\bibitem[\citeproctext]{ref-disipavlic_justification_2025}
Disi Pavlic, R., Medel, R. M., Bargsted, M., \& Somma, N. M. (2025).
Justification of violence, ideological preferences, and exposure to
protests: Causal evidence from the 2019 {Chilean} social unrest.
\emph{Social Forces}, soaf102. \url{https://doi.org/10.1093/sf/soaf102}

\bibitem[\citeproctext]{ref-easterbrook_social_2021}
Easterbrook, M. J. (2021). \emph{The social psychology of economic
inequality} (43rd ed., Vol. 2021). UNU-WIDER.
\url{https://doi.org/10.35188/UNU-WIDER/2021/981-5}

\bibitem[\citeproctext]{ref-engelhardt_what_2018}
Engelhardt, C., \& Wagener, A. (2018). What do {Germans} think and know
about income inequality? {A} survey experiment. \emph{Socio-Economic
Review}, \emph{16}(4), 743--767.
\url{https://doi.org/10.1093/ser/mwx036}

\bibitem[\citeproctext]{ref-espinoza_movilidad_2014}
Espinoza, V., \& Núñez, J. (2014). Movilidad ocupacional en {Chile}
2001-2009. {\textquestiondown}{Desigualdad} de ingresos con igualdad de
oportunidades? \emph{Revista Internacional de Sociolog{í}a},
\emph{72}(1), 57--82. \url{https://doi.org/10.3989/ris.2011.11.08}

\bibitem[\citeproctext]{ref-ferre_welfare_2023}
Ferre, J. C. (2023). Welfare regimes in twenty-first-century {Latin
America}. \emph{Journal of International and Comparative Social Policy},
\emph{39}(2), 101--127. \url{https://doi.org/10.1017/ics.2023.16}

\bibitem[\citeproctext]{ref-flores_top_2020}
Flores, I., Sanhueza, C., Atria, J., \& Mayer, R. (2020). Top {Incomes}
in {Chile}: {A Historical Perspective} on {Income Inequality},
1964--2017. \emph{Review of Income and Wealth}, \emph{66}(4), 850--874.
\url{https://doi.org/10.1111/roiw.12441}

\bibitem[\citeproctext]{ref-garcia-castro_perceiving_2020}
García-Castro, J. D., Rodríguez-Bailón, R., \& Willis, G. B. (2020).
Perceiving economic inequality in everyday life decreases tolerance to
inequality. \emph{Journal of Experimental Social Psychology}, \emph{90},
104019. \url{https://doi.org/10.1016/j.jesp.2020.104019}

\bibitem[\citeproctext]{ref-garcia-sanchez_creencias_2022}
García-Sánchez, E., \& De Carvalho, S. (2022). Las creencias que
justifican la desigualdad moderan la relaci{ó}n entre el estatus
socioecon{ó}mico y el apoyo a la redistribuci{ó}n. \emph{Revista
Internacional de Sociolog{í}a}, \emph{80}(3), e210.
\url{https://doi.org/10.3989/ris.2022.80.3.21.29}

\bibitem[\citeproctext]{ref-garcia-sanchez_attitudes_2020}
García-Sánchez, E., Osborne, D., Willis, G. B., \& Rodríguez-Bailón, R.
(2020). Attitudes towards redistribution and the interplay between
perceptions and beliefs about inequality. \emph{British Journal of
Social Psychology}, \emph{59}(1), 111--136.
\url{https://doi.org/10.1111/bjso.12326}

\bibitem[\citeproctext]{ref-garcia-sanchez_vicious_2019}
García-Sánchez, E., Van Der Toorn, J., Rodríguez-Bailón, R., \& Willis,
G. B. (2019). The {Vicious Cycle} of {Economic Inequality}: {The Role}
of {Ideology} in {Shaping} the {Relationship Between} {``{What Is}''}
and {``{What Ought} to {Be}''} in 41 {Countries}. \emph{Social
Psychological and Personality Science}, \emph{10}(8), 991--1001.
\url{https://doi.org/10.1177/1948550618811500}

\bibitem[\citeproctext]{ref-garcia-sanchez_perceptions_2018}
García-Sánchez, E., Willis, G. B., Rodríguez-Bailón, R., Palacio Sañudo,
J., David Polo, J., \& Rentería Pérez, E. (2018). Perceptions of
{Economic Inequality} and {Support} for {Redistribution}: {The} role of
{Existential} and {Utopian Standards}. \emph{Social Justice Research},
\emph{31}(4), 335--354. \url{https://doi.org/10.1007/s11211-018-0317-6}

\bibitem[\citeproctext]{ref-gijsberts_thelegitimation_1999}
Gijsberts, M. (1999). \emph{{Thelegitimation of inequality in state-
socialist and market societies, 1987 - 1996}}. Amsterdam: Thela Thesis.

\bibitem[\citeproctext]{ref-gimpelson_misperceiving_2018}
Gimpelson, V., \& Treisman, D. (2018). Misperceiving inequality.
\emph{Economics \& Politics}, \emph{30}(1), 27--54.
\url{https://doi.org/10.1111/ecpo.12103}

\bibitem[\citeproctext]{ref-hadler_why_2005}
Hadler, M. (2005). Why {Do People Accept Different Income Ratios}?: {A
Multi-level Comparison} of {Thirty Countries}. \emph{Acta Sociologica},
\emph{48}(2), 131--154. \url{https://doi.org/10.1177/0001699305053768}

\bibitem[\citeproctext]{ref-hoxMultilevelAnalysisTechniques2017a}
Hox, J. J., Moerbeek, M., \& Van de Schoot, R. (2017). \emph{Multilevel
{Analysis}: {Techniques} and {Applications}}.

\bibitem[\citeproctext]{ref-hoyt_mindsets_2023}
Hoyt, C. L., Burnette, J. L., Billingsley, J., Becker, W., \& Babij, A.
D. (2023). Mindsets of poverty: {Implications} for redistributive policy
support. \emph{Analyses of Social Issues and Public Policy},
\emph{23}(3), 668--693. \url{https://doi.org/10.1111/asap.12367}

\bibitem[\citeproctext]{ref-immergut_it_2020}
Immergut, E. M., \& Schneider, S. M. (2020). Is it unfair for the
affluent to be able to purchase {``better''} healthcare? {Existential}
standards and institutional norms in healthcare attitudes across 28
countries. \emph{Social Science \& Medicine}, \emph{267}, 113146.
\url{https://doi.org/10.1016/j.socscimed.2020.113146}

\bibitem[\citeproctext]{ref-janmaat_subjective_2013}
Janmaat, J. G. (2013). Subjective inequality: {A} review of
international comparative studies on people's views about inequality.
\emph{Archives Europeennes de Sociologie}, \emph{54}(3), 357--389.
\url{https://doi.org/10.1017/S0003975613000209}

\bibitem[\citeproctext]{ref-kluegel_social_1995a}
Kluegel, J. R., Mason, D. S., \& Wegener, B. (Eds.). (1995).
\emph{Social {Justice} and {Political Change}: {Public Opinion} in
{Capitalist} and {Post-Communist States}} (1st ed.). Routledge.

\bibitem[\citeproctext]{ref-kluegel_legitimation_1999}
Kluegel, J. R., Mason, D. S., \& Wegener, B. (1999). The {Legitimation}
of {Capitalism} in the {Postcommunist Transition}: {Public Opinion}
about {Market Justice}, 1991-1996. \emph{European Sociological Review},
\emph{15}(3), 251--283. Retrieved from
\url{https://www.jstor.org/stable/522731}

\bibitem[\citeproctext]{ref-kluegel_beliefs_1981}
Kluegel, J. R., \& Smith, E. R. (1981). Beliefs {About Stratification}.
\emph{Annual Review of Sociology}, 29--56.

\bibitem[\citeproctext]{ref-kluegel_beliefs_1986}
Kluegel, J. R., \& Smith, E. R. (1986). \emph{Beliefs about
{Inequality}: {Americans}' {Views} of {What Is} and {What Ought} to
{Be}} (1st ed.). Routledge.

\bibitem[\citeproctext]{ref-knell_perceptions_2020}
Knell, M., \& Stix, H. (2020). Perceptions of inequality. \emph{European
Journal of Political Economy}, \emph{65}, 101927.
\url{https://doi.org/10.1016/j.ejpoleco.2020.101927}

\bibitem[\citeproctext]{ref-koos_moral_2019}
Koos, S., \& Sachweh, P. (2019). The moral economies of market
societies: Popular attitudes towards market competition, redistribution
and reciprocity in comparative perspective. \emph{Socio-Economic
Review}, \emph{17}(4), 793--821.
\url{https://doi.org/10.1093/ser/mwx045}

\bibitem[\citeproctext]{ref-kuhn_eye_2011}
Kuhn, A. (2011). In the eye of the beholder: {Subjective} inequality
measures and individuals' assessment of market justice. \emph{European
Journal of Political Economy}, \emph{27}(4), 625--641.
\url{https://doi.org/10.1016/j.ejpoleco.2011.06.002}

\bibitem[\citeproctext]{ref-lane_political_1962}
Lane, R. E. (1962). \emph{Political ideology: Why the {American} common
man believes what he does}. Oxford, England: Free Press of Glencoe.

\bibitem[\citeproctext]{ref-lane_market_1986}
Lane, Robert E. (1986). Market {Justice}, {Political Justice}.
\emph{American Political Science Review}, \emph{80}(2), 383--402.
\url{https://doi.org/10.2307/1958264}

\bibitem[\citeproctext]{ref-lee_fairness_2023}
Lee, J.-S., \& Stacey, M. (2023). Fairness perceptions of educational
inequality: The effects of self-interest and neoliberal orientations.
\emph{The Australian Educational Researcher}.
\url{https://doi.org/10.1007/s13384-023-00636-6}

\bibitem[\citeproctext]{ref-lindh_public_2015}
Lindh, A. (2015). Public {Opinion} against {Markets}? {Attitudes}
towards {Market Distribution} of {Social Services} -- {A Comparison} of
17 {Countries}. \emph{Social Policy \& Administration}, \emph{49}(7),
887--910. \url{https://doi.org/10.1111/spol.12105}

\bibitem[\citeproctext]{ref-lindh_bringing_2023}
Lindh, A., \& McCall, L. (2023). Bringing the market in: An expanded
framework for understanding popular responses to economic inequality.
\emph{Socio-Economic Review}, \emph{21}(2), 1035--1055.
\url{https://doi.org/10.1093/ser/mwac018}

\bibitem[\citeproctext]{ref-lopez-roldan_comparative_2021}
López-Roldán, P., \& Fachelli, S. (Eds.). (2021). \emph{Towards a
{Comparative Analysis} of {Social Inequalities} between {Europe} and
{Latin America}}. Cham: Springer International Publishing.
\url{https://doi.org/10.1007/978-3-030-48442-2}

\bibitem[\citeproctext]{ref-mac-clure_justicia_2024}
Mac-Clure, O., Barozet, E., \& Franetovic, G. (2024). {Justicia
distributiva y posici{ó}n social subjetiva: {\textquestiondown}la
meritocracia justifica la desigualdad de ingresos?} \emph{Convergencia
Revista de Ciencias Sociales}, \emph{31}, 1.
\url{https://doi.org/10.29101/crcs.v31i0.22258}

\bibitem[\citeproctext]{ref-madariaga_three_2020}
Madariaga, A. (2020). The three pillars of neoliberalism: {Chile}'s
economic policy trajectory in comparative perspective.
\emph{Contemporary Politics}, \emph{26}(3), 308--329.
\url{https://doi.org/10.1080/13569775.2020.1735021}

\bibitem[\citeproctext]{ref-mccall_exposure_2017}
McCall, L., Burk, D., Laperrière, M., \& Richeson, J. A. (2017).
Exposure to rising inequality shapes {Americans}' opportunity beliefs
and policy support. \emph{Proceedings of the National Academy of
Sciences}, \emph{114}(36), 9593--9598.
\url{https://doi.org/10.1073/pnas.1706253114}

\bibitem[\citeproctext]{ref-mijs_stratified_2016}
Mijs, J. (2016). Stratified {Failure}: {Educational Stratification} and
{Students}' {Attributions} of {Their Mathematics Performance} in 24
{Countries}. \emph{Sociology of Education}, \emph{89}(2), 137--153.
\url{https://doi.org/10.1177/0038040716636434}

\bibitem[\citeproctext]{ref-mijs_paradox_2021}
Mijs, J. (2021). The paradox of inequality: Income inequality and belief
in meritocracy go hand in hand. \emph{Socio-Economic Review},
\emph{19}(1), 7--35. \url{https://doi.org/10.1093/ser/mwy051}

\bibitem[\citeproctext]{ref-pedersen_attitudes_2019}
Pedersen, R. T., \& Mutz, D. C. (2019). Attitudes {Toward Economic
Inequality}: {The Illusory Agreement}. \emph{Political Science Research
and Methods}, \emph{7}(04), 835--851.
\url{https://doi.org/10.1017/psrm.2018.18}

\bibitem[\citeproctext]{ref-pierson_when_1993}
Pierson, P. (1993). When {Effect Becomes Cause}: {Policy Feedback} and
{Political Change}. \emph{World Politics}, \emph{45}(4), 595--628.
\url{https://doi.org/10.2307/2950710}

\bibitem[\citeproctext]{ref-reynolds_perceptions_2014}
Reynolds, J., \& Xian, H. (2014). Perceptions of meritocracy in the land
of opportunity. \emph{Research in Social Stratification and Mobility},
\emph{36}, 121--137. \url{https://doi.org/10.1016/j.rssm.2014.03.001}

\bibitem[\citeproctext]{ref-sandel_tyranny_2020}
Sandel, M. J. (2020). \emph{The tyranny of merit: {What}'s become of the
common good?} (First edition). New York: {Farrar, Straus and Giroux}.

\bibitem[\citeproctext]{ref-schneider_poverty_2015}
Schneider, S. M., \& Castillo, J. C. (2015). Poverty {Attributions} and
the {Perceived Justice} of {Income Inequality} : {A Comparison} of
{East} and {West Germany}.
\url{https://doi.org/10.1177/0190272515589298}

\bibitem[\citeproctext]{ref-schroder_income_2017}
Schröder, M. (2017). Is {Income Inequality Related} to {Tolerance} for
{Inequality}? \emph{Social Justice Research}, \emph{30}(1), 23--47.
\url{https://doi.org/10.1007/s11211-016-0276-8}

\bibitem[\citeproctext]{ref-singer_applied_2009}
Singer, J. D., \& Willett, J. B. (2009). \emph{Applied longitudinal data
analysis: Modeling change and event occurence}. New York: Oxford
University Press, Incorporated.

\bibitem[\citeproctext]{ref-somma_no_2021}
Somma, N. M., Bargsted, M., Disi Pavlic, R., \& Medel, R. M. (2021). No
water in the oasis: The {Chilean Spring} of 2019--2020. \emph{Social
Movement Studies}, \emph{20}(4), 495--502.
\url{https://doi.org/10.1080/14742837.2020.1727737}

\bibitem[\citeproctext]{ref-svallfors_political_2007}
Svallfors, S. (Ed.). (2007). \emph{The {Political Sociology} of the
{Welfare State}: {Institutions}, {Social Cleavages}, and {Orientations}}
(1st ed.). Stanford University Press.
\url{https://doi.org/10.2307/j.ctvr0qv0q}

\bibitem[\citeproctext]{ref-tejero-peregrina_perceived_2025}
Tejero-Peregrina, L., Willis, G., Sánchez-Rodríguez, Á., \&
Rodríguez-Bailón, R. (2025). From {Perceived Economic Inequality} to
{Support} for {Redistribution}: {The Role} of {Meritocracy Perception}.
\emph{International Review of Social Psychology}, \emph{38}(1), 4.
\url{https://doi.org/10.5334/irsp.1013}

\bibitem[\citeproctext]{ref-tilly_contentious_2015}
Tilly, C., \& Tarrow, S. (2015). \emph{Contentious {Politics}} (Second
Edition, New to this Edition:, Second Edition, New to this Edition:).
Oxford, New York: Oxford University Press.

\bibitem[\citeproctext]{ref-torche_intergenerational_2014a}
Torche, F. (2014). Intergenerational {Mobility} and {Inequality}: {The
Latin American Case}. \emph{Annual Review of Sociology}, \emph{40}(1),
619--642. \url{https://doi.org/10.1146/annurev-soc-071811-145521}

\bibitem[\citeproctext]{ref-trump_income_2018}
Trump, K.-S. (2018). Income {Inequality Influences Perceptions} of
{Legitimate Income Differences}. \emph{British Journal of Political
Science}, \emph{48}(4), 929--952.
\url{https://doi.org/10.1017/S0007123416000326}

\bibitem[\citeproctext]{ref-vondemknesebeck_are_2016}
Von Dem Knesebeck, O., Vonneilich, N., \& Kim, T. J. (2016). Are health
care inequalities unfair? {A} study on public attitudes in 23 countries.
\emph{International Journal for Equity in Health}, \emph{15}(1), 61.
\url{https://doi.org/10.1186/s12939-016-0350-8}

\bibitem[\citeproctext]{ref-willis_legitimacy_2015}
Willis, G. B., Rodríguez-Bailón, R., López-Rodríguez, L., \&
García-Sánchez, E. (2015). Legitimacy {Moderates} the {Relation Between
Perceived} and {Ideal Economic Inequalities}. \emph{Social Justice
Research}, \emph{28}(4), 493--508.
\url{https://doi.org/10.1007/s11211-015-0253-7}

\bibitem[\citeproctext]{ref-wilson_role_2003}
Wilson, C. (2003). The {Role} of a {Merit Principle} in {Distributive
Justice}. \emph{The Journal of Ethics}, \emph{7}(3), 277--314.
\url{https://doi.org/10.1023/A:1024667228488}

\bibitem[\citeproctext]{ref-young_rise_1962}
Young, M. (1962). \emph{The rise of the meritocracy}. Baltimore: Penguin
Books.

\end{CSLReferences}

\end{document}