### 2025-07-28 resubmission/05-discussion.qmd
update discussion and conclusion (no render)

---
author: "Equipo EDUMER"
bibliography: "../input/bib/merit-edjust.bib"
csl: "../input/bib/apa6.csl"
---

# Discussion

This study tests several key hypotheses concerning the relationships between perceptions of economic inequality, meritocratic perceptions, and attitudes toward market justice in Chile. It posits that higher perceived inequality and stronger meritocratic perceptions both independently contribute to greater support for market-based distributions of social goods. Furthermore, the study examines how perceived economic inequality moderates the influence of meritocratic perceptions on market justice preferences. Additionally, it considers how major social and political mobilizations, such as those occurring around 2019 and subsequent years, may have weakened these relationships, reflecting increased critical perspectives on systemic inequalities and market mechanisms. The subsequent discussion interprets the empirical findings in light of these hypotheses, considering their implications for understanding attitudes toward inequality and social justice in a context marked by profound social change.

The first hypothesis suggested that perceptions of economic inequality were strongly linked to attitudes toward market justice. The findings demonstrated that individuals perceiving higher income disparities showed less support for the distribution of social services based on individual income, aligning with theories emphasizing that awareness of inequality could fuel a critical view of such policies [@Castillo2012a_justice; @mijs_paradox_2021]. This relationship reflected a moral economy where perceptions of systemic unfairness threatened the legitimacy of existing distributions, prompting calls for greater equity. Similarly, the negative within-effects observed over time ($H1_{b}$) implied that as perceptions of inequality became more salient, some individuals experienced a decline in support for market justice, possibly due to increasing distrust in market mechanisms or disillusionment with the fairness of the system. This indicated that while perception heightened awareness, prolonged exposure or social mobilization could also further foster skepticism about whether redistribution effectively addressed inequality or simply perpetuated systemic flaws.

Regarding the second hypothesis, the results reinforced that meritocratic perceptions ---particularly those emphasizing individual effort--- were associated with stronger support for market-based distribution systems. Individuals who viewed success as primarily earned through effort were more inclined to justify existing inequalities, seeing them as outcomes of individual merit rather than systemic injustice. This aligned with research indicating that meritocratic narratives served as moral justifications that reinforced societal stratification [@Castillo2012a_justice; @hoyt_mindsets_2023]. These perceptions functioned symbolically to legitimize structural disparities, reducing support for redistributive measures by framing inequality as fair and deserved. The findings suggested that these perceptions bolstered the social hierarchy by fostering acceptance of inequality as a reflection of individual virtue rather than structural failure, which was especially pertinent in neoliberal contexts like Chile's, where market logic heavily influenced social attitudes [@canalesceron_sujeto_2021]. Interestingly, intraindividual changes in meritocratic perceptions over time ($H2_{b}$) reveal mixed effects. While an increase in the perception that rewards are based on individual effort is associated with stronger preferences for market fairness, an increase in the belief that rewards come from talent is linked to lower support for such principles. One possible explanation is that effort is generally perceived as a controllable and malleable trait, which aligns with the normative foundations of market fairness. Conversely, talent may be viewed as an innate and less alterable attribute, making talent-based reward allocation seem less fair or legitimate. Moreover, greater exposure to real-life situations where outcomes are strongly determined by inherent characteristics, rather than effort or hard work, may lead people to question the fairness of market-based reward systems.

The third hypothesis concerned the moderating role of perceptions of inequality on the relationship between meritocracy and market justice attitudes. The analysis indicated that the positive association between meritocratic perceptions and support for market-based distribution tended to strengthen when perceptions of economic inequality were particularly high. While this moderation is not significant for effort-related effects, it is significant for talent in the within- and between-individual effects. This appears as contrary to our initial hypothesis, as it would mean that effect of meritocratic perceptions becomes stronger as perception of inequality increases. Nevertheless, we need to be cautious about this interpretation, as the main effect of inequality perception on market justice is negative. Therefore, as perceived inequality become less negative (i.e. more positive), or closer to zero, then the meritocratic perception has an stronger positive relationship on market justice preferences. This could mean that meritocratic perceptions play an stronger role in justification of market justice inequalities for those who are less capable or less prone to perceive inequality.

Finally, the fourth hypothesis addressed the impact of recent social movements—particularly those that emerged in 2019—on the relationship between meritocratic beliefs, perceptions of inequality, and attitudes toward market justice. The main effect of time suggests that, on average, support for market justice increased after 2019. However, the results do not reveal any significant interactions between distributive mechanisms and market justice attitudes, indicating that the 2019 social mobilizations did not substantially alter the underlying associations between these variables. This may suggest a certain stability or resilience in the normative frameworks that shape individuals’ evaluations of justice, even in the context of major collective political events. Alternatively, it is possible that while the protests challenged institutional legitimacy, they did not fundamentally shift individuals’ views regarding how rewards should be allocated in society.