### 2025-08-04 resubmission/02-antecedents.qmd

# Theoretical views on market justice, inequality perception, and meritocracy

## The justification of market inequality

Conceptually, *market justice* has been discussed in the literature as a normative principle that legitimates the distribution of economic rewards based on individual merit. It is possible to trace the concept to the understanding of Lane [-@lane_market_1986], who makes a contrast between market justice and political justice. The author defines market justice as a system of "earned deserts", whereby individuals are seen as deserving of a determined distributive outcome due to their effort and skills. In contrast, political justice emphasizes principles of equality and need, which are often represented by the welfare state action through social policies. An important remark is that the principles of market justice rely on the assumption that markets are neutral and self-regulating arenas, where individuals are treated fairly because they face the same formal rules of engagement and procedures [@lane_market_1986]. Consequently, the legitimacy of market justice stems from the assumption that inequalities are not only inevitable but fair---so long as the rules are transparent and opportunities are open. In this way, market justice provides a moral justification for inequality by framing it as a necessary outcome of individual responsibility [@kluegel_legitimation_1999; @lindh_public_2015].

Empirical studies have shown different strategies for the study of market justice preferences. A common approach in the literature is to gauge attitudes toward the legitimacy of inequality in specific domains, especially when linked to income differences. This can be traced to the seminal work of Kluegel and Smith [-@kluegel_beliefs_1981] who assessed the normative foundations that explain public support for economic inequality. Over time, this approach has been extended beyond income to include other market-mediated outcomes, such as education, healthcare, and/or pensions. For example, @vondemknesebeck_are_2016 and Immergut and Schneider [-@immergut_it_2020] examine whether citizens consider it fair that individuals with higher incomes can access better healthcare, while Lee and Stacey [-@lee_fairness_2023] apply a similar method in the context of education in Australia. These studies usually rely on a survey item asking respondents to evaluate the fairness of income-based access to welfare services, allowing for comparing justice perceptions across different contexts. Similarly, comparative studies have also considered these indicators to study the support for market-based distribution in welfare systems [@svallfors_political_2007;@lindh_public_2015]. More recently, Castillo et al. [-@castillo_socialization_2024] introduced a single-item composite measure of market justice to assess student attitudes toward income-based access to education, healthcare, and pensions in Chile. These empirical strategies all aim to capture the extent to which individuals accept inequality when framed as a reflection of market outcomes.

The study of market justice preferences has increasingly focused on how they are shaped by individuals' socioeconomic position, normative beliefs, and the institutional context in which they are embedded. Across the literature, there is empirical evidence suggesting that individuals in more advantaged socioeconomic positions---those with higher occupational class, income, and education---are more likely to support market justice principles [@svallfors_political_2007; @koos_moral_2019; @lindh_public_2015]. This tendency reflects not only material self-interest but also a broader moral economy, in which winners of the market system internalize justifications for the status quo. At the same time, political ideology also plays a role ---such as economically conservative values ---  where more right-wing individuals show higher support for meritocracy and more skepticism towards redistribution. This is particularly salient in countries with more restricted public provision of social services. For example, Castillo et al. [-@castillo_socialization_2024] in Chile and, Lee and Stacey [-@lee_fairness_2023] in Australia show that right-wing individuals are more supportive of market-based distribution of welfare. Beyond individual characteristics, country-level institutions also play a central role. In liberal welfare regimes like those of the United States or the United Kingdom, market justice preferences are more widespread, while in coordinated or social-democratic regimes---such as Sweden or Germany---citizens are generally more critical of market-based inequalities [@immergut_it_2020; @lindh_public_2015]. These findings suggest that market justice is also shaped contextually, which is represented by the development of welfare institutions, in line with the policy-feedbacks literature [@campbell_policy_2012; @pierson_when_1993].

## Perception of inequality

Perceptions of inequality have been associated with a range of distributive attitudes [@castillo_perception_2022; @garcia-sanchez_vicious_2019; @trump_income_2018; @mccall_exposure_2017]. Research indicates that perceptions of inequality influence redistributive preferences to a greater extent than objective or actual economic inequality [@gimpelson_misperceiving_2018]. Others have shown that the perception of inequality, through the perceived wage gap, is related to the justification of greater inequality [@Castillo2011]. In that sense, lower perceived inequality can reinforce support for market-based distributive arrangements by suggesting that the system is fair and that outcomes reflect effort and ability [@kuhn_eye_2011]. In contrast, when inequality is perceived as excessive or structurally determined, individuals could question the legitimacy of market justice and become more supportive of redistributive policies [@garcia-sanchez_vicious_2019].

Perceptions of inequality refer to individuals' subjective evaluations of the extent, causes, and consequences of income and wealth disparities [@akyelken_urban_2020]. Unlike objective measures such as the Gini index, perceived inequality captures how individuals make sense of distributive hierarchies in their everyday lives, shaped by reference groups, social comparisons, and information environments [@mijs_stratified_2016; @gimpelson_misperceiving_2018; @garcia-castro_perceiving_2020]. Scholars have proposed multiple dimensions of perceived inequality, including its magnitude (how significant are the gaps), vertical structure (between which groups), the trend over time (increasing or decreasing), and legitimacy (whether it is just or not) [@engelhardt_what_2018; @garcia-sanchez_vicious_2019]. These dimensions encompass both cognitive and normative aspects of perceptions of inequality and can vary across societies and social groups, depending on exposure, ideology, and personal experience [@castillo_perception_2022; @garcia-sanchez_perceptions_2018].

Among the different approaches to measuring perceived inequality, one of the most widely used is the estimation of wage gaps between occupational extremes, such as between a CEO and a manual worker. This type of item provides a concrete frame that respondents can relate to more easily than abstract questions about national income distribution [@castillo_percepcion_2019; @easterbrook_social_2021; @willis_legitimacy_2015]. While it enables the estimation of inequality using simple heuristics, this method is not without challenges. For instance, people often lack reliable knowledge about the earnings of those at the top of the income ladder, which leads to high variability in responses and the use of biased mental shortcuts [@knell_perceptions_2020]. Despite this, perceived wage gaps are strong predictors of political attitudes [@pedersen_attitudes_2019; @garcia-sanchez_perceptions_2018], making them a valuable tool for understanding public responses to economic disparities.

Another relevant critique of this approach lies in its conflation of different psychological constructs. Many surveys assess perceived inequality through Likert-type items that ask respondents to agree or disagree with statements, such as "income differences are too large," which captures general concern or discomfort rather than a specific perception [@Castillo2011; @garcia-sanchez_vicious_2019]. These items mix cognitive estimations with affective evaluations, complicating the interpretation of what respondents perceive versus what they morally reject. As a result, the conceptual clarity between perceived inequality and inequality aversion remains blurred in many empirical studies. To address this limitation, recent work has emphasized the need to distinguish between absolute and comparative measures, as well as between ideal and actual estimates of economic gaps [@garcia-sanchez_creencias_2022; @auspurg_why_2017].

## Perception of meritocracy

Meritocracy constitutes a central ideological framework for legitimizing different types of social inequality, for instance through market justice beliefs. Rooted in the belief that rewards and positions should be allocated based on individual effort and talent, meritocracy operates as a normative ideal and a descriptive belief about how society functions. As initially conceptualized by Michael Young [-@young_rise_1962], the term was meant to critique a system in which merit-based stratification becomes a new form of inequality. However, over time, meritocracy has been widely supported in many societies as a fair and desirable principle of distribution, particularly within liberal democracies and market-oriented economies [@mijs_paradox_2021; @sandel_tyranny_2020]. From a sociological perspective, the belief in meritocracy is more than a cognitive assessment; it reflects a moral lens through which individuals interpret inequality. People who believe that success results from hard work and talent are more likely to view social and economic disparities as legitimate [@castillo_percepcion_2019; @batruch_belief_2023]. Conversely, if they see outcomes as driven by luck, social origin, or systemic barriers, inequality is more likely to be perceived as unjust. This distinction becomes crucial in societies with persistent structural inequality, where public narratives often emphasize personal responsibility and merit while overlooking entrenched disadvantages.

We adopt a multidimensional perspective on meritocracy, distinguishing between two key dimensions: effort-based and talent-based perceptions. This distinction is essential, as it captures different pathways through which individuals justify inequality. Effort-based meritocracy emphasizes hard work and perseverance as the basis for social rewards, aligning closely with cultural narratives of personal responsibility. A talent-based meritocracy, by contrast, emphasizes intelligence and innate abilities, which are often perceived as less malleable and more unequally distributed. Both dimensions have been shown to correlate with acceptance of inequality, but they may carry distinct implications for how inequality is justified in specific domains [@castillo_multidimensional_2023]. The relevance of this distinction is supported by recent studies, which show that individuals respond differently to these dimensions. For instance, perceptions that effort is rewarded in society are more strongly associated with positive evaluations of fairness and acceptance of unequal outcomes [@batruch_belief_2023]. This may be because effort is seen as a controllable and morally virtuous trait, whereas talent is often perceived as a natural advantage. Consequently, effort-based meritocracy is likely more potent in legitimizing inequality, particularly in neoliberal contexts.

These dimensions of meritocracy reflect how respondents perceive society's distributive logic, regardless of whether they endorse meritocratic principles. This distinction aligns with recent findings indicating that individuals distinguish between how merit is perceived in society and how it should ideally operate, which in turn shapes their preferences for redistribution and justice [@tejero-peregrina_perceived_2025]. Meritocratic beliefs serve as symbolic justifications for unequal outcomes, particularly when access is stratified by income or social opportunity. Prior studies have shown that individuals who perceive higher levels of meritocracy tend to express stronger support for unequal distributions that reflect market outcomes [@castillo_percepcion_2019; @castillo_socialization_2024].

In addition to influencing individual attitudes toward inequality, meritocratic perceptions can contribute to social division and the stigmatization of disadvantaged groups. Recent research has demonstrated that exposure to meritocratic narratives can reinforce the belief that poverty results from individual failings rather than systemic conditions, reducing support for redistributive measures and increasing the stigmatization of the poor [@hoyt_mindsets_2023]. This reinforces negative stereotypes and reduces empathy toward individuals from lower socioeconomic backgrounds. Moreover, @busemeyer_positive_2021 argues that meritocratic narratives can serve as feedback mechanisms that shape public opinion and well-being by framing individuals' understanding of welfare outcomes as deserved or undeserved within existing institutional structures. This psychological mechanism highlights the normative power of meritocracy in stabilizing unequal systems by shaping political attitudes and personal perceptions of success and failure.

## The Chilean context

Chile offers an interesting case for examining how public attitudes toward the distribution of social services evolve amid declining poverty and persistently high income inequality within a residual welfare framework [@ferre_welfare_2023]. Following the neoliberal reforms of the 1980s introduced during the civil-military dictatorship (1973-1989), the Chilean welfare state has become increasingly reliant on private provision, with essential services often commodified and largely reserved for those with sufficient purchasing power [@madariaga_three_2020; @boccardo_30_2020]. This model has favored the affluent, while lower-income groups are left dependent on limited and underfunded public alternatives. Research on social stratification characterizes Chile as a society where mobility into intermediate classes is feasible, but access to elite strata remains largely restricted [@lopez-roldan_comparative_2021; @torche_intergenerational_2014a]. While some indicators—such as education and occupation—suggest a relatively fluid class structure, income mobility remains constrained [@espinoza_movilidad_2014].

Despite steady economic expansion, Chile continues to rank among the most unequal nations in the OECD, marked by a high Gini coefficient and substantial wealth concentration at the top [@flores_top_2020]. In this context, Chile experienced a wave of mass protests from October 2019 to March 2020. Initially led by high school students, the demonstrations quickly gained momentum as broader segments of society mobilized around demands for more equality of opportunities, particularly regarding precarious access to welfare services. The political system interpreted this unrest as a call to reform existing political institutions, with a special focus on equalizing opportunities and providing more egalitarian access to welfare by reducing reliance on out-of-pocket mechanisms [@somma_no_2021].

Concerning public opinion, the ELSOC survey shows that Chile has a low level of agreement with the idea that it is fair for those with higher incomes to buy better welfare in the areas of education, healthcare, and pensions, with average support of 8.5% between 2016 and 2023. At the same time, around 55% and 47% of the public between 2016 and 2023 *disagreed* that effort and talent are rewarded accordingly. This coincides with 90% of citizens *agreeing* that income differences in Chile are *too large* [@castillo_changes_2025].

In sum, Chile represents an illustrative case of a particular form of *moral economy* [@koos_moral_2019], in which national socioeconomic characteristics, institutional arrangements, and public attitudes coexist. This suggests that the country combines persistently high levels of economic and social inequality with a strong institutionalization of market-based mechanisms within the welfare system [@madariaga_three_2020]. At the same time, perceptions of inequality are high, and there is widespread public criticism of the commodification of social services [@somma_no_2021]. For the above reasons, we argue that the Chilean case represents a relevant context to study the role of perceived income inequality and meritocracy in shaping public preferences regarding the commodification of welfare services, particularly from the perspective of individual change.

## This study

Building upon the previous literature, this study proposes that attitudes toward market justice are shaped by a dynamic interplay between individuals' perceptions of economic inequality and their perceptions of meritocracy. Specifically, we argue that both perceptions independently and interactively influence the extent to which individuals endorse market-based distributions of social goods and services in Chile.

First, consistent with previous findings about redistributive preferences [@castillo_perception_2022], we expect that a higher perception of economic inequality will be associated with lower market justice preferences. When individuals perceive smaller income gaps, they are more likely to view market mechanisms as fair and legitimate, reinforcing the acceptance of outcomes based on competition and ability to pay. Conversely, a heightened perception of inequality may erode confidence in market fairness, weakening support for market-based distribution. This relationship is particularly relevant in the context of Chile, where the neoliberal economic model has been a dominant force in shaping public attitudes toward inequality and justice [@canalesceron_sujeto_2021; @mac-clure_justicia_2024].

Second, higher perceived meritocracy is expected to be positively associated with market justice preferences. Individuals who believe that effort and talent primarily determine success are more likely to justify unequal outcomes and endorse the notion that markets allocate resources fairly according to individual merit [@castillo_socialization_2024]. This aligns with the idea that meritocratic perceptions serve as a moral framework that legitimizes market-based inequalities, as individuals perceive the system as just when they believe that rewards are based on individual merit. This is particularly relevant in contexts where neoliberal ideologies dominate, as they often emphasize individual responsibility and competition as the basis for social order.

Third, we propose that perceptions of meritocracy and perceptions of economic inequality interact in shaping market justice attitudes. Specifically, we argue that the legitimizing effect of perceived meritocracy on market justice preferences is moderated by perceived economic inequality: when perceived inequality is low, the positive association between meritocratic perceptions and market justice preferences will be stronger. However, when perceived inequality is high, this association will weaken, as greater awareness of large economic gaps may challenge the view that outcomes are purely merit-based.

Additionally, this study examines whether these relationships vary over time, particularly in light of broader social and political transformations. We are especially interested in whether changes in individuals’ perceptions of inequality and meritocracy are associated with shifts in their market justice attitudes across different time points. The primary focus of this study is thus to scrutinize how individual-level changes in perceptions relate to changes in preferences over time. While we acknowledge the relevance of contextual events — such as the 2019 political outburst or the subsequent constitutional process — for public opinion and political orientation [@disipavlic_justification_2025], identifying causal effects of such events (e.g., mass protest exposure) on individual attitudes lies outside the scope of this research. Instead, we focus on how intra-individual trajectories in perceived inequality and merit shape evolving attitudes toward the commodification of social services.

Based on these arguments, we propose the following hypotheses:

$H1_{a} \; (\text{Between-person}):$ Individuals who perceive higher levels of economic inequality will report lower support for market justice preferences, on average, compared to those who perceive lower levels of inequality.

$H1_{b} \; (\text{Within-person}):$ Increases in an individual's perception of economic inequality over time will be associated with decreases in their support for market justice preferences.

$H2_{a} \; (\text{Between-person}):$ Individuals who endorse stronger meritocratic perceptions will report higher support for market justice preferences, on average, compared to those with weaker meritocratic perceptions

$H2_{b} \; (\text{Within-person}):$ Increases in an individual's meritocratic perceptions over time will be associated with increases in their support for market justice preferences.

$H3_{a} \; (\text{Between-person}):$ The positive association between meritocratic perceptions and market justice preferences will be weaker among individuals who perceive higher levels of economic inequality, compared to those who perceive lower inequality.

$H3_{b} \; (\text{Within-person}):$ The positive association between meritocratic perceptions and market justice preferences will be weaker at times when individuals perceive higher levels of inequality, relative to times when they perceive lower inequality.

$H4_{a} \; (\text{Growth moderation – Inequality}):$ The longitudinal association between perceived economic inequality and market justice preferences will strengthen over time, both within and between individuals.

$H4_{b} \; (\text{Growth moderation – Meritocracy}):$ The longitudinal association between meritocratic perceptions and market justice preferences will weaken over time, both within and between individuals.