### 2025-08-05 resubmission/01-introduction.qmd
review introduction & antecedentes

---
author: "Equipo EDUMER"
bibliography: "../input/bib/merit-edjust.bib"
csl: "../input/bib/apa6.csl"
---

# Introduction

Since 1980, economic inequality and wealth concentration have dramatically increased worldwide, becoming one of the main challenges for the social sciences. Globally, in 2021, less than 50% of the world population owned only 2% of the wealth, while the richest 10% concentrated 76%. Futhermore, the wealthiest 1% captured nearly 38% of total assets [@chancel_world_2022]. This context of economic disparity has sparked renewed interest in studying not only the objective aspects of inequality, such as income and access to resources, but also its subjective dimensions, including perceptions, beliefs, and associated attitudes [@janmaat_subjective_2013]. Such dimensions are deemed as key to understand the  delegitimation of political institutions that economic inequality can elicit [@castillo_perception_2022]. Therefore, understanding how people perceive inequality is crucial, as these perceptions can influence how societies comprehend and justify (or not) inequality in the distribution of goods and services, as well as their capacity to funnel political discontent and eventually opposition to the current economic distribution.

The justification of economic inequality is a topic with a long tradition in sociology, with seminal works such as the ones from Kluegel [-@kluegel_social_1995a; -@kluegel_beliefs_1986] and Lane [-@lane_political_1962; -@lane_market_1986]. Lane's main aim was to "help explain the tenacious hold of market capitalism on the public mind" [-@lane_market_1986, p.384], proposing the distinction between market and political justice. Whereas political justice refer to government actions aligned with justice principles such equality or need, market justice involves the allocation linked to earned desserts, particularly through individual contributions based on individual merit. Part of this agenda has focused on analyzing the extent to which individuals consider as just to get better access to social services (such as pensions, health and education) based on payment capacity [@busemeyer_skills_2014; @immergut_it_2020; @castillo_socialization_2024; @koos_moral_2019; @lindh_bringing_2023; @lindh_public_2015], usually under the label of market justice preferences. The empirical evidence to date suggests that a significant portion of the population justifies inequality in these areas [@koos_moral_2019], and that such preferences vary according to objective socioeconomic factors and countries contextual characteristics [@lindh_public_2015; @busemeyer_skills_2014]. Within this framework, we argue that there are two relevant factors that are usually sidelined in this literature and that are key to understand market justice preferences: the perception of economic inequality and the perception of meritocracy.

The perception of economic inequality can be understood as an individual's subjective assessment of how resources are allocated among members of a given society [@akyelken_urban_2020]. Regardless of their measurement, various studies have shown that this perception often underestimates the gap between the rich and the poor, revealing inconsistencies between objective and subjective aspects of inequality [@trump_income_2018]. Moreover, there is consistent evidence that perceived inequality, even under control of objective inequality, could have implications for attitudes toward the distribution of goods and services [@castillo_perception_2022; @schroder_income_2017], redistributive preferences, justification of inequality, and legitimacy of the economic system [@castillo_perception_2022; @garcia-sanchez_perceptions_2018; @garcia-sanchez_attitudes_2020]. From this perspective, in this paper we argue that perceptions of economic inequality would play a relevant role for understanding market justice preferences, as a lower perceived inequality could reinforce the idea that the differences in access to social services could be justified --- a link that becomes harder to hold when high inequality is perceived.

Along with perception of inequality, a second focus of this paper is the perception of meritocracy. Meritocracy posits that inequality can be legitimized through distributive criteria such as effort and talent [@young_rise_1962; @davis_principles_2001]. Previous studies have shown that those with stronger meritocratic perceptions tend to justify larger inequality, as economic differences are legitimized by being attributed to individual achievements [@mijs_paradox_2021; @wilson_role_2003; @batruch_belief_2023]. We argue that in contexts where the distribution of goods and services is predominantly governed by market logics, as in the chilean case [@madariaga_three_2020; @boccardo_30_2020], meritocratic perceptions could play a role in the preferences of market justice.

The primary objective of this study is to analyze the interplay between perceptions of inequality and meritocracy and their joint influence on preferences for market justice along time in Chile using the panel survey ELSOC (Social Chilean Longitudinal Study, 2016-2023). Furthermore, we explore how political and social contingencies---such as the 2019 and 2022 social movements---might have moderated these relationships by prompting more critical reflection on the commodification of social services [@somma_no_2021]. Recognizing the temporal dimension in shaping preferences is essential, given that historical and contextual factors that could challenge or reaffirm the way in which individuals perceive inequality and meritocracy, resulting in changes in market justice preferences.