### 2025-08-05 resubmission/05-discussion.qmd
Update + render
Creo que solo falta mirar en detalle la conclusión @juancarloscastillo y luego lo cerramos con un vistazo final

---
author: "Equipo EDUMER"
bibliography: "../input/bib/merit-edjust.bib"
csl: "../input/bib/apa6.csl"
---

# Discussion

This study tests several key hypotheses concerning the relationships between perceptions of economic inequality, meritocratic perceptions, and preferences toward market justice in Chile. It posits that higher perceived inequality and stronger meritocratic perceptions both independently contribute to greater support for market-based distributions of social goods. Furthermore, the study examines how perceived economic inequality moderates the influence of meritocratic perceptions on market justice preferences. Additionally, it considers how major social and political mobilizations, such as those occurring around 2019 and subsequent years, may have weakened these relationships, reflecting increased critical perspectives on systemic inequalities and market mechanisms. The subsequent discussion interprets the empirical findings in light of these hypotheses, considering their implications for understanding attitudes toward inequality and social justice in a context marked by profound social change.

The first set of hypothesis suggested that perceptions of economic inequality were strongly linked to preferences for market justice. The findings demonstrated that individuals perceiving higher income disparities showed less support for the idea that the access to core social services depends on individual income, aligning with theories emphasizing that awareness of inequality could fuel a critical view of such policies [@Castillo2012a_justice; @mijs_paradox_2021; @garcia-sanchez_vicious_2019].  Such a relationship reflects a moral economy in which perceptions of systemic unfairness undermine the legitimacy of existing distributions and fuel demands for greater equity. In particular, the negative within-individual effect observed over time ($H1_{b}$) suggests that as inequality becomes more salient for some individuals, their support for market justice declines—potentially due to rising distrust in market mechanisms or growing disillusionment with the fairness of the system. However, although the effect of perceived inequality at the between-individual level is also negative, it is not statistically significant. Therefore, we cannot conclude that individuals who, on average, perceive higher inequality are systematically less supportive of market justice. As a result, hypothesis $H1_{a}$ does not receive empirical support.

Regarding the second set of hypotheses, the results confirmed that meritocratic perceptions—particularly those emphasizing individual effort—were associated with stronger support for market-based distribution systems. Individuals who believed that success is primarily achieved through personal effort were more likely to justify unequal access to core social services based on income, interpreting such disparities as outcomes of individual merit rather than systemic injustice ($H2_{a}$). This finding aligns with previous research showing that meritocratic narratives serve as moral justifications that legitimize social stratification [@Castillo2012a_justice; @hoyt_mindsets_2023]. These perceptions operate symbolically to reinforce structural inequalities by reducing support for redistributive policies, framing inequality as both fair and deserved. Consistent with prior work by @castillo_socialization_2024 on Chilean students, the findings suggest that such meritocratic perceptions uphold existing hierarchies by promoting the acceptance of inequality as a reflection of individual virtue rather than structural failure. This mechanism is particularly salient in neoliberal contexts like Chile, where market logics heavily shape social attitudes [@canalesceron_sujeto_2021].

Interestingly, intra-individual changes in meritocratic perceptions over time ($H2_{b}$) reveal mixed effects. While increases in the belief that rewards are based on individual effort are associated with stronger preferences for market justice, increases in the belief that rewards derive from talent are linked to weaker support for such principles. One possible explanation is that effort is generally viewed as a controllable and malleable trait, aligning with the normative foundations of market fairness. In contrast, talent tends to be perceived as innate and less subject to personal control, rendering talent-based inequality less legitimate. Furthermore, increased exposure to real-world scenarios in which outcomes are clearly shaped by inherent traits rather than hard work may lead individuals to question the fairness and legitimacy of market-based reward systems.

The third set of hypotheses addressed the moderating role of perceptions of inequality in the relationship between meritocratic perceptions and preferences for market justice. The analysis revealed that the positive association between meritocratic perceptions and support for market-based distribution systems tends to be stronger when perceived economic inequality is high—contradicting our initial expectations ($H3_{a}$ and $H3_{b}$). Although this moderating effect is not statistically significant for effort-related meritocratic perceptions, it is significant for talent-related perceptions at both the within- and between-individual levels. This finding runs counter to our hypotheses, as it suggests that the effect of meritocratic perceptions—specifically, talent-based meritocracy—intensifies as perceptions of inequality increase. However, this interpretation requires caution. The main effect of perceived inequality on support for market justice is negative. Thus, as perceived inequality becomes less negative (i.e., closer to zero), the positive relationship between talent-based meritocratic perceptions and market justice preferences becomes stronger. This suggests a more nuanced dynamic: meritocratic perceptions, particularly those emphasizing talent, may serve as a stronger justificatory mechanism for market-based inequalities among individuals who are less inclined to perceive inequality as a problem. In other words, among those who do not strongly perceive systemic disparities, meritocratic narratives may play a more influential role in legitimizing unequal outcomes.

Finally, the fourth hypothesis examined the potential impact of recent social mobilizations—particularly those that erupted in 2019—on the relationship between meritocratic perceptions, perceptions of inequality, and preferences for market justice. The main effect of time indicates that, on average, support for market justice increased following the 2019 protests. However, the analysis did not detect any significant interactions between distributive beliefs and support for market justice, suggesting that the social mobilizations did not substantially modify the underlying associations between these variables. As a result, hypotheses $H4_{a}$ and $H4_{b}$ are not supported by the data. This may point to a certain stability—or even resilience—in the normative frameworks that guide individuals’ evaluations of distributive justice, despite the occurrence of major collective political events. Alternatively, while the protests may have eroded trust in institutional arrangements or governance, they may not have fundamentally altered individuals’ beliefs about how rewards and resources should be allocated in society.