### 2025-08-06 resubmission/02-antecedents.qmd
Up antecedentes meritocracia + render
@juancarloscastillo todo tuyo

# Market justice preferences

Conceptually, *market justice* has been discussed in the literature as a normative principle that legitimates the distribution of economic rewards based on individual merit. It is possible to trace the concept to the understanding of Lane [-@lane_market_1986], who makes a contrast between market justice and political justice. The author defines market justice as a system of "earned deserts", whereby individuals are seen as deserving of a determined distributive outcome due to their effort and skills. In contrast, political justice emphasizes principles of equality and need, which are often represented by the welfare state action through social policies. An important remark is that the principles of market justice rely on the assumption that markets are neutral and self-regulating arenas, where individuals are treated fairly because they face the same formal rules of engagement and procedures [@lane_market_1986]. Consequently, the legitimacy of market justice stems from the assumption that inequalities are not only inevitable but fair---so long as the rules are transparent and opportunities are open. In this way, market justice provides a moral justification for inequality by framing it as a necessary outcome of individual responsibility [@kluegel_legitimation_1999; @lindh_public_2015].

Empirical studies have shown different strategies for the study of market justice preferences. A common approach in the literature is to gauge attitudes toward the legitimacy of inequality in specific domains, especially when linked to income differences. This can be traced to the seminal work of Kluegel and Smith [-@kluegel_beliefs_1981] who assessed the normative foundations that explain public support for economic inequality. Over time, this approach has been extended beyond income to include other market-mediated outcomes, such as education, healthcare, and/or pensions. For example, @vondemknesebeck_are_2016 and Immergut and Schneider [-@immergut_it_2020] examine whether citizens consider it fair that individuals with higher incomes can access better healthcare, while Lee and Stacey [-@lee_fairness_2023] apply a similar method in the context of education in Australia. These studies usually rely on survey items asking respondents to evaluate the fairness of income-based access to welfare and social services[@svallfors_political_2007;@lindh_public_2015]. More recently, Castillo et al. [-@castillo_socialization_2024] introduced a single-item composite measure of market justice to assess student attitudes toward income-based access to education, healthcare, and pensions in Chile. Such empirical strategies aim to capture the extent to which individuals accept inequality when framed as a reflection of market outcomes.

The study of market justice preferences has increasingly focused on how they are shaped by individuals' socioeconomic position, normative beliefs, and the institutional context in which they are embedded [@pierson_when_1993]. Across the literature, there is empirical evidence suggesting that individuals in more advantaged socioeconomic positions---those with higher occupational class, income, and education---are more likely to support market justice principles [@svallfors_political_2007; @koos_moral_2019; @lindh_public_2015]. This tendency reflects not only material self-interest but also a broader moral economy, in which winners of the market system internalize justifications for the status quo. At the same time, political ideology also plays a role ---such as economically conservative values ---  where  right-wing individuals show higher support for meritocracy and more skepticism towards redistribution. This is particularly salient in countries with more restricted public provision of social services. For example, in cases such as Chile [@castillo_socialization_2024] and Australia[@lee_fairness_2023] the evidence suggests that right-wing individuals are more supportive of market-based distribution of welfare. Regarding contextual factors, in liberal welfare regimes like those of the United States or the United Kingdom, market justice preferences are more widespread, while in coordinated or social-democratic regimes---such as Sweden or Germany---citizens are generally more critical of market-based inequalities [@immergut_it_2020; @lindh_public_2015].

# Perceptions of inequality and meritocracy

_The perception of economic inequality_

Perceptions of inequality refer to individuals' subjective evaluations of the extent, causes, and consequences of income and wealth disparities [@akyelken_urban_2020]. Unlike objective measures such as the Gini index, perceived inequality captures how individuals make sense of distributive hierarchies in their everyday lives, shaped by reference groups, social comparisons, and information environments [@mijs_stratified_2016; @gimpelson_misperceiving_2018; @garcia-castro_perceiving_2020]. Scholars have proposed multiple dimensions of perceived inequality, including its magnitude (how significant are the gaps), vertical structure (between which groups), the trend over time (increasing or decreasing), and legitimacy (whether it is just or not) [@engelhardt_what_2018; @garcia-sanchez_vicious_2019]. These dimensions encompass both cognitive and normative aspects of perceptions of inequality and can vary across societies and social groups, depending on exposure, ideology, and personal experience [@castillo_perception_2022; @garcia-sanchez_perceptions_2018].

Perceptions of inequality have been associated with a range of distributive attitudes [@castillo_perception_2022; @garcia-sanchez_vicious_2019; @trump_income_2018; @mccall_exposure_2017]. Research indicates that perceptions of inequality influence redistributive preferences to a greater extent than objective or actual economic inequality [@gimpelson_misperceiving_2018; @Castillo2011]. Lower perceived inequality can reinforce support for market-based distributive arrangements [@kuhn_eye_2011]. In contrast, when inequality is perceived as excessive, individuals could question the legitimacy of market justice and become more supportive of redistributive policies [@garcia-sanchez_vicious_2019].

Assessing the perception of inequality in empirical research is an ongoing challenge. Specifically in survey research, the assessment of perceived inequality through Likert-type items such as "income differences are too large" have been questioned as they captures general concern or discomfort rather than a specific perception [@Castillo2011; @garcia-sanchez_vicious_2019]. As a result, the conceptual clarity between perceived inequality and inequality aversion remains blurred in many empirical studies. To address this limitation, recent work has emphasized the need to distinguish between absolute and comparative measures, as well as between ideal and perceived estimates of economic gaps [@garcia-sanchez_creencias_2022; @auspurg_why_2017]. Through asking perceived salaries for occupations in the extremes of the status continuum (such as a CEO and a manual worker), an indirect measure of perceived inequality is built by the ratio of the high and the low status occupation. This type of measure provides a concrete frame that respondents can relate to more easily than abstract questions about national income distribution [@castillo_percepcion_2019; @easterbrook_social_2021; @willis_legitimacy_2015]. Evidence to date shows that perceived wage gaps are strong predictors of political attitudes [@pedersen_attitudes_2019; @garcia-sanchez_perceptions_2018], making them a valuable tool for understanding public responses to economic disparities.

_Perception of meritocracy_

Meritocracy constitutes a central ideological framework for legitimizing various forms of social inequality [@mijs_unfulfillable_2016; @wilson_role_2003]. Rooted in the belief that rewards and social positions should be allocated based on individual effort and talent, meritocracy operates both as a descriptive account of how societies function and as a normative ideal of how they ought to function. As originally conceptualized by Michael Young [-@young_rise_1962], the term was intended as a critique of a system in which merit-based stratification gives rise to a new form of inequality. However, over time, meritocracy has come to be widely endorsed as a fair and desirable principle of distribution, particularly within liberal democracies and market-oriented societies [@mijs_paradox_2021; @sandel_tyranny_2020].

From a sociological perspective, belief in meritocracy extends beyond a cognitive judgment; it constitutes a moral lens through which individuals interpret social and economic disparities [@mijs_paradox_2021]. Individuals who believe that success results from hard work and talent are more likely to regard inequality as legitimate [@tejero-peregrina_perceived_2025; @castillo_percepcion_2019; @batruch_belief_2023]. In contrast, when outcomes are attributed to luck, social origin, or structural barriers, inequality is more often perceived as unjust [@reynolds_perceptions_2014; @hauser_misperceptions_2017]. This distinction becomes especially salient in societies marked by persistent structural inequality, where dominant narratives emphasize personal responsibility while neglecting the role of entrenched disadvantage.

Recent research has highlighted the importance of disaggregating the concept of "meritocratic beliefs," particularly by distinguishing between meritocratic preferences and meritocratic perceptions [@castillo_multidimensional_2023]. While preferences refer to normative ideals about how rewards should be allocated (i.e., based on merit), perceptions refer to subjective evaluations of how meritocracy actually operates in practice (i.e., how rewards are allocated) [@janmaat_subjective_2013]. This distinction allows for a more precise understanding of how individuals make sense of inequality: people may endorse meritocracy normatively while simultaneously recognizing its failure in practice. In stratified societies, the gap between preferences and perceptions is particularly evident, as individuals may continue to uphold the ideal of meritocracy despite growing awareness of unequal opportunities [@mijs_belief_2022].

Within the domain of meritocratic perceptions [@castillo_multidimensional_2023], we further distinguish between two core dimensions: effort-based and talent-based perceptions. This distinction is analytically important, as it captures different mechanisms through which individuals may justify inequality [@young_rise_1962; @reynolds_perceptions_2014]. Effort-based perceptions emphasize hard work and perseverance as the key drivers of success, aligning closely with cultural narratives of individual responsibility and moral deservingness [@dubet_repensar_2011]. In contrast, talent-based perceptions emphasize innate ability and intelligence—traits often regarded as less malleable and more unequally distributed [@young_rise_1962]. We argue that this latter dimension may shape attitudes differently, as talent tends to be seen as a given attribute rather than the result of personal effort.

Meritocratic perceptions function as a normative framework that legitimizes unequal outcomes, particularly when access to social goods is stratified by income or socioeconomic background. Prior studies have consistently shown that individuals who perceive society as meritocratic express lower support for redistribution [@tejero-peregrina_perceived_2025; @hoyt_mindsets_2023], greater legitimization of class inequality [@batruch_belief_2023], increased tolerance for inequality [@day_movin_2017], and lower perceptions of income inequality itself [@castillo_percepcion_2019]. In the regard of market justice, prior studies in Chile have shown that individuals who perceive higher levels of meritocracy tend to express stronger support for unequal distributions that reflect market outcomes in healthcare, education and pensions [@castillo_socialization_2024]. Taken together, and as @busemeyer_positive_2021 argues, meritocratic narratives can operate as feedback mechanisms that shape public opinion and individual well-being by framing welfare outcomes as either deserved or undeserved. This psychological mechanism underscores the normative power of meritocracy in stabilizing unequal systems by shaping political attitudes and personal understandings of success and failure.

## The Chilean context

Chile offers an interesting case for examining how public attitudes toward the distribution of social services evolve amid declining poverty and persistently high income inequality within a residual welfare framework [@ferre_welfare_2023]. Following the neoliberal reforms of the 1980s introduced during the civil-military dictatorship (1973-1989), the Chilean welfare state has become increasingly reliant on private provision, with essential services often commodified and largely reserved for those with sufficient purchasing power [@madariaga_three_2020; @boccardo_30_2020]. This model has favored the affluent, while lower-income groups are left dependent on limited and underfunded public alternatives. Research on social stratification characterizes Chile as a society where mobility into intermediate classes is feasible, but access to elite strata remains largely restricted [@lopez-roldan_comparative_2021; @torche_intergenerational_2014a]. While some indicators—such as education and occupation—suggest a relatively fluid class structure, income mobility remains constrained [@espinoza_movilidad_2014].

Despite steady economic expansion, Chile continues to rank among the most unequal nations in the OECD, marked by a high Gini coefficient and substantial wealth concentration at the top [@flores_top_2020]. In this context, Chile experienced a wave of mass protests from October 2019 to March 2020. Initially led by high school students, the demonstrations quickly gained momentum as broader segments of society mobilized around demands for more equality of opportunities, particularly regarding precarious access to welfare services. The political system interpreted this unrest as a call to reform existing political institutions, with a special focus on equalizing opportunities and providing more egalitarian access to welfare by reducing reliance on out-of-pocket mechanisms [@somma_no_2021].

Concerning public opinion, the ELSOC survey shows that Chile has a low level of agreement with the idea that it is fair for those with higher incomes to buy better welfare in the areas of education, healthcare, and pensions, with average support of 8.5% between 2016 and 2023. At the same time, around 55% and 47% of the public between 2016 and 2023 *disagreed* that effort and talent are rewarded accordingly. This coincides with 90% of citizens *agreeing* that income differences in Chile are *too large* [@castillo_changes_2025].

In sum, Chile represents an illustrative case of a particular form of *moral economy* [@koos_moral_2019], in which national socioeconomic characteristics, institutional arrangements, and public attitudes coexist. This suggests that the country combines persistently high levels of economic and social inequality with a strong institutionalization of market-based mechanisms within the welfare system [@madariaga_three_2020]. At the same time, perceptions of inequality are high, and there is widespread public criticism of the commodification of social services [@somma_no_2021]. For the above reasons, we argue that the Chilean case represents a relevant context to study the role of perceived income inequality and meritocracy in shaping public preferences regarding the commodification of welfare services, particularly from the perspective of individual change.

## This study

Building upon the previous literature, this study proposes that attitudes toward market justice are shaped by a dynamic interplay between individuals' perceptions of economic inequality and their perceptions of meritocracy. Specifically, we argue that both perceptions independently and interactively influence the extent to which individuals endorse market-based distributions of social goods and services in Chile.

First, consistent with previous findings about redistributive preferences [@castillo_perception_2022], we expect that a higher perception of economic inequality will be associated with lower market justice preferences. When individuals perceive smaller income gaps, they are more likely to view market mechanisms as fair and legitimate, reinforcing the acceptance of outcomes based on competition and ability to pay. Conversely, a heightened perception of inequality may erode confidence in market fairness, weakening support for market-based distribution. This relationship is particularly relevant in the context of Chile, where the neoliberal economic model has been a dominant force in shaping public attitudes toward inequality and justice [@canalesceron_sujeto_2021; @mac-clure_justicia_2024].

Second, higher perceived meritocracy is expected to be positively associated with market justice preferences. Individuals who believe that effort and talent primarily determine success would be more likely to justify unequal outcomes and endorse the notion that markets allocate resources fairly according to individual merit [@castillo_socialization_2024]. This aligns with the idea that meritocratic perceptions serve as a moral framework that legitimizes market-based inequalities, as individuals perceive the system as just when they believe that rewards are based on individual merit. This is particularly relevant in contexts where neoliberal ideologies dominate, as they often emphasize individual responsibility and competition as the basis for social order.

Third, we propose that perceptions of meritocracy and perceptions of economic inequality interact in shaping market justice attitudes. Specifically, we argue that the legitimizing effect of perceived meritocracy on market justice preferences is moderated by perceived economic inequality: when perceived inequality is low, the positive association between meritocratic perceptions and market justice preferences will be stronger. However, when perceived inequality is high, this association will weaken, as greater awareness of large economic gaps may challenge the view that outcomes are purely merit-based.

Additionally, this study examines whether these relationships vary over time, particularly in light of broader social and political transformations. We are especially interested in whether changes in individuals’ perceptions of inequality and meritocracy are associated with shifts in their market justice attitudes across different time points. While we acknowledge the relevance of contextual events — such as the 2019 political outburst or the subsequent constitutional process — for public opinion and political orientation [@disipavlic_justification_2025], identifying causal effects of such events (e.g., mass protest exposure) on individual attitudes lies outside the scope of this research. Instead, we focus on how intra-individual trajectories in perceived inequality and merit shape evolving attitudes toward the commodification of social services.

Based on these arguments, we propose the following hypotheses:

$H1_{a} \; (\text{Between-person}):$ Individuals who perceive higher levels of economic inequality will report lower support for market justice preferences, on average, compared to those who perceive lower levels of inequality.

$H1_{b} \; (\text{Within-person}):$ Increases in an individual's perception of economic inequality over time will be associated with decreases in their support for market justice preferences.

$H2_{a} \; (\text{Between-person}):$ Individuals who endorse stronger meritocratic perceptions will report higher support for market justice preferences, on average, compared to those with weaker meritocratic perceptions

$H2_{b} \; (\text{Within-person}):$ Increases in an individual's meritocratic perceptions over time will be associated with increases in their support for market justice preferences.

$H3_{a} \; (\text{Between-person}):$ The positive association between meritocratic perceptions and market justice preferences will be weaker among individuals who perceive higher levels of economic inequality, compared to those who perceive lower inequality.

$H3_{b} \; (\text{Within-person}):$ The positive association between meritocratic perceptions and market justice preferences will be weaker at times when individuals perceive higher levels of inequality, relative to times when they perceive lower inequality.

$H4_{a} \; (\text{Growth moderation – Inequality}):$ The longitudinal association between perceived economic inequality and market justice preferences will strengthen over time, both within and between individuals.

$H4_{b} \; (\text{Growth moderation – Meritocracy}):$ The longitudinal association between meritocratic perceptions and market justice preferences will weaken over time, both within and between individuals.