### 2025-08-08 resubmission/06-conclusion.qmd

---
author: "Equipo EDUMER"
bibliography: "../input/bib/merit-edjust.bib"
csl: "../input/bib/apa6.csl"
---

# Conclusions

This study examined the complex interplay between perceptions of economic inequality, meritocratic perceptions, and preferences toward market justice in Chile from 2016 to 2023, drawing on longitudinal data from the ELSOC survey. By exploring how subjective assessments and social contexts influence support for redistribution and market-based resource allocation, the research offers different elements that contribute to the understanding of the normative foundations underpinning social justice attitudes in a highly unequal and commodified environment.

The findings support that higher perceptions of economic inequality are associated with less support for market justice preferences, this is, the belief that it is fair that those with higher income have better social services such as education, pensions and health. At the same time, meritocratic perceptions — particularly those emphasizing individual effort — are strongly associated with support for market-based distributions, suggesting that meritocracy serve as a moral justification for structural inequalities. However, changes in meritocratic perceptions over time reveal a more nuanced picture: while increased emphasis on effort reinforces support for market justice, increased emphasis on talent tends to reduce it. This distinction likely reflects broader beliefs about the controllability and fairness of different meritocratic traits. Moreover, the interaction between inequality perceptions and meritocratic perceptions indicates that the legitimizing power of meritocracy becomes stronger as perceptions of inequality become less negative — that is, when individuals are less aware or concerned about economic disparities. Such finding highlights a potential feedback mechanism by which lower sensitivity to inequality may enable stronger endorsement of merit-based explanations for unequal outcomes.

This research advances the extant literature by integrating subjective perceptions with social and political contexts to explain attitudes toward economic inequality and distributional justice. While previous studies primarily focused on objective measures or individual characteristics [@busemeyer_skills_2014; @immergut_it_2020; @lindh_public_2015], this work emphasizes the dynamic and interactional nature of perceptions and beliefs over time. Furthermore, it highlights the importance of socio-political upheavals in reshaping normative attitudes, underscoring the role of collective action in challenging entrenched narratives of meritocracy and fairness. The longitudinal approach provides a deeper temporal perspective on how societal events could influence individual perceptions and preferences.

Regarding avenues for future research, experimental evidence could help expand the understanding of the association between perceptual variables and market justice preferences. International comparative studies are also necessary in order to assess the role of national contexts and the universality or specificity of these dynamics. Additionally, investigating the role of media, political communication, and education in shaping perceptions of inequality and meritocracy would deepen understanding of the normative foundations of social justice attitudes and their variation over time. Finally, examining how these perceptions influence behavioral outcomes, such as political participation or support for social movements, would provide valuable insights into the pathways from beliefs to collective action and policy change.