### 2025-08-08 resubmission/05-discussion.qmd

---
author: "Equipo EDUMER"
bibliography: "../input/bib/merit-edjust.bib"
csl: "../input/bib/apa6.csl"
---

# Discussion

The first set of hypotheses proposed that perceptions of economic inequality are strongly associated with preferences for market justice. In the between-individual specification, the effect of perceived inequality is not statistically significant. Therefore, we cannot conclude that individuals who, on average, perceive higher levels of inequality are systematically less supportive of market justice. As a result, hypothesis $H1_{a}$ is not supported by the data. However, in the within-individual specification, the findings reveal a significant negative association: individuals who, at a given point in time, perceive higher income disparities are less supportive of the idea that access to core social services should depend on individual income. This result aligns with theories suggesting that increased awareness of inequality fosters a more critical stance toward market-based distributive arrangements [@Castillo2012a_justice; @mijs_paradox_2021; @garcia-sanchez_vicious_2019]. Such a pattern reflects a moral economy logic in which perceptions of systemic unfairness undermine the legitimacy of existing distributions and strengthen demands for greater equity. Specifically, the negative within-individual effect over time ($H1_{b}$) suggests that as inequality becomes more salient for some individuals, their support for market justice declines—possibly due to growing distrust in market mechanisms or increasing disillusionment with the perceived fairness of the system.

Regarding the second set of hypotheses, the results confirmed that meritocratic perceptions—particularly those emphasizing individual effort—were associated with stronger support for market-based distribution systems. Individuals who believed that success is primarily achieved through personal effort were more likely to justify unequal access to core social services based on income, interpreting such disparities as outcomes of individual merit rather than systemic injustice ($H2_{a}$). This finding aligns with previous research showing that meritocratic narratives serve as moral justifications that legitimize social stratification [@Castillo2012a_justice; @hoyt_mindsets_2023]. These perceptions operate symbolically to reinforce structural inequalities by reducing support for redistributive policies, framing inequality as both fair and deserved. Consistent with prior work by @castillo_socialization_2024 on Chilean students, the results suggest that such meritocratic perceptions uphold existing hierarchies by promoting the acceptance of inequality as a reflection of individual virtue rather than structural failure. This mechanism is particularly salient in neoliberal contexts like Chile, where market logics heavily shape social attitudes [@canalesceron_sujeto_2021].

Interestingly, intra-individual changes in meritocratic perceptions over time ($H2_{b}$) reveal mixed effects. While increases in the belief that rewards are based on individual effort are associated with stronger preferences for market justice, increases in the belief that rewards derive from talent are linked to weaker support for such principles. One possible explanation is that effort is generally viewed as a controllable and malleable trait, whereas talent tends to be perceived as innate and less subject to personal control, rendering talent-based inequality less legitimate. Furthermore, increased exposure to real-world scenarios in which outcomes are clearly shaped by inherent traits rather than hard work may lead individuals to question the fairness and legitimacy of market-based reward systems.

The third set of hypotheses addressed the moderating role of perceptions of inequality in the relationship between meritocratic perceptions and preferences for market justice. The analysis revealed that the positive association between meritocratic perceptions and support for market-based distribution systems tends to be stronger when perceived economic inequality is high, contradicting our initial expectations ($H3_{a}$ and $H3_{b}$). Although this moderating effect is not statistically significant for effort-related meritocratic perceptions, it is significant for talent-related perceptions at both the within- and between-individual levels. However, this interpretation requires caution: the main effect of perceived inequality on support for market justice is negative, meaning that as perceived inequality becomes less negative (i.e., closer to zero), the positive relationship between talent-based meritocratic perceptions and market justice preferences becomes stronger. This suggests a more nuanced dynamic: meritocratic perceptions, particularly those emphasizing talent, may serve as a stronger justificatory mechanism for market-based inequalities among individuals who are less inclined to perceive inequality as a problem. In other words, among those who do not strongly perceive systemic disparities, meritocratic narratives may play a more influential role in legitimizing unequal outcomes.

Finally, the fourth exploratory hypothesis examined the potential impact of social events — particularly those that erupted in 2019 — on the relationship between meritocratic perceptions, perceptions of inequality, and preferences for market justice. The main effect of time indicates that, on average, support for market justice increased after 2019 protests. However, the analysis did not detect any significant interactions between distributive beliefs and support for market justice, suggesting that events in this period (as the mobilizations, the pandemic and/or the constitucional processes) did not substantially modify the underlying associations between these variables. As a result, hypotheses $H4_{a}$ and $H4_{b}$ are not supported by the analysis. This may point to a certain stability or even resilience in the normative frameworks that guide individuals’ evaluations of distributive justice, despite the occurrence of major collective political events. Alternatively, while the protests may have eroded trust in institutional arrangements or governance, they may not have fundamentally altered individuals’ beliefs about how rewards and resources should be allocated in society.